### 2015-09-28 paper.tex
Initial version of journal paper

% tADRguide.tex
% v1.0 released January 2013

\documentclass{tADR2e}

% The following packages can be found on http:\\www.ctan.org
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{algorithm}
\usepackage{algpseudocode}  %\usepackage[noend]{algpseudocode}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{proof}
\newtheorem{property}{Property}
\usepackage[numbers]{natbib}

\makeatletter
%\let\NAT@parse\undefined % allow hyperref to work with ieeeconf.cls for bib numbers
\makeatother
\usepackage{hyperref}
%\usepackage{cite}

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\pii{\mathbf{p_{i,i+1}}}
\newcommand\pij{\mathbf{p_{i,j}}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}}
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
\newcommand\tcolli{t_{coll\ i}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\graphicspath{{images/}}

\jvol{00} \jnum{00} \jyear{2013} \jmonth{January}

\articletype{GUIDE}

\title{{\itshape Advanced Robotics} A simple path optimization method for motion planning }

\author{Myl\`{e}ne Campana$^{a,b}$ $^{\ast}$ \thanks{$^\ast$Corresponding author. Email: mcampana@laas.fr}, Florent Lamiraux$^{a,b}$ and Jean-Paul Laumond$^{a,b}$
\\\vspace{6pt}
$^{a}${\em{LAAS-CNRS, Toulouse, France}};
$^{b}${\em{ This work has been supported by the project ERC Advanced Grant 340050 Actanthrope and by the FP~7 project Factory in a Day under grant agreement n°~609206 }}
}
\maketitle

\begin{abstract}
Most algorithms in probabilistic sampling-based path planning compute 
collision-free paths made of straight line segments lying in the configuration 
space. Due
to the randomness of sampling, the paths make detours that need to be optimized.
The contribution of this paper is to propose a gradient-based algorithm that
transforms a polygonal collision-free path into a shorter one, while both:
\begin{itemize}
\item requiring mainly collision checking, and few time-consuming obstacle 
distance computation,
\item constraining only part of the configuration variables that may cause a 
collision, and not the entire configurations, and
\item reducing parasite motions that are not useful for the problem resolution.
\end{itemize}
The algorithm is simple and requires few parameter tuning. Experimental
results include navigation and manipulation tasks, e.g. an robotic arm 
manipulating throw a window and a PR2 robot working in a kitchen environment, 
and comparisons with a random shortcut optimizer.

\medskip

\begin{keywords}path optimization; motion planning; robotics
\end{keywords}\medskip

\end{abstract}

\section{Introduction}

Motion planning for systems in cluttered environments has been addressed for more
than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
today randomly sample the system configuration space~\cite{ref-spaceconfig} in
order to find a collision-free path. The main issue using these techniques is
 that the computed path makes unnecessary detours and needs to be post-processed 
 before being executed by a virtual or real robot. Alternative
strategies exist however.
\begin{itemize}
\item Planning by path-optimization~\cite{itomp2012},~\cite{voronoiOMP} where
obstacle avoidance is handled by constraints or cost using computation of the
nearest obstacle distance. Most of these planners are using non-linear
optimization~\cite{BettsNonlinopt} under constraints.

Such planners provide close-to-optimality paths and have smaller time computation
for easy problems, but they are mostly unable to solve narrow passage issues.

\item Optimal random sampling~\cite{KaramanPRMstarRRTstar} are also close to an
optimal solution, but computation time is significantly higher than classical
approaches. Moreover the roadmap is only valid for a given motion planning
   problem. 
\end{itemize}

In this paper, we propose a method aimed at shortening path length after a path
planning step. Note that we do not address path planning, but that we take the
result of a probabilistic motion planner as the input to our path optimization 
method.

For this shortening purpose, random shortcut methods are
still very popular~\cite{Sekhavat-Svestka1998,HauserFastSmooth}. However, random 
shortcut requires fine tuning of the termination condition (see 
Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long trajectories where only a small part needs to be optimized. Figure~\ref{decoupled_DOF_optimization} presents another situation where random shortcut will always fail to optimize the initial path, since it cannot decouple the robot degrees of freedom (DOF) on which the optimization occurs, contrary to our algorithm.

\begin{figure}[H]
	\centering
	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
	\caption{Example of a path which random shortcut will never manage to 
	optimize: each shortcut tentative will provide a collision or will not 
	decrease the path length. The optimal path belongs to the $x-y$ plane 
	containing $q_{init}$ and $q_{end}$.}
	\label{decoupled_DOF_optimization}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[height=4.9cm,width=1.8cm]{pr2_initial_final_vertical.png}
	\includegraphics[width=5.7cm]{p0_pr2_alone_merged3.png}
	\includegraphics[height=4.9cm]{p1RS_pr2_alone_merged3.png}
	\includegraphics[height=4.9cm]{p1GB_pr2_alone_merged3.png}\\
	\caption{In this motion planning problem, PR2 robot has just to 
exchange the positions of its arms. The task is simple, however, in absence of 
explicit indication, any probabilistic motion planner will compute a path that 
makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
algorithm (left). Path optimization is expected to remove unnecessary motions. 
The popular random shortcut algorithm fails in this case (middle). Our algorithm 
succeeds (right).}
	\label{pr2-final}
\end{figure}

On the other hand, numerical
optimization methods like CHOMP~\cite{RatliffChomp} can be used as a
post-processing step. They have clear termination conditions, but collision
avoidance is handled by inequality constraints sampled at many points along
the trajectory. These methods therefore require a pre-processing step of the
robot (and/or environment) model in order to make 
it simpler:~\cite{RatliffChomp} covers PR2 bodies with spheres, 
while~\cite{convexOptimMotplan} needs to decompose objects into convex subsets.
Moreover, in the applications we address, optimality is not desirable as
such since the shortest path between two configurations in a cluttered
environment usually contains contacts with the obstacle,
or is most of the time
out of the scope of numerical optimization algorithms~\cite{LaumondOptim}.

The idea of our method is to find a good trade-off between
the simplicity of “blind” methods like shortcut algorithm,
and the complexity of distance based optimization techniques.
The method iteratively shortens the initial path with gradient-based information.
When a collision is detected at a given iteration, the method backtracks to the
latest valid iteration and inserts a linearized transformation constraint
between the objects detected in collision. Only few distances between objects are evaluated, therefore no pre-processing of the
robot or environment models is necessary. The underlying optimization
algorithm is a quadratic program.

Related work is presented in Section II. Section III explains how the 
path-optimizer works, from the formulation of the problem to the implemented
algorithm. Finally, we conclude on experimental results in Section IV.

\section{RELATED WORK}

CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
input. It minimizes a time invariant cost function using efficient covariant
hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
high velocities) and an obstacle avoidance term, provided by the distance to the 
nearest obstacle for each iteration of the trajectory. Calculating these nearest 
distances however is time-consuming because the distances between all pairs of 
objects must be computed at each time step along the path. To reduce the 
computation time, the method starts by building offline a map of distances that 
will be called during the optimization at the requested time. Besides, meshes 
are pre-processed into bounding spheres so that distances are computed faster 
at the cost of a geometry approximation.

STOMP method~\cite{KalakrishnanStomp} avoids to compute an 
explicit gradient for cost optimization using a stochastic analysis of local 
random samples. But as for CHOMP, the obstacle cost term requires a voxel map to 
perform its Euclidian Distance Transforms, and represents the robot bodies with 
overlapping spheres. Such technique provides lots of distance and penetration 
information but remains very time consuming and is not as precise as some 
distance computation techniques based on the problem meshes as 
Gilbert-Johnson-Keerthi~\cite{gilbertGjk}.

Some optimization-based planners may not require an initial guess but some naive 
straight-line manually or randomly-sampled initialization as 
TrajOp~\cite{SchulmanConvexOptim}. Its trajectory is iteratively optimized with 
sequential convex optimization by minimizing at each step its square length, 
linear and non-linear constraints considered as penalties. To compute the 
collision-constraints, nearest obstacle distances are calculated at each discrete 
time of the trajectory vector. This can be a burden for a high-dimensional robot 
or a complex environment as we propose to use, and may be compensated with a 
short path composed of only one or two waypoints.

The elastic strips framework~\cite{BrockElasticStrips} is also an optimization 
based method. The path is modeled as a spring and obstacles give rise to a 
repulsive potential field. Although designed for on-line control purposes, this 
method may be used for path shortening. In this case however, the number of 
distance computation is very high. The authors also proposed to approximate the 
robot geometry by spheres.

Some heuristics use random shortcuts on the initial guess combined with a 
trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
are time-optimal since they comply with acceleration and velocity constraints.

In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
since this latter method relies on collision checking and backtracks when an 
iteration is detected in collision, instead of trying to constantly satisfy 
distance constraints. Unlike our method however, the iterations are composed of 
Cubic-B-splines. The benefit of this method is that the result is a 
differentiable path.

TO DESCRIBE:

~\cite{Guernane2011} (guiding shoots) is ``locally optimal'', not addressing high-DOF problems,

~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide the shortcut.

\section{PATH OPTIMIZATION} \label{section:path_optim}

\subsection {Kinematic chain}

A robot is defined by a kinematic chain composed of a tree of joints. We denote 
by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
is the dimension of $J_i$ in the configuration space, to the space 
of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
in the frame of its parent. In the examples shown in this paper, we consider 4 
types of joints described in Table~\ref{tab:joints}.
\begin{table}[b]
\centerline {
  \begin{tabular}{cccc}
    Name & dimension & config space & velocity\\
    \hline
    translation & 1 & $\real$ & $\real$\\
    bounded rotation & 1& $\real$ & $\real$\\
    unbounded rotation & 2 & $\Sone\subset\real^2$ & $\real$\\
    $SO(3)$ & 4 & $\Sthree\subset\real^4$ & $\real^3$
  \end{tabular}
}
\caption {Translation and rotation joint position are defined by 1 
parameter corresponding respectively to the translation along an axis and a 
rotation angle around an axis. Unbounded rotation is defined by a point on the 
unit circle of the plane: 2 parameters corresponding to the cosine and the 
sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
velocity of translation and unbounded rotation joints is the derivative of the 
configuration variable. The velocity of an unbounded rotation joint corresponds 
to the angular velocity. The velocity of a $SO(3)$ joint is defined by the 
angular velocity vector $\omega\in\real^3$.}
\label{tab:joints}
\end{table}
A configuration of the robot
$$\conf = (\underbrace{q_1,\cdots,q_{n_1}}_{J_1},\underbrace{q_{n_1+1},
\cdots,q_{n_1+n_2}}_{J_2},\cdots q_n),\ n\triangleq\sum_{i=1}^m n_i$$
is defined by the concatenation of the joint configurations. The configuration 
space of the robot is denoted by $\CS\subset\real^n$.

Note that the configuration of the robot belongs to a sub-manifold of $\real^n$.

The velocity of each joint $J_i$, $1\leq i \leq m$ is defined by a vector of $
\real^{p_i}$, where $p_i$ is the number of DOF of $J_i$. Note that 
the velocity vector does not necessarily have the same dimension as the 
configuration vector.

The velocity of the robot is defined as the concatenation of the velocities of 
each joint.
$$\dot{\conf} = (\underbrace{\dot{q}_{1},\cdots,\dot{q}_{p_1}}_{J_1},
\underbrace{\dot{q}_{p_1+1},\cdots,\dot{q}_{p_1+p_2}}_{J_2},\cdots \dot{q}_p),\ p
\triangleq\sum_{i=1}^m p_i$$

\paragraph {Operations on configurations and vectors} by analogy with the case 
where the configuration space is a vector space, we define the following 
operators between configurations and vectors:
$$
\conf_2 - \conf_1 \in \real^p, \ \conf_1, \conf_2\in\CS
$$
is the constant velocity moving from $\conf_1$ to $\conf_2$ in unit time, and
$$
\conf + \dot{\conf}\in\CS, \ \conf\in\CS \ \dot{\conf}\in\real^p
$$
is the configuration reached from $\conf$ after following constant velocity $
\dot{\conf}$ during unit time.

Note that the definitions above stem from the Riemanian structure of the 
configuration space of the robot. The above sum corresponds to the exponential 
map. We do not have enough space in this paper to develop the theory in a more 
rigorous way. The reader can easily state that ``following a 
constant velocity'' makes sense for the four types of joints defined in 
Table~\ref{tab:joints}. We refer to~\cite{riemanian-optim2008} Chapter~5 for 
details about Riemanian geometry.

\subsection {Straight interpolation}

Let $\conf_1, \conf_2\in\CS$ be two configurations. We define the straight 
interpolation between $\conf_1$ and $\conf_2$ as the curve in $\CS$ defined on 
interval $[0,1]$ by:
$$
t \rightarrow \conf_1 + t (\conf_2 - \conf_1)
$$
This interpolation corresponds to the linear interpolation for translation and 
bounded rotations, to the shortest arc on $\Sone$ for unbounded rotation and to 
the so called slerp interpolation for $SO(3)$.

\subsection{Problem definition}

We consider as input a path composed of a concatenation of straight 
interpolations between $wp+2$ configurations: $(\conf_0, \conf_1,\cdots,\conf_{wp
+1})$. This path is the output of a random sampling path planning algorithm 
between $\conf_0$ and $\conf_{wp+1}$. We wish to find a sequence of waypoints $
\conf_{1}$,...,$\conf_{wp}$ such that the new path $(\conf_0, \conf_1,\cdots,
\conf_{wp+1})$ is shorter and collision-free. Note that $\conf_0$ and $\conf_{wp
+1}$ are unchanged and that the workspace of the robot contains obstacles. We 
denote by $\xx$ the optimization variable:
$$
\xx \triangleq (\conf_1,\cdots,\conf_{wp})
$$

\subsubsection {Cost}

let $W\in\real^{p\times p}$ be a diagonal matrix of weights:
$$
W=\left(\begin{array}{cccccccccc}
w_1 I_{p_1}       &        &  0  \\
    & w_2 I_{p_2} &        &   \\
    &            & \ddots &   \\
  0 &            &        & w_m I_{p_m}
\end{array}\right)
$$
where $I_{p_i}$ is the identity matrix of size $p_i$ and $w_i$ is the weight 
associated to joint $J_i$. We define the length of the straight interpolation 
between two configurations as:
$$
\|\conf_2 - \conf_1\|_{W} \triangleq \sqrt{(\conf_2 - \conf_1)^T W^2 (\conf_2 - 
\conf_1)}.
$$
Weights are used to homogenize translations and rotations in the velocity vector. 
For rotations, the weight is equal to the maximal distance of the robot bodies 
moved by the joint to the center of the joint.

Given $\conf_0$ and $\conf_{wp+1}$ fixed, the cost we want to minimize is defined 
by
$$
\cost (\xx) \triangleq \frac{1}{2}\sum_{i=1}^{wp+1} \lambda_{i-1} \|\conf_{i}-\conf_{i-1}\|_{W}
^{2}
$$
where the $\lambda_{i-1}$ coefficients benefit will be explained in the results 
section. For now we can assume that $\forall i,\, \lambda_{i-1}=1$.

Note that $\cost$ is not exactly the length of the path, but it can be 
established that minimal length paths also minimize $\cost$. This latter cost is 
better conditioned for optimization purposes.

\subsection {Resolution}
We assume that the direct interpolation between the initial and final configurations contains collisions.
Let $H$ denotes the constant Hessian of the cost function, an iteration is 
described as follow:
\begin{equation}\label{eq:iteration-1}
\begin{split}
& \pii =  -H^{-1} \nabla c(\xx_i)^{T} \\
& \xx_{i+1} =  \xx_{i} + \alpha_i \pii
\end{split} 
\end{equation}
where $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
between $q_0$ and $q_{n+1}$. Since this solution is in collision, we set 
$\alpha_i = \alpha$ 
where $\alpha$ is a parameter that will be explained in the algorithm section.

We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. 
When a collision is detected, we introduce a constraint and perform a new 
iteration from $\xx_i$ as explained in the next section. We use a continuous 
collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in Algorithm~\ref{algo:gradient}.

\subsection{Constraints}

\begin{figure}[b]
	\centering
	\includegraphics[width=8cm]{optim_grad.pdf}
	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
	appears 
	to be in collision with the obstacle, the first colliding configuration 
	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The corresponding constraint will be computed in configuration $\xx_{i}(\tcolli)$.}
	\label{optim_grad}
\end{figure}

Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. Let 
us denote by $\tcolli$ the abscissa of the first collision detected on path 
$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free (see 
Figure~\ref{optim_grad}). Thus in 
configuration $\xx_{i+1}(\tcolli)$ a collision has been 
detected. Two cases are possible:
\begin{enumerate}
\item the collision occurred between two bodies of the robot: $\body_1$ and $
\body_2$, or
\item the collision occurred between a body of the robot $\body_1$ and the 
environment.
\end{enumerate}
In the rest of this section, we will only consider the first case. Reasoning about the second case is similar, except that the 
constraint is on the transformation of $\body_1$ with respect to the environment.

\begin {figure}
  \centerline {
    \def\svgwidth {\linewidth}
    \graphicspath {{./images/}}
    \input {images/transformation-constraint.pdf_tex}
  }
  \caption {Transformation constraint notations.}
  \label{fig:transformation-constraint}
\end {figure}
Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$. Let $R\in SO(3)$ (resp. $R^{*}$) and $\translation\in\real^3$ (resp. $\translation^{*}$) be the rotation and translation parts of $M$ (resp. $M^{*}$).
See also Figure~\ref{fig:transformation-constraint} for notations.

We denote by $F$ the mapping from $\CS^{wp}$ to $\real^6$ defined by
\begin{equation}\label{eq:constraint}
F (\xx) = \left(\begin{array}{c} R^T (\translation^{*} - \translation) \\
  \log (R^T R^{*})\end{array}\right)
\end{equation}
$\log (R^T R^{*})$ represents here a vector $\mathbf{v}$ of $\real^3$ such that
$R^T R^{*}$ is the rotation of axis $\mathbf{v}$ and of angle $\|\mathbf{v}\|$.
Note that $F$ vanishes for any path $\xx$ such that in configuration
$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $M^{*}$.
Let $r$ denotes the \textit{radius} of the bounding sphere of $\body_2$ centered at the origin of $\body_2$ local frame. Let us denote by $d$ the minimal distance between $\body_1$ and $\body_2$
in configuration $\xx_i(\tcolli)$. We have the following property.
\begin {property} \label {prop:collision free}
Let $\xx\in\CS^{wp}$ be a path and let us denote by $\translation$ and $\mathbf{v}$ respectively the 3 first coordinates and the 3 last coordinates of $F(\xx)$. If
$$
\|\translation\| + r \|\mathbf{v}\| \leq d
$$
then in configuration $\xx(\tcolli)$, $\body_1$ and $\body_2$ are collision-free.
\end {property}
The proof is straightforward: expressed in reference frame of $\body_1$, the motion of $\body_2$ between configurations $\xx(\tcolli)$ and $\xx_i(\tcolli)$ is a translation of vector $\translation$ followed by a rotation of angle $\|\mathbf{v}\|$. No point of $\body_2$ has moved by more than $\|\translation\| + r \|\mathbf{v}\|$. Therefore no collision is possible.

When a collision is detected at $\tcolli$ along path $\xx_{i+1}$, we wish to add
constraint
\begin{equation}\label{eq:non-linear constraint}
F (\xx) = 0
\end{equation}
However, as this constraint is non-linear, we linearize it around $\xx_i$:
$$
\frac{\partial F}{\partial \xx}(\xx_i)(\xx -\xx_i) = 0
$$
for any later iteration $\xx_j$, $j\geq i+1$.

We refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
programs (QP). This corresponds to \texttt{computeIterate} in Algorithm~\ref{algo:gradient}.

\subsection{Relinearization of constraints}

As constraint~(\ref{eq:non-linear constraint}) is not exactly enforced, but linearized, a new collision may appear at $\tcolli$ on a later iteration between
$\body_1$ and $\body_2$. To avoid this undesired effect, we check the values of all the
constraints at each iteration. The constraints that fail to satisfy
Property~\ref{prop:collision free} are relinearized around the latest valid path $\xx_k$, $ k > i+1$, as follows:
$$
\frac{\partial F}{\partial \xx}(\xx_k)(\xx -\xx_k) = -F(\xx_k)
$$
This step is performed by \texttt{solveConstraints} in Algorithm~\ref{algo:gradient}.

\subsection{Algorithm}

In this part we describe the path-optimizer Algorithm~\ref{algo:gradient} 
according to the previous step definitions. The main difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
be traveled through, and also to return a collision-free path. Typically, we 
chose $\alpha_{init} = 0.2$ to process small steps. $\alpha_{init} = 0.5$ is more 
encountered in the optimization literature, procuring larger steps but with more 
risks of collision.

\begin{algorithm}
\begin{algorithmic}[1] % line numerotation
\Procedure{Optimize}{$\xx_0$}
	\State $\alpha \gets \alpha_{init}$; $\epsilon \gets 10^{-3}$
	\State $minReached \gets false$
	\State $validConstraints \gets true$
	\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
		\State $\p = \texttt{computeIterate}()$
		\State $minReached = (||\p||<\epsilon \textbf{ or } \alpha=1)$\label{algo:minReached}
		\State $\xx_1 \gets \xx_0 + \alpha\,\p$
		\If{($\alpha \neq 1$)}
			%\State \texttt{solveConstraints($\xx_1$)}
			\If{(\textbf{not}(\texttt{solveConstraints($\xx_1$)}))} \label{algo:solveCons}
				\State $\alpha \gets \alpha*0.5$
				\State $validConstraints \gets false$\label{algo:solveCons2}
			\Else
				\State $validConstraints \gets true$
			\EndIf
		\EndIf
		\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
			\State $noCollision \gets false$
			\If{$(\alpha \neq 1)$}
				\State \texttt{addCollisionConstraints}()\label{algo:addCons}
				\State $\alpha \gets 1$
			\Else
				\If{($validConstraints$)}\label{algo:towardMin}
					\State $\alpha \gets \alpha_{init}$\label{algo:towardMin2}
				\EndIf
			\EndIf
		\Else
			\State $\xx_0 \gets \xx_1$
			\State $noCollision \gets true$\label{algo:noCollision}
			\State $\alpha \gets 0.5*(1+\alpha)$
		\EndIf
	\EndWhile
	\Return $\xx_0$
\EndProcedure
\end{algorithmic}
\caption{Gradient-based path-optimization.} \label{algo:gradient}
\end{algorithm}

Each time a collision-constraint is added, 
the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution (lines~\ref{algo:minReached}-\ref{algo:noCollision}). Otherwise, 
smaller steps are iteratively applied and tested toward the minimum (lines~\ref{algo:towardMin}-\ref{algo:towardMin2}), and constraints are added each time a collision is found (line~\ref{algo:addCons}).

If the constraint relinearization fails to validate Property~\ref{prop:collision free}, $\alpha$ is halved and the algorithm restarts from the latest valid path (lines~\ref{algo:solveCons}-\ref{algo:solveCons2}).

\section{RESULTS}

This part gathers optimization results performed on the planning software 
Humanoid Path Planner~\cite{hpp}. The initial trajectory is obtained with two 
kind of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
RRT-connect~\cite{rrt-connect}.

\subsection{From 2D basic examples...}

\begin{figure}[t]
	\centering
	\includegraphics[width=7.5cm]{2d_final_green.png}
	%\noindent\rule{8cm}{0.8pt}
	\includegraphics[width=7.5cm]{concave_final_green.png}
	\caption{Path-optimization results on a 2D-ponctual robot, moving around 
	obstacles. Initial paths are in red, optimized ones in black. Grey paths 
	represent intermediate iterations, red dots colliding configurations and blue 
	dots the associated collision-free configurations of the backtracked paths 
	where constraints have been computed.}
	\label{2D_long}
\end{figure}

Figure~\ref{2D_long} shows several iterations of our optimizer on 2D cases. Since 
in this special case, transformation constraints are equivalent to compulsory 
configurations to pass through, we can verify that our optimized path is applying 
all computed constraints. We can also understand which collisions have led to 
the constraints.

Besides, these examples give a better understanding of how the tuning of $\alpha$ 
has to balance lot of iterations and relevant collision-constraint addition. For 
example, if we alter a lot the initial path with a large gradient step and 
compute the corresponding collisions, constraints will be chosen on a very-not 
optimal path and will not be pertinent w.r.t. the obstacles we wanted to 
avoid.

Figure~\ref{2D_long} illustrates a path example that random shortcut will not manage to 
optimize in an affordable time, because of probabilistically failing to sample 
configurations in the box. Our gradient-based algorithm succeeds to optimize the 
path contained in the box, with the following cost coefficients 
$$
\lambda_{i-1} = \frac{1}{\sqrt{(\conf_{i\,0}-\conf_{i-1\,0})^T \weight^2 
(\conf_{i\,0}-\conf_{i-1\,0})}}
$$
aiming at keeping the same ratio between path segment lengths at 
minimum as at 
initial path, represented by the waypoints ($\conf_{i\,0})_{1\leq i\leq n+1}$.
Without these coefficients, the path that minimizes the cost corresponds to a 
straight line with the waypoints equidistantly allocated, which is not adapted for 
the Figure~\ref{local_box_optim} type of problems with a local passage very
constrained by obstacles. Note that this cost is also working with all other 
examples presented in this paper.

\begin{figure}[t]
	\centering
	\includegraphics[width=7.5cm]{local_box_optim.png}
	\caption{Case of a long initial path (above) containing a small part that can 
	be optimized (below). Random shortcut is unlikely to optimize the part containing detours in the box, whereas our method 
	succeeds (in blue).}
	\label{local_box_optim}
\end{figure}

\begin{algorithm}
\begin{algorithmic}[1] % line numerotation
\Procedure{randomShortcut}{$\xx$}
   \State nbFailures $\gets 0$
   \While {$nbFailures < maxNbFailures$}
        \State $failure \gets true$
        \State $T \gets \mbox{upper bound of }\xx\mbox{ definition interval}$
        \State $t_1 < t_2 \gets \mbox{random numbers in }[0,T]$
        \State $lp0 \gets \texttt{steeringMethod} (\xx(0), \xx(t_1))$
        \State $lp1 \gets \texttt{steeringMethod} (\xx(t_1), \xx(t_2))$
        \State $lp2 \gets \texttt{steeringMethod} (\xx(t_2), \xx(T))$
        \State $newPath \gets \mbox{empty path defined on }[0,0]$
   	\If{$lp0$ is collision-free}
          \State $newPath \gets lp0$; $failure \gets false$
        \Else
          \State $newPath \gets \xx_{|[0,t_1]}$
        \EndIf
   	\If{$lp1$ is collision-free}
          \State $newPath \gets \texttt{concatenate} (newPath, lp1)$
          \State $failure \gets false$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_1,t_2]})$
        \EndIf
   	\If{$lp2$ is collision-free}
          \State $newPath \gets \texttt{concatenate} (newPath, lp2)$
          \State $failure \gets false$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_2,T]})$
        \EndIf
        \State $\xx \gets newPath$
        \If {$failure$} $nbFailures \gets nbFailures + 1$
        \EndIf
      \EndWhile
    \Return $\xx$
\EndProcedure
\end{algorithmic}
\caption{Random shortcut as adapted from~\cite{Sekhavat-Svestka1998} 
Section~6.4.1. \texttt{steeringMethod} returns the linear interpolation between 
two configurations. $\xx_{|I}$ denotes path $\xx$ restricted to interval $I$. 
$maxNbFailures$ is a parameter that affects time of computation and quality of 
the result.}\label{algo:random-shortcut}
\end{algorithm}

\subsection{To 3D complex problems}

We also experiment our algorithm on more complex robots, with transformation 
collision-constraints. Unless another value is given, $\alpha_{init}$ is set to 0.2.

%\begin{figure}[b]
%	\centering
%	\includegraphics[width=5cm,height=3.5cm]{ur5_v2.png}
%	\caption{6-DOF robot arm on which our path-optimization has been tested.}
%	\label{ur5}
%\end{figure}

In the included video, we present four situations where our algorithm has been 
% TODO puzzle if included => 5 situations
tested and compared to random shortcut~\cite{randomShortcutHPP}
(Algorithm~\ref{algo:random-shortcut}).
The termination condition of random shortcut allows it to try 
shortening the path until 5 iterations of non-improvement are reached 
(corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
%But we admit that restrincting the number of random shortcut iterations is 
%neither a acceptable criterion because of randomness.
%We also discuss the good properties and limits of our algorithm.

% analysis UR2:
On the 5-DOF double-arm problem, one arm has to get around an obstacle while the 
other is relatively far from obstacles. As expected, the initial path given by 
RRT-connect activates both arms to solve the problem. Contrary to random shortcut, 
our optimizer manages to cancel the rotation of the free-arm while optimizing the 
first arm motion. Initial path length 5.17 has been decreased to 4.55 by random shortcut and to 3.20 by the gradient-based optimizer.

% analysis UR5:
For the following example with a 6-axis manipulator arm 
%(Figure~\ref{ur5})
in a  cluttered environment, path reduction only manages to reduce some but not 
all detours. This behavior can be explained by a path over-constraining due to the 
important proximity of the robot to the obstacles during the whole path, and 
because there are not so many DOF to play on to easily avoid collisions and 
return a small-cost solution.
We advance this explanation because on the two following high-DOF examples of the video (and Figure~\ref{pr2_final}), results are better 
in terms of path length and quality.

% analysis PR2 alone:
In the example presented Figure~\ref{pr2_final} and in the video, the mobile 
40-DOF PR2 simply has to cross its arms from 
the left arm up position to the right arm up one. Lengths of the initial path, the 
random shortcut optimized path and the gradient-based optimized path are 
respectively 14.5, 6.34 and 3.62. The RRT-connect planner 
returns detours and activates non-useful DOF such as the head, the torso lift 
and the translation on 
the ground. Once again, random shortcut will hardly optimize the mobile base 
translation of the robot and other unnecessary DOF uses, unlike our 
optimized-path which mainly results in moving the arms as expected 
(Figure~\ref{pr2_final} right).
%The removal 
%of the translation is also a consequence of our cost ponderation as described in 
%the previous section.
% NON ???

% analysis PR2 in kitchen:
We obtain similar results on the PR2 performing a manipulation task 
in a kitchen environment (see the video). The robot has to move 
his hands from the top to the bottom of a table.
Our optimizer manages to reduce the initial length 15.8 from the visibility-PRM 
planner and improves the path quality 
just adding transformation constraints between the table and the 
robot's arms and grippers. Thus, the robot just slightly moves 
backward and uses its arm DOF to avoid the table, instead of 
processing a large motion to get away from the table. With our method, the path length is downed to 4.59, against 8.24 for random shortcut.

Optimization computation time and path length averages are presented for 30 runs of the PR2-crossing-arms in the Table~\ref{tab:results}. In the same way, time computation averages for the PR2-in-kitchen example have been 
calculated: $59.5\,\text{s}$ (gradient-based optimizer) and $76.2\,\text{s}$ (random shortcut).
Thus our method presents 
similar computation times and path lengths for mobile manipulation tasks. However, the extinction of the mobile base motion in the gradient-based optimized path is advanced in the right column of Table~\ref{tab:results}. Thus the problem addressed 
Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.

\begin{table}[t]
\centering
  \begin{tabular}{cccc}
  \\
  \toprule
     & Computation & Path  & Traveled distance\\
     & time & length & by the base\\
    \midrule
    \midrule
    Initial &  & $13.7$ & $10.9$\\
    Gradient-based & $\,781\text{ms}$ & $3.69$ & $10^{-6}$\\
    Random shortcut & $\,823\text{ms}$ & $3.82$ & $2.94$\\
    \bottomrule
  \end{tabular}
\caption {Results for 30 runs of the PR2-crossing-arms example. For each run, a solution path is planned by Visibility-PRM as initial guess for each optimizer. $\alpha_{init}$ is set to 0.2. The right column represents the distance that is traveled by the robot mobile base during the path.}
\label{tab:results}
\end{table}

\section{CONCLUSIONS}
We managed to settle a path optimization for decoupled-DOF robots such as mobile 
manipulators. Our algorithm uses standard numerical tools as collision checking 
and QP resolution, and correlates them in a 
simple but effective way, playing on the scalar iteration parameter. Therefore, 
our method only require few distances computation, so geometry pre-processing or 
offline optimization are not necessary to remain time competitive. We demonstrate 
that the optimizer is 
time-competitive comparing to random shortcut and proposes better quality paths 
for high-DOF robots, removing unnecessary DOF motions. Our optimizer also manages 
to reduce a local detour in a long path while random shortcut methods will mostly 
fail.

The experimental results show that transformation constraints may in some cases be overconstraining. We also tried distance constraints without much success due to the non-continuous differentiability of distance. In a future work, we will study other one-dimensional constraints in order to achieve better results.

\bibliographystyle{tADR} % plain, alpha, abbr, unsrt
\bibliography{paper}

\end{document}

### 2015-09-30 paper.tex
Fix image and algo-lines references

@@
 	\label{decoupled_DOF_optimization}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[height=4.9cm,width=1.8cm]{pr2_initial_final_vertical.png}
-	\includegraphics[width=5.7cm]{p0_pr2_alone_merged3.png}
-	\includegraphics[height=4.9cm]{p1RS_pr2_alone_merged3.png}
-	\includegraphics[height=4.9cm]{p1GB_pr2_alone_merged3.png}\\
-	\caption{In this motion planning problem, PR2 robot has just to 
-exchange the positions of its arms. The task is simple, however, in absence of 
-explicit indication, any probabilistic motion planner will compute a path that 
-makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
-algorithm (left). Path optimization is expected to remove unnecessary motions. 
-The popular random shortcut algorithm fails in this case (middle). Our algorithm 
-succeeds (right).}
-	\label{pr2-final}
 \end{figure}
@@
 \begin{algorithm}
-\begin{algorithmic}[1] % line numerotation
+\begin{algorithmic}%[1] % line numerotation
 \Procedure{Optimize}{$\xx_0$}
-	\State $\alpha \gets \alpha_{init}$; $\epsilon \gets 10^{-3}$
+	\State $\alpha \gets \alpha_{init}$
 	\State $minReached \gets false$
-	\State $validConstraints \gets true$
 	\While{($\textbf{not}(noCollision \textbf{ and } minReached)$)}
 		\State $\p = \texttt{computeIterate}()$
-		\State $minReached = (||\p||<\epsilon \textbf{ or } \alpha=1)$\label{algo:minReached}
+		\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
 		\State $\xx_1 \gets \xx_0 + \alpha\,\p$
-		\If{($\alpha \neq 1$)}
-			%\State \texttt{solveConstraints($\xx_1$)}
-			\If{(\textbf{not}(\texttt{solveConstraints($\xx_1$)}))} \label{algo:solveCons}
-				\State $\alpha \gets \alpha*0.5$
-				\State $validConstraints \gets false$\label{algo:solveCons2}
-			\Else
-				\State $validConstraints \gets true$
-			\EndIf
-		\EndIf
 		\If{(\textbf{not}(\texttt{validatePath($\xx_1$)}))}
@@
 			\If{$(\alpha \neq 1)$}
-				\State \texttt{addCollisionConstraints}()\label{algo:addCons}
+				\State \texttt{addCollisionConstraints}()
 				\State $\alpha \gets 1$
 			\Else
-				\If{($validConstraints$)}\label{algo:towardMin}
-					\State $\alpha \gets \alpha_{init}$\label{algo:towardMin2}
-				\EndIf
+				\State $\alpha \gets \alpha_{init}$
 			\EndIf
@@
 			\State $\xx_0 \gets \xx_1$
-			\State $noCollision \gets true$\label{algo:noCollision}
-			\State $\alpha \gets 0.5*(1+\alpha)$
+			\State $noCollision \gets true$
 		\EndIf
@@
 Each time a collision-constraint is added, 
-the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution (lines~\ref{algo:minReached}-\ref{algo:noCollision}). Otherwise, 
-smaller steps are iteratively applied and tested toward the minimum (lines~\ref{algo:towardMin}-\ref{algo:towardMin2}), and constraints are added each time a collision is found (line~\ref{algo:addCons}).
+the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
+smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found.
 
-If the constraint relinearization fails to validate Property~\ref{prop:collision free}, $\alpha$ is halved and the algorithm restarts from the latest valid path (lines~\ref{algo:solveCons}-\ref{algo:solveCons2}).
+If the constraint relinearization fails to validate Property~\ref{prop:collision free}, $\alpha$ is halved and the algorithm restarts from the latest valid path.
 
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[height=4.9cm,width=1.8cm]{pr2_initial_final_vertical.png}
+	\includegraphics[width=5.7cm]{p0_pr2_alone_merged3.png}
+	\includegraphics[height=4.9cm]{p1RS_pr2_alone_merged3.png}
+	\includegraphics[height=4.9cm]{p1GB_pr2_alone_merged3.png}\\
+	\caption{In this motion planning problem, PR2 robot has just to 
+exchange the positions of its arms. The task is simple, however, in absence of 
+explicit indication, any probabilistic motion planner will compute a path that 
+makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
+algorithm (left). Path optimization is expected to remove unnecessary motions. 
+The popular random shortcut algorithm fails in this case (middle). Our algorithm 
+succeeds (right).}
+	\label{pr2_final}
+\end{figure}
+
 % analysis PR2 alone:

### 2015-10-07 paper.tex
Remove re-linearization part

@@
 
-\subsection{Relinearization of constraints}
-
-As constraint~(\ref{eq:non-linear constraint}) is not exactly enforced, but linearized, a new collision may appear at $\tcolli$ on a later iteration between
-$\body_1$ and $\body_2$. To avoid this undesired effect, we check the values of all the
-constraints at each iteration. The constraints that fail to satisfy
-Property~\ref{prop:collision free} are relinearized around the latest valid path $\xx_k$, $ k > i+1$, as follows:
-$$
-\frac{\partial F}{\partial \xx}(\xx_k)(\xx -\xx_k) = -F(\xx_k)
-$$
-This step is performed by \texttt{solveConstraints} in Algorithm~\ref{algo:gradient}.
-
 \subsection{Algorithm}
@@
 smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found.
-
-If the constraint relinearization fails to validate Property~\ref{prop:collision free}, $\alpha$ is halved and the algorithm restarts from the latest valid path.
 

### 2015-10-08 paper.tex
Replace transf constraints by orthonogal ones, re-introduce stack of jacobians

@@
 \newcommand\tcolli{t_{coll\ i}}
+\newcommand\po{\mathbf{po}}
 
@@
 are time-optimal since they comply with acceleration and velocity constraints.
+~\cite{Guernane2011}, also based on random shooting, is guiding configuration-generation with local holonimic considerations.
+But this method remains only locally optimal, and is not addressing high-DOF problems.
+~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide which DOF to choose for the partial shortcut.
 
@@
 differentiable path.
-
-TO DESCRIBE:
-
-~\cite{Guernane2011} (guiding shoots) is ``locally optimal'', not addressing high-DOF problems,
-
-~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide the shortcut.
 
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 	\centering
-	\includegraphics[width=8cm]{optim_grad.pdf}
+	\includegraphics[width=9cm]{optim_grad.pdf}
 	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
@@
 \end{enumerate}
-In the rest of this section, we will only consider the first case. Reasoning about the second case is similar, except that the 
-constraint is on the transformation of $\body_1$ with respect to the environment.
+In the rest of this section, we will only consider the first case. Reasoning 
+about the second case is similar, except that the 
+constraint is on the position of $\body_1$ with respect to the environment.
 
-\begin {figure}
-  \centerline {
-    \def\svgwidth {\linewidth}
-    \graphicspath {{./images/}}
-    \input {images/transformation-constraint.pdf_tex}
-  }
-  \caption {Transformation constraint notations.}
-  \label{fig:transformation-constraint}
-\end {figure}
-Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$.
-Let $M\in SE(3)$ denote the same transformation in configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
-Let $R\in SO(3)$ (resp.
-$R^{*}$) and $\translation\in\real^3$ (resp.
-$\translation^{*}$) be the rotation and translation parts of $M$ (resp.
-$M^{*}$).
-See also Figure~\ref{fig:transformation-constraint} for notations.
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=15cm]{contact-points.pdf}
+	\caption{Constraint definition based on contact points.}
+	\label{contact-points}
+\end{figure}
 
-We denote by $F$ the mapping from $\CS^{wp}$ to $\real^6$ defined by
+The principle of the method is to compute a constraint on the backtracked free-
+configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
+(\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
+orthogonal direction of the encountered collision.
+
+In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be the 
+contact point expressed in the global frame ${\cal R}_0$ of each of the two 
+bodies. We compute (see Figure~\ref{contact-points}) the associated points $\po_1$ 
+and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and thus 
+$\mathbf{u}=\po_2-\po_1$.
+
+Let $\translation^{*}\in \real^3$ denotes the position between frames of $\body_1$  
+and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
+\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
+configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
+
+We denote by $F$ the mapping from $\CS^{wp}$ to $\real$ defined by
 \begin{equation}\label{eq:constraint}
-F (\xx) = \left(\begin{array}{c} R^T (\translation^{*} - \translation) \\
-  \log (R^T R^{*})\end{array}\right)
+F (\xx) =  \mathbf{u}^T \, R^T (\translation^{*} - \translation)
 \end{equation}
-$\log (R^T R^{*})$ represents here a vector $\mathbf{v}$ of $\real^3$ such that
-$R^T R^{*}$ is the rotation of axis $\mathbf{v}$ and of angle $\|\mathbf{v}\|$.
 Note that $F$ vanishes for any path $\xx$ such that in configuration
-$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $M^{*}$.
-Let $r$ denotes the \textit{radius} of the bounding sphere of $\body_2$ centered at the origin of $\body_2$ local frame. Let us denote by $d$ the minimal distance between $\body_1$ and $\body_2$
-in configuration $\xx_i(\tcolli)$. We have the following property.
-\begin {property} \label {prop:collision free}
-Let $\xx\in\CS^{wp}$ be a path and let us denote by $\translation$ and $\mathbf{v}$ respectively the 3 first coordinates and the 3 last coordinates of $F(\xx)$. If
-$$
-\|\translation\| + r \|\mathbf{v}\| \leq d
-$$
-then in configuration $\xx(\tcolli)$, $\body_1$ and $\body_2$ are collision-free.
-\end {property}
-The proof is straightforward: expressed in reference frame of $\body_1$, the motion of $\body_2$ between configurations $\xx(\tcolli)$ and $\xx_i(\tcolli)$ is a translation of vector $\translation$ followed by a rotation of angle $\|\mathbf{v}\|$.
-No point of $\body_2$ has moved by more than $\|\translation\| + r \|\mathbf{v}\|$.
-Therefore no collision is possible.
+$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $
+\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is 
+orthogonal to $\mathbf{u}$.
 
@@
 
-We refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
-programs (QP). This corresponds to \texttt{computeIterate} in Algorithm~\ref{algo:gradient}.
+\vspace{0.2cm}
+
+Let $q_{k\,i}$ denotes waypoint $k$ along path $x_i$.
+The constraint on the position at $t_{coll\,i}$ is a function that depends only
+on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
+$\beta\in[0,1]$ such that for any $j$:
+$$
+\xx_j (t_{coll\,i}) = (1-\beta) \conf_{k\,j} + \beta \conf_{k+1\,j}
+$$
+Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
+$x$ is
+$$
+J_F = \left(\begin{array}{cccccc}
+0 & \cdots & (1-\beta)\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) &
+\beta\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
+\end{array}
+\right)
+$$
+Both $F$ and $\frac{\partial F}{\partial q}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.
+
+\vspace{0.2cm}
+
+Each time a new constraint is added to the resolution, we stack its jacobian $J_F$ to the previous ones. This corresponds to \texttt{addCollisionConstraints} in 
+Algorithm~\ref{algo:gradient}. Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic programs (QP) (function \texttt{computeIterate} in 
+Algorithm~\ref{algo:gradient}).
 

### 2015-10-09 paper.tex
Change table results, introduce partial random shortcut

@@
 
-\articletype{GUIDE}
-
 \title{{\itshape Advanced Robotics} A simple path optimization method for motion planning }
@@
 to the randomness of sampling, the paths make detours that need to be optimized.
-The contribution of this paper is to propose a gradient-based algorithm that
+The contribution of this paper is to propose a gradient-based (GB) algorithm that
 transforms a polygonal collision-free path into a shorter one, while both:
@@
 
-For this shortening purpose, random shortcut methods are
-still very popular~\cite{Sekhavat-Svestka1998,HauserFastSmooth}. However, random 
-shortcut requires fine tuning of the termination condition (see 
-Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long trajectories where only a small part needs to be optimized.
-Figure~\ref{decoupled_DOF_optimization} presents another situation where random shortcut will always fail to optimize the initial path, since it cannot decouple the robot degrees of freedom (DOF) on which the optimization occurs, contrary to our algorithm.
+For this shortening purpose, random shortcut (RS) methods are
+still very popular~\cite{Sekhavat-Svestka1998,HauserFastSmooth}. However, RS requires fine tuning of the termination condition (see 
+Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long 
+trajectories where only a small part needs to be optimized. 
+Figure~\ref{decoupled_DOF_optimization} presents another situation where RS will 
+always fail to optimize the initial path, since it cannot decouple the 
+robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
+algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} that shortcut only one or a group of DOF.
 
@@
 are time-optimal since they comply with acceleration and velocity constraints. 
-~\cite{Guernane2011}, also based on random shooting, is guiding configuration-generation with local holonimic considerations.
-But this method remains only locally optimal, and is not addressing high-DOF problems.
-~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is using medial axis retractation for clearance and a partial random shortcut (applied only on random DOF) which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer computation time and is not using the information returned by the checker of which links are colliding to guide which DOF to choose for the partial shortcut.
+~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
+generation with local holonimic considerations. But this method remains only locally 
+optimal, and is not addressing high-DOF problems.
+~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
+using medial axis retractation for clearance and a PRS (applied only on random DOF) 
+which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer 
+computation time and is not using the information returned by the checker of which 
+links are colliding to guide which DOF to choose for the PRS.
 
@@
 where $I_{p_i}$ is the identity matrix of size $p_i$ and $w_i$ is the weight 
-associated to joint $J_i$. We define the length of the straight interpolation 
+associated to the lever arm of the joint $J_i$. We define the length of the straight interpolation 
 between two configurations as:
@@
 bodies. We compute (see Figure~\ref{contact-points}) the associated points $\po_1$ 
-and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and thus 
-$\mathbf{u}=\po_2-\po_1$.
+and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and the difference 
+$\mathbf{u}=\po_2-\po_1$ which will be used as the orthogonal direction the bodies can not move along at the path parameter $\tcolli$.
 
@@
 
-We denote by $F$ the mapping from $\CS^{wp}$ to $\real$ defined by
+We denote by $F$ the mapping from $\CS^{wp}$ to $\real^3$ defined by
 \begin{equation}\label{eq:constraint}
-F (\xx) =  \mathbf{u}^T \, R^T (\translation^{*} - \translation)
+F (\xx) =  R^T (\translation^{*} - \translation)
 \end{equation}
@@
 J_F = \left(\begin{array}{cccccc}
-0 & \cdots & (1-\beta)\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) &
-\beta\frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
+0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
+\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
 \end{array}
@@
 
+%The solution step is given by:
+%$$
+%\mathbf{\pii}^{*} = -V_0(V_0^TH V_0)^{-1}\,V_0^T\nabla c(\xx_i)
+%$$
+%where $V_0$ is the base of the nullspace of the stack of $J_F$ jacobians.
+
 \subsection{Algorithm}
 
-In this part we describe the path-optimizer Algorithm~\ref{algo:gradient} 
+In this part we describe the GB path-optimizer Algorithm~\ref{algo:gradient} 
 according to the previous step definitions. The main difficulty here is to handle 
@@
 \end{algorithmic}
-\caption{Gradient-based path-optimization.} \label{algo:gradient}
+\caption{Gradient-based (GB) path-optimization.} \label{algo:gradient}
 \end{algorithm}
@@
 
-Figure~\ref{2D_long} illustrates a path example that random shortcut will not manage to 
+Figure~\ref{2D_long} illustrates a \textit{very long} path example that RS or PRS will not manage to 
 optimize in an affordable time, because of probabilistically failing to sample 
-configurations in the box. Our gradient-based algorithm succeeds to optimize the 
+configurations in the box. GB succeeds to optimize the 
 path contained in the box, with the following cost coefficients 
@@
 \end{figure}
+
+\subsection{To 3D complex problems}
+
+We also experiment our algorithm on more complex robots, with transformation 
+collision-constraints. Unless another value is given, $\alpha_{init}$ is set to 0.2.
+
+%\begin{figure}[b]
+%	\centering
+%	\includegraphics[width=5cm,height=3.5cm]{ur5_v2.png}
+%	\caption{6-DOF robot arm on which our path-optimization has been tested.}
+%	\label{ur5}
+%\end{figure}
 
@@
 
-\subsection{To 3D complex problems}
+\begin{figure}[b]
+	\centering
+	\includegraphics[height=4.9cm,width=1.8cm]{pr2_initial_final_vertical.png}
+	\includegraphics[width=5.7cm]{p0_pr2_alone_merged3.png}
+	\includegraphics[height=4.9cm]{p1RS_pr2_alone_merged3.png}
+	\includegraphics[height=4.9cm]{p1GB_pr2_alone_merged3.png}\\
+	\caption{PR2-crossing-arms example. In this motion planning problem, PR2 robot 
+	has just to 
+exchange the positions of its arms. The task is simple, however, in absence of 
+explicit indication, any probabilistic motion planner will compute a path that 
+makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
+algorithm (left). Path optimization is expected to remove unnecessary motions. 
+The popular RS algorithm fails in this case (middle). Our GB algorithm 
+succeeds (right).}
+	\label{pr2_final}
+\end{figure}
 
-We also experiment our algorithm on more complex robots, with transformation 
-collision-constraints. Unless another value is given, $\alpha_{init}$ is set to 0.2.
-
-%\begin{figure}[b]
-%	\centering
-%	\includegraphics[width=5cm,height=3.5cm]{ur5_v2.png}
-%	\caption{6-DOF robot arm on which our path-optimization has been tested.}
-%	\label{ur5}
-%\end{figure}
-
-In the included video, we present four situations where our algorithm has been 
-% TODO puzzle if included => 5 situations
-tested and compared to random shortcut~\cite{randomShortcutHPP}
-(Algorithm~\ref{algo:random-shortcut}).
-The termination condition of random shortcut allows it to try 
+In the included video, we present multiple situations where the GB algorithm has 
+been tested and compared to RS~\cite{randomShortcutHPP}
+(Algorithm~\ref{algo:random-shortcut}) and PRS ~\cite{partialrandomShortcutHPP} 
+adapted from~\cite{Geraerts04clearancebased}.
+The termination condition of RS allows it to try 
 shortening the path until 5 iterations of non-improvement are reached 
 (corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
+For PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{n+1}$.
+Then, for each DOF, configurations are shot as in Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by an interpolation on only the current DOF and the previous subpath for other DOF.
+If this path is collision-free, it is replaced as in RS.
+This process is also stopped when 5 iterations of non-improvement are reached.
+
 %But we admit that restrincting the number of random shortcut iterations is 
@@
 % analysis UR2:
-On the 5-DOF double-arm problem, one arm has to get around an obstacle while the 
+On the double-arm problem (see Figure~\ref{trajectories} top-right), one arm has to get around an obstacle while the 
 other is relatively far from obstacles. As expected, the initial path given by 
-RRT-connect activates both arms to solve the problem. Contrary to random shortcut, 
-our optimizer manages to cancel the rotation of the free-arm while optimizing the 
-first arm motion. Initial path length 5.17 has been decreased to 4.55 by random shortcut and to 3.20 by the gradient-based optimizer.
+RRT-connect activates both arms to solve the problem. Contrary to RS, 
+GB manages to cancel the rotation of the free-arm while optimizing the 
+first arm motion. Initial path length 5.17 has been decreased to 4.55 by RS and to 3.20 by GB.
 
 % analysis UR5:
-For the following example with a 6-axis manipulator arm 
+For the following example with a 6-axis manipulator arm (see Figure~\ref{trajectories} bottom-left)
 %(Figure~\ref{ur5})
@@
 return a small-cost solution.
-We advance this explanation because on the two following high-DOF examples of the video (and Figure~\ref{pr2_final}), results are better 
+We advance this explanation because on the two following high-DOF examples of the 
+video (and Figure~\ref{pr2_final}), results are better 
 in terms of path length and quality.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[height=4.9cm,width=1.8cm]{pr2_initial_final_vertical.png}
-	\includegraphics[width=5.7cm]{p0_pr2_alone_merged3.png}
-	\includegraphics[height=4.9cm]{p1RS_pr2_alone_merged3.png}
-	\includegraphics[height=4.9cm]{p1GB_pr2_alone_merged3.png}\\
-	\caption{In this motion planning problem, PR2 robot has just to 
-exchange the positions of its arms. The task is simple, however, in absence of 
-explicit indication, any probabilistic motion planner will compute a path that 
-makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
-algorithm (left). Path optimization is expected to remove unnecessary motions. 
-The popular random shortcut algorithm fails in this case (middle). Our algorithm 
-succeeds (right).}
-	\label{pr2_final}
-\end{figure}
 
@@
 the left arm up position to the right arm up one. Lengths of the initial path, the 
-random shortcut optimized path and the gradient-based optimized path are 
+RS optimized path and the GB optimized path are 
 respectively 14.5, 6.34 and 3.62. The RRT-connect planner 
@@
 and the translation on 
-the ground. Once again, random shortcut will hardly optimize the mobile base 
+the ground. Once again, RS will hardly optimize the mobile base 
 translation of the robot and other unnecessary DOF uses, unlike our 
@@
 backward and uses its arm DOF to avoid the table, instead of 
-processing a large motion to get away from the table. With our method, the path length is downed to 4.59, against 8.24 for random shortcut.
+processing a large motion to get away from the table. With GB, the path length is downed to 4.59, against 8.24 for RS.
 
-Optimization computation time and path length averages are presented for 30 runs of the PR2-crossing-arms in the Table~\ref{tab:results}.
-In the same way, time computation averages for the PR2-in-kitchen example have been
-calculated: $59.5\,\text{s}$ (gradient-based optimizer) and $76.2\,\text{s}$ (random shortcut).
+Optimization computation time and path length averages are presented for 30 runs of 
+the PR2-crossing-arms in the Table~\ref{tab:results}. In the same way, time 
+computation averages for the PR2-in-kitchen example have been 
+calculated: $59.5\,\text{s}$ (GB) and $76.2\,\text{s}$ (RS).
 Thus our method presents 
-similar computation times and path lengths for mobile manipulation tasks.
-However, the extinction of the mobile base motion in the gradient-based optimized path is advanced in the right column of Table~\ref{tab:results}.
-Thus the problem addressed
+similar computation times and path lengths for mobile manipulation tasks. However, 
+the extinction of the mobile base motion in the gradient-based optimized path is 
+advanced in the right column of Table~\ref{tab:results}. Thus the problem addressed 
 Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
 
-\begin{table}[t]
+\begin{figure}[t]
 	\centering
-  \begin{tabular}{cccc}
-  \\
+	\includegraphics[width=7.0cm,height=6.8cm]{pr2_orth_constr_results.png}
+	\includegraphics[width=7.0cm,height=6.8cm]{ur2_orth_constr_results.png}\\
+	\vspace{0.05cm}
+	\includegraphics[width=7.0cm,height=6.3cm]{ur5sph_orth_constr_results.png}
+	\includegraphics[width=7.0cm,height=6.3cm]{puzzle_orth_constr_results.png}
+	\caption{Trajectories of wrists, end-effectors or center allong different paths: the initial path is represented in red, the RS optimized path in blue and the GB optimized path ($\alpha_{init}=0.1$) in green.
+For the freeflyer puzzle (bottom right), points $\po_1$, $\po_2$ and vectors $\mathbf{u}$ for constraints computation are also shown in grey and red.}
+	\label{trajectories}
+\end{figure}
+
+Trajectories results of the robots end-effectors or center are presented 
+Figure\ref{trajectories}.
+
+DESCRIBE TRAJECTORIES
+
+\begin{table}
+\centering
+\scalebox{0.9}{
+  \begin{tabular}{cccccc}
   \toprule
-     & Computation & Path  & Traveled distance\\
-     & time & length & by the base\\
+    \textbf{Problem} & Path obtained from & Computation time  & Path length & Cost & Base traveled distance\\
     \midrule
     \midrule
-    Initial &  & $13.7$ & $10.9$\\
-    Gradient-based & $\,781\text{ms}$ & $3.69$ & $10^{-6}$\\
-    Random shortcut & $\,823\text{ms}$ & $3.82$ & $2.94$\\
+    & Initial &  & $14.3$ & $ $ &\\
+    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $ & $ $ &\\
+    \textbf{puzzle} & PRS & $6.53\,\text{ms}$ & $14.3$ & $ $ &\\
+    & RS & $68.3\,\text{ms}$ & $6.05$ & $ $ &\\
+    \midrule
+    & Initial &  & $11.3$ & $ $ &\\
+    \textbf{Double-arm} & GB & $318\,\text{ms}$ & $4.95$ & $ $ &\\
+    \textbf{5-DOF} & PRS & $8.42\,\text{ms}$ & $8.73$ & $ $ &\\
+    & RS & $61.9\,\text{ms}$ & $6.96$ & $ $ &\\
+    \midrule
+    & Initial &  & $10.8$ & $ $ & $8.03$\,\text{m}\\
+    \textbf{PR2-crossing-} & GB & $24.8\,\text{s}$ & $1.90$ & $ $ & $10^{-7}\,\text{m}$\\
+    \textbf{arms} & PRS & $41.1\,\text{s}$ & $ 4.73$ & $ $ & $0\,\text{m}$\\
+    & RS & $31.5\,\text{s}$ & $1.60$ & $ $ & $2.70\,\text{m}$\\
+    \midrule
+    & Initial &  & $ $ & $ $ &\\
+    \textbf{PR2-in-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
+    \textbf{kitchen} & PRS & $\,\text{s}$ & $ $ & $ $ &\\
+    & RS & $\,\text{s}$ & $ $ & $ $ &\\
+    \midrule
+    & Initial &  & $ $ & $ $ &\\
+    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
+    \textbf{with-spheres} & PRS & $\,\text{s}$ & $ $ & $ $ &\\
+    & RS & $\,\text{s}$ & $ $ & $ $ &\\
+    \midrule
+    & Initial &  & $ $ & $ $ &\\
+    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
+    \textbf{through-window} & PRS & $\,\text{s}$ & $  $ & $ $ &\\
+    & RS &  $\,\text{s}$ & $ $ & $ $ &\\
+    \midrule
+    & Initial &  & $19.0$ & $ $ &\\
+    \textbf{Baxter-in-} & GB & $2449\,\text{s}$ & $5.85$ & $ $ &\\
+    \textbf{office} & PRS & $638\,\text{s}$ & $17.5$ & $ $ &\\
+    & RS & $139\,\text{s}$ & $13.3$ & $ $ &\\
     \bottomrule
   \end{tabular}
-\caption {Results for 30 runs of the PR2-crossing-arms example.
-For each run, a solution path is planned by Visibility-PRM as initial guess for each optimizer.
-$\alpha_{init}$ is set to 0.2.
-The right column represents the distance that is traveled by the robot mobile base during the path.}
+  }%scalebox
+\caption{Average results for 30 runs of several examples presented in the paper or 
+the video. For each run, a solution path is planned by Visibility-PRM as initial 
+guess for each optimizer. $\alpha_{init}$ is set to 0.1.
+For the PR2-crossing-arms example only, the right column represents the 
+distance that is traveled by the robot mobile base during the different paths.
+}
 \label{tab:results}
@@
 
-The experimental results show that transformation constraints may in some cases be overconstraining.
-We also tried distance constraints without much success due to the non-continuous differentiability of distance.
-In a future work, we will study other one-dimensional constraints in order to achieve better results.
+FUTURE WORK
 

### 2015-10-13 paper.tex
Add FIAD figures, correct mistakes

@@
 
-\jvol{00} \jnum{00} \jyear{2013} \jmonth{January}
+\jvol{00} \jnum{00} \jyear{2015} \jmonth{October}
 
@@
 For this shortening purpose, random shortcut (RS) methods are
-still very popular~\cite{Sekhavat-Svestka1998,HauserFastSmooth}. However, RS requires fine tuning of the termination condition (see 
+still very 
+popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
+However, RS requires fine 
+tuning of the termination condition (see 
 Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long 
@@
 robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
-algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} that shortcut only one or a group of DOF.
+algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} 
+that shortcut only one or a group of DOF.
 
@@
 When a collision is detected at a given iteration, the method backtracks to the
-latest valid iteration and inserts a linearized transformation constraint
-between the objects detected in collision. Only few distances between objects are evaluated, therefore no pre-processing of the
-robot or environment models is necessary. The underlying optimization
-algorithm is a quadratic program.
+latest valid iteration and inserts a one-dimensional constraint
+between the objects detected in collision. Only collisions between objects are 
+evaluated, therefore no pre-processing of the
+robot or environment models is necessary to increase distance computation speed. 
+Preserving the problem geometry also 
+guarantees that a solution can still be found, e.g. for narrow passages as holes or 
+grippers. The underlying optimization algorithm is a quadratic program.
 
@@
 using medial axis retractation for clearance and a PRS (applied only on random DOF) 
-which can address the problem of ~\ref{decoupled_DOF_optimization} but has longer 
-computation time and is not using the information returned by the checker of which 
-links are colliding to guide which DOF to choose for the PRS.
+which can address the problem of Figure~\ref{decoupled_DOF_optimization} but is 
+relatively slower than RS. Moreover, PRS is not using the information returned by 
+the checker of which links are colliding to guide which DOF or group of DOF are 
+relevant to shortcut, contrary to our method.
 
@@
 iteration is detected in collision, instead of trying to constantly satisfy 
-distance constraints. Unlike our method however, the iterations are composed of 
-Cubic-B-splines. The benefit of this method is that the result is a 
-differentiable path.
+distance constraints. Unlike our method however, interpolating configuration -which 
+can be viewed as constraints- are freezing the whole robot configuration instead of 
+a relevant subpart. Thus, this method cannot solve the 
+Figure~\ref{decoupled_DOF_optimization} problem.
+
+%Unlike our method however, the iterations are composed of 
+%Cubic-B-splines. The benefit of this method is that the result is a 
+%differentiable path.
 
@@
 	\includegraphics[width=15cm]{contact-points.pdf}
-	\caption{Constraint definition based on contact points.}
+	\caption{Bodies representation in collision-configuration (left) and 
+	backtracked free-configuration (right). A contact point (i.e. any point in 
+	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
+	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
+	line is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
+	direction which is allowed by the constraint.}
 	\label{contact-points}
@@
 
-In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be the 
+In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
 contact point expressed in the global frame ${\cal R}_0$ of each of the two 
-bodies. We compute (see Figure~\ref{contact-points}) the associated points $\po_1$ 
+bodies (Figure~\ref{contact-points} left). We compute the associated points $\po_1$ 
 and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and the difference 
-$\mathbf{u}=\po_2-\po_1$ which will be used as the orthogonal direction the bodies can not move along at the path parameter $\tcolli$.
+$\mathbf{u}=\po_2-\po_1$ which will be used as the orthogonal direction the 
+bodies can not move along at the path parameter $\tcolli$ 
+(Figure~\ref{contact-points} right).
 
-Let $\translation^{*}\in \real^3$ denotes the position between frames of $\body_1$  
-and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
-\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
-configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
+%Let $\translation^{*}\in \real^3$ denotes the position between frames of $
+%\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
+%\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
+%configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
 
-We denote by $F$ the mapping from $\CS^{wp}$ to $\real^3$ defined by
+Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
+\body_1$  and $\body_2$ in configuration 
+$\x_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
+$\x (\tcolli)$ where $\x$ is any path 
+in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
+(resp. $\translation^{*}$) be the rotation 
+and translation parts of $M$ (resp. $M^{*}$).
+
+We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
+\real^3$ defined by
 \begin{equation}\label{eq:constraint}
-F (\xx) =  R^T (\translation^{*} - \translation)
+F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
 \end{equation}
+where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.
+
 Note that $F$ vanishes for any path $\xx$ such that in configuration
 $\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $
-\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is 
-orthogonal to $\mathbf{u}$.
+\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is orthogonal to $\mathbf{u}$.
 
@@
 $$
-for any later iteration $\xx_j$, $j\geq i+1$.
+for any later iteration $\xx_j$, $j\geq i+1$. Both $F$ and $\frac{\partial F}{\partial \conf}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.
 
@@
 
-Let $q_{k\,i}$ denotes waypoint $k$ along path $x_i$.
+Let $q_{k\,i}$ denote the waypoint $k$ along path $x_i$.
 The constraint on the position at $t_{coll\,i}$ is a function that depends only
@@
 $$
-\xx_j (t_{coll\,i}) = (1-\beta) \conf_{k\,j} + \beta \conf_{k+1\,j}
+\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
 $$
-Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
-$x$ is
-$$
-J_F = \left(\begin{array}{cccccc}
-0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
-\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
-\end{array}
-\right)
-$$
-Both $F$ and $\frac{\partial F}{\partial q}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.
+The path Jacobian $J_F$ of constraint~(\ref{eq:constraint}) w.r.t. $x$ is thus built by block using the previous constraint jacobian $\frac{\partial F}{\partial \conf}$ and $\beta$. ????
+%Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
+%$x$ is
+%$$
+%J_F = \left(\begin{array}{cccccc}
+%0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
+%\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
+%\end{array}
+%\right)
+%$$
 
@@
 	\centering
-	\includegraphics[width=7.5cm]{2d_final_green.png}
+	\includegraphics[width=7.8cm]{contact_points6.png}
 	%\noindent\rule{8cm}{0.8pt}
-	\includegraphics[width=7.5cm]{concave_final_green.png}
-	\caption{Path-optimization results on a 2D-ponctual robot, moving around 
-	obstacles. Initial paths are in red, optimized ones in black. Grey paths 
-	represent intermediate iterations, red dots colliding configurations and blue 
-	dots the associated collision-free configurations of the backtracked paths 
+	\includegraphics[width=8cm]{contact_points2potential.png}
+	\caption{Path-optimization results on 2D-ponctual robots, moving around 
+	green obstacles. Initial paths are in red, optimized ones in black. Grey paths 
+	represent intermediate iterations, red and black dots resp. $\po_1$ and $\po_2$ 
+	points, red segments $\mathbf{u}$ and blue circles collision-free configurations 
+	of the backtracked paths 
 	where constraints have been computed.}
@@
 Figure~\ref{2D_long} shows several iterations of our optimizer on 2D cases. Since 
-in this special case, transformation constraints are equivalent to compulsory 
-configurations to pass through, we can verify that our optimized path is applying 
+in this special case, orthogonal constraints are equivalent to forbidden motion 
+direction, we can verify that our optimized path is applying 
 all computed constraints. We can also understand which collisions have led to 
@@
 
-Figure~\ref{2D_long} illustrates a \textit{very long} path example that RS or PRS will not manage to 
+Figure~\ref{local_box_optim} illustrates a \textit{very long} path example that RS or PRS will not manage to 
 optimize in an affordable time, because of probabilistically failing to sample 
@@
 	\caption{Case of a long initial path (above) containing a small part that can 
-	be optimized (below). Random shortcut is unlikely to optimize the part containing detours in the box, whereas our method 
-	succeeds (in blue).}
+	be optimized (below). Random shortcut is unlikely to optimize the part 
+	containing detours in the box, whereas our method 
+	succeeds (in blue). This type of issue is common in navigation problems in 
+	environments with long corridors.}
 	\label{local_box_optim}
@@
 explicit indication, any probabilistic motion planner will compute a path that 
-makes the PR2 mobile basis purposelessly moving. This is the case of RRT-connect 
+makes the PR2 mobile base purposelessly moving. This is the case of RRT-connect 
 algorithm (left). Path optimization is expected to remove unnecessary motions. 
@@
 first arm motion. Initial path length 5.17 has been decreased to 4.55 by RS and to 3.20 by GB.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=6cm]{fiad_qinit.png}
+	\includegraphics[width=6cm]{fiad_qfinal.png}\\
+	\includegraphics[width=6cm]{fiad_real.png}
+	\includegraphics[width=6cm]{fiad-trajectorires.png}
+
+	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
+	the 
+	Factory-in-a-Day project~\cite{factory-day-video}. A similar environment have 
+	been created (top) to 
+	illustrate that our method can comply with an industrial problem, where 
+	initial and final configurations are constrained in boxes. Trajectories results 
+	are presented (bottom right).}
+	\label{fiad}
+\end{figure}
 
@@
 similar computation times and path lengths for mobile manipulation tasks. However, 
-the extinction of the mobile base motion in the gradient-based optimized path is 
-advanced in the right column of Table~\ref{tab:results}. Thus the problem addressed 
+the disappearance of the mobile base motion in the gradient-based optimized path is 
+shown in the right column of Table~\ref{tab:results}. Thus the problem addressed 
 Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
@@
     \midrule
-    & Initial &  & $ $ & $ $ &\\
-    \textbf{PR2-in-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
-    \textbf{kitchen} & PRS & $\,\text{s}$ & $ $ & $ $ &\\
-    & RS & $\,\text{s}$ & $ $ & $ $ &\\
+    & Initial &  & $15.0$ & $ $ &\\
+    \textbf{PR2-in-} & GB & $288\,\text{s}$ & $4.40$ & $ $ &\\
+    \textbf{kitchen} & PRS & $431\,\text{s}$ & $12.1$ & $ $ &\\
+    & RS & $95.6\,\text{s}$ & $5.33$ & $ $ &\\
     \midrule

### 2015-10-16 paper.tex
Remove cost column, fix typo

@@
 algorithm. Finally, we conclude on experimental results in Section IV.
+
+????? This paper is based on a similar work, which has been sent to ICRA, with a different type of collision constraints.
 
@@
 
-We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. 
+We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. This 
+corresponds to the unconstrained function \texttt{computeIterate} in 
+Algorithm~\ref{algo:gradient}.
 When a collision is detected, we introduce a constraint and perform a new 
@@
 \body_1$  and $\body_2$ in configuration 
-$\x_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
-$\x (\tcolli)$ where $\x$ is any path 
+$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
+$\xx (\tcolli)$ where $\xx$ is any path 
 in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
@@
 	\includegraphics[width=7.0cm,height=6.3cm]{puzzle_orth_constr_results.png}
-	\caption{Trajectories of wrists, end-effectors or center allong different paths: the initial path is represented in red, the RS optimized path in blue and the GB optimized path ($\alpha_{init}=0.1$) in green.
-For the freeflyer puzzle (bottom right), points $\po_1$, $\po_2$ and vectors $\mathbf{u}$ for constraints computation are also shown in grey and red.}
+	\caption{Trajectories of wrists, end-effectors or center allong different paths: the initial path is represented in red, the RS optimized path in blue and the GB optimized path ($\alpha_{init}=0.1$) in green.}
 	\label{trajectories}
@@
 \scalebox{0.9}{
-  \begin{tabular}{cccccc}
+  \begin{tabular}{ccccc}
   \toprule
-    \textbf{Problem} & Path obtained from & Computation time  & Path length & Cost & Base traveled distance\\
+    \textbf{Problem} & Path obtained from & Computation time  & Path length & Base traveled distance\\
     \midrule
     \midrule
-    & Initial &  & $14.3$ & $ $ &\\
-    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $ & $ $ &\\
-    \textbf{puzzle} & PRS & $6.53\,\text{ms}$ & $14.3$ & $ $ &\\
-    & RS & $68.3\,\text{ms}$ & $6.05$ & $ $ &\\
+    & Initial &  & $14.3$  &\\
+    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $  &\\
+    \textbf{puzzle} & PRS & $6.53\,\text{ms}$ & $14.3$  &\\
+    & RS & $68.3\,\text{ms}$ & $6.05$  &\\
     \midrule
-    & Initial &  & $11.3$ & $ $ &\\
-    \textbf{Double-arm} & GB & $318\,\text{ms}$ & $4.95$ & $ $ &\\
-    \textbf{5-DOF} & PRS & $8.42\,\text{ms}$ & $8.73$ & $ $ &\\
-    & RS & $61.9\,\text{ms}$ & $6.96$ & $ $ &\\
+    & Initial &  & $11.3$  &\\
+    \textbf{Double-arm} & GB & $318\,\text{ms}$ & $4.95$  &\\
+    \textbf{5-DOF} & PRS & $8.42\,\text{ms}$ & $8.73$  &\\
+    & RS & $61.9\,\text{ms}$ & $6.96$  &\\
     \midrule
-    & Initial &  & $10.8$ & $ $ & $8.03$\,\text{m}\\
-    \textbf{PR2-crossing-} & GB & $24.8\,\text{s}$ & $1.90$ & $ $ & $10^{-7}\,\text{m}$\\
-    \textbf{arms} & PRS & $41.1\,\text{s}$ & $ 4.73$ & $ $ & $0\,\text{m}$\\
-    & RS & $31.5\,\text{s}$ & $1.60$ & $ $ & $2.70\,\text{m}$\\
+    & Initial &  & $10.8$  & $8.03$\,\text{m}\\
+    \textbf{PR2-crossing-} & GB & $24.8\,\text{s}$ & $1.90$  & $0\,\text{m}$\\
+    \textbf{arms} & PRS & $41.1\,\text{s}$ & $ 4.73$ & $0\,\text{m}$\\
+    & RS & $31.5\,\text{s}$ & $1.60$ & $2.70\,\text{m}$\\
     \midrule
-    & Initial &  & $15.0$ & $ $ &\\
-    \textbf{PR2-in-} & GB & $288\,\text{s}$ & $4.40$ & $ $ &\\
-    \textbf{kitchen} & PRS & $431\,\text{s}$ & $12.1$ & $ $ &\\
-    & RS & $95.6\,\text{s}$ & $5.33$ & $ $ &\\
+    & Initial &  & $15.0$ &\\
+    \textbf{PR2-in-} & GB & $288\,\text{s}$ & $4.40$ &\\
+    \textbf{kitchen} & PRS & $431\,\text{s}$ & $12.1$ &\\
+    & RS & $95.6\,\text{s}$ & $5.33$ &\\
     \midrule
-    & Initial &  & $ $ & $ $ &\\
-    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
-    \textbf{with-spheres} & PRS & $\,\text{s}$ & $ $ & $ $ &\\
-    & RS & $\,\text{s}$ & $ $ & $ $ &\\
+    & Initial &  & $ $ &\\
+    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ &\\
+    \textbf{with-spheres} & PRS & $\,\text{s}$ & $ $ &\\
+    & RS & $\,\text{s}$ & $ $ &\\
     \midrule
-    & Initial &  & $ $ & $ $ &\\
-    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ & $ $ &\\
-    \textbf{through-window} & PRS & $\,\text{s}$ & $  $ & $ $ &\\
-    & RS &  $\,\text{s}$ & $ $ & $ $ &\\
+    & Initial &  & $ $ &\\
+    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ &\\
+    \textbf{through-window} & PRS & $\,\text{s}$ & $  $ &\\
+    & RS &  $\,\text{s}$ & $ $ &\\
     \midrule
-    & Initial &  & $19.0$ & $ $ &\\
-    \textbf{Baxter-in-} & GB & $2449\,\text{s}$ & $5.85$ & $ $ &\\
-    \textbf{office} & PRS & $638\,\text{s}$ & $17.5$ & $ $ &\\
-    & RS & $139\,\text{s}$ & $13.3$ & $ $ &\\
+    & Initial &  & $19.0$ &\\
+    \textbf{Baxter-in-} & GB & $2449\,\text{s}$ & $5.85$ &\\
+    \textbf{office} & PRS & $638\,\text{s}$ & $17.5$ &\\
+    & RS & $139\,\text{s}$ & $13.3$ &\\
     \bottomrule

### 2015-10-20 paper.tex
Add results description, change images, re-fill table

@@
 \usepackage{caption}
+\usepackage{wrapfig}
 \usepackage{color}
@@
 \usepackage[numbers]{natbib}
-
+\algnewcommand\algorithmicinput{\textbf{input:}}
+\algnewcommand\INPUT{\item[\algorithmicinput]}
+\algnewcommand\algorithmicoutput{\textbf{output:}}
+\algnewcommand\OUTPUT{\item[\algorithmicoutput]}
 \makeatletter
@@
 \newcommand\conf{\mathbf{q}}
-\newcommand\xx{\mathbf{x}}
+\newcommand\xx{\mathbf{x}} % \x already defined in the class
 \newcommand\cost{C}
@@
 \newcommand\po{\mathbf{po}}
+\newcommand\Jf{{J_f}}
+\newcommand\kernel{\mbox{Ker }}
 
@@
 \begin{itemize}
-\item requiring mainly collision checking, and few time-consuming obstacle 
-distance computation,
+\item requiring only collision checking, and not any time-consuming obstacle distance computation,
 \item constraining only part of the configuration variables that may cause a 
@@
 robot or environment models is necessary to increase distance computation speed. 
-Preserving the problem geometry also 
-guarantees that a solution can still be found, e.g. for narrow passages as holes or 
-grippers. The underlying optimization algorithm is a quadratic program.
+Respecting the problem geometry also 
+preserves that a solution can still be found, e.g. for narrow passages as holes or 
+grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a quadratic program.
 
@@
 
-????? This paper is based on a similar work, which has been sent to ICRA, with a different type of collision constraints.
+This paper is based on a similar work sent to a conference, where a different type of collision constraints was studied.
 
@@
 
+Let us assume that at iteration $i$, $k$ constraints have been inserted before the current interation. These constraints are stored as lines of a Jacobian matrix
+$$
+\Jf_{i} = \left(\begin{array}{c}L_1 \\ \vdots \\ L_k\end{array}\right)
+$$
+where iterate $\p_i$ are constrained to be in the kernel of $\Jf_{i}$
+$$
+\Jf_{i} \p_i = 0
+$$
+
+\subsection*{New constraint}
+
 Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 	\centering
-	\includegraphics[width=15cm]{contact-points.pdf}
+	\includegraphics[width=15.8cm]{contact-points.pdf}
 	\caption{Bodies representation in collision-configuration (left) and 
@@
 	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
-	line is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
+	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
 	direction which is allowed by the constraint.}
@@
 In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
-contact point expressed in the global frame ${\cal R}_0$ of each of the two 
-bodies (Figure~\ref{contact-points} left). We compute the associated points $\po_1$ 
-and $\po_2$ in the free-configuration $\xx_{i}(\tcolli)$, and the difference 
-$\mathbf{u}=\po_2-\po_1$ which will be used as the orthogonal direction the 
-bodies can not move along at the path parameter $\tcolli$ 
-(Figure~\ref{contact-points} right).
+contact point expressed in the global frame ${\cal R}_0$ 
+(Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
+$\po_2$ of $\po_c$ in 
+the local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
 
-%Let $\translation^{*}\in \real^3$ denotes the position between frames of $
-%\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
-%\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
-%configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
+Let $M_1 \in SE(3)$ (resp. $M_2 \in SE(3)$) denote the homogeneous transformation 
+between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
+$\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
+linking the contact points in configuration 
+$\xx_{i}(t_{coll\ i})$ (the latest collision-free path), expressed in the local 
+frame of $\body_1$:
+$$
+\mathbf{u} = \frac{M_{1}^{-1} (\xx_i) M_2 (\xx_i) \po_2 - \po_1}{\|M_{1}^{-1} 
+(\xx_i) M_2 (\xx_i) \po_2 - \po_1\|}
+$$
 
-Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
-\body_1$  and $\body_2$ in configuration 
-$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
-$\xx (\tcolli)$ where $\xx$ is any path 
-in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
-(resp. $\translation^{*}$) be the rotation 
-and translation parts of $M$ (resp. $M^{*}$).
-
-We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
-\real^3$ defined by
-\begin{equation}\label{eq:constraint}
-F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
-\end{equation}
-where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.
-
-Note that $F$ vanishes for any path $\xx$ such that in configuration
-$\xx(\tcolli)$ the relative position of $\body_2$ w.r.t. $\body_1$ is equal to $
-\translation^{*}$ or if the relative position $\translation^{*} - \translation$ is orthogonal to $\mathbf{u}$.
-
-When a collision is detected at $\tcolli$ along path $\xx_{i+1}$, we wish to add
-constraint
-\begin{equation}\label{eq:non-linear constraint}
-F (\xx) = 0
-\end{equation}
-However, as this constraint is non-linear, we linearize it around $\xx_i$:
+Notice that $\mathbf{u}$ is well defined since configuration 
+$\xx_{i}(t_{coll\ i})$ is collision-free.
+Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
 $$
-\frac{\partial F}{\partial \xx}(\xx_i)(\xx -\xx_i) = 0
+f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
 $$
-for any later iteration $\xx_j$, $j\geq i+1$. Both $F$ and $\frac{\partial F}{\partial \conf}$ are provided by our path planning framework~\cite{hpp} as differentiable functions.
-
-\vspace{0.2cm}
+The constraint is defined by
+\begin {equation}\label{eq:new-constraint}
+f(\xx) - f(\xx_{i}) = 0.
+\end {equation}
+As this constraint is non-linear, it is linearized around $\xx_{i}$.
 
@@
 $$
-The path Jacobian $J_F$ of constraint~(\ref{eq:constraint}) w.r.t. $x$ is thus built by block using the previous constraint jacobian $\frac{\partial F}{\partial \conf}$ and $\beta$. ????
+The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
+blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
+and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
+Algorithm~\ref{algo:gradient}.
+Then, a line is added in the contraint Jacobian matrix (function 
+\texttt{addCollisionConstraint}), and the constraint 
+Jacobian becomes:
+\begin {align*}
+\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
+L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
+\end{align*}
+
+%Let $\translation^{*}\in \real^3$ denotes the position between frames of $
+%\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
+%\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
+%configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
+
+%Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
+%\body_1$  and $\body_2$ in configuration 
+%$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
+%$\xx (\tcolli)$ where $\xx$ is any path 
+%in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
+%(resp. $\translation^{*}$) be the rotation 
+%and translation parts of $M$ (resp. $M^{*}$).
+
+%We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
+%\real^3$ defined by
+%\begin{equation}\label{eq:constraint}
+%F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
+%\end{equation}
+%where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.
+
+\vspace{0.2cm}
+
 %Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
@@
 
-Each time a new constraint is added to the resolution, we stack its jacobian $J_F$ to the previous ones. This corresponds to \texttt{addCollisionConstraints} in 
-Algorithm~\ref{algo:gradient}. Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic programs (QP) (function \texttt{computeIterate} in 
+Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
+programs (QP) (function \texttt{computeIterate} in 
 Algorithm~\ref{algo:gradient}).
@@
 
+\subsection*{Convergence Analysis}
+For iteratively constrained problem, the risk of falling in an infinite loop 
+that keeps adding a redundant constraint exists.
+In this section, we demonstrate that, given the constraint definition and the 
+function \texttt{solveRedundantConstraint}, our 
+algorithm will converge.
+Under the previous notations, it is straight-forward that
+\begin{equation}\label{eq:decreasing-kernel}
+\kernel \Jf_{i+1} \subset \kernel \Jf_{i}
+\end{equation}
+Let us assume that
+\begin{equation}\label{eq:asumption-1}
+L_{k+1}\p_i \not= 0.
+\end{equation}
+We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then
+$$
+\kernel J_i \not= \kernel J_{i+1}.
+$$
+From~(\ref{eq:decreasing-kernel}), we deduce that
+$$
+\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
+$$
+This results proves that each added constraint is linearly independant from the 
+previous ones, and thus our algorithm terminates in a finite number of iterations.
+
+\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
+Let us consider the following expression:
+$$
+\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
+$$
+This expression represents the velocity of point $\po_2$ in the reference frame of Body~1 projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
+\begin{equation}\label{eq:trajectory}
+\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
+\end{equation}
+The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
+path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
+reached for $t=1$.
+This expression being equal to 0 requires either
+\begin {enumerate}
+\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
+\item the velocity of $\po_2$ in the reference frame of Body~1 is orthogonal to 
+$\mathbf{u}$ at contact ($t=1$).
+\end {enumerate}
+The first case is very unlikely and almost never encountered in real applications, 
+except if the user builds such a case on purpose.
+The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
+case, we do not directly insert the new constraint. Instead, we process intermediate 
+iterations until the redundancy vanishes, as explained in function 
+\texttt{solveRedundantConstraint}.
+
+\begin{algorithm}
+\begin{algorithmic}%[1] % line numerotation
+\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
+\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
+\State $solved \gets false$
+\State $minReached \gets false$
+\While{($\textbf{not}(solved)$)}
+	\State $\p = \texttt{computeIterate}()$
+	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
+	\State $\xx \gets \xx_{Free} + \frac{\alpha}{2}\,\p$
+	\If{(\texttt{validatePath}($\xx_1$))}
+		\State $\xx_{Free} \gets \xx$
+	\Else
+		\State $\xx_{Coll} \gets \xx$
+	\EndIf
+	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
+	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
+\EndWhile
+\end{algorithmic}
+\caption{\texttt{solveRedundantConstraint}()} \label{algo:solveRedundant}
+\end{algorithm}
+
+After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ was 
+halved, the obtained path is tested for collisions. If it is collision-free, the 
+constraint is now linearized around it. Otherwise, new contact points are taken 
+on it, and the constraint is linearized around the latest free path.
+
+Doing so, either the new constraint makes the dimension of the kernel decrease, or 
+makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
+case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
+closer and closer. And the continuity of $f$ implies that
+$$
+\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
+$$
+converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
+body~1 along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
+contact. We already handled this case previously, admitting that the velocity norm 
+would not vanish. Therefore, is~\ref{eq:asumption-1} verified.
+
 \subsection{Algorithm}
@@
 be traveled through, and also to return a collision-free path. Typically, we 
-chose $\alpha_{init} = 0.2$ to process small steps. $\alpha_{init} = 0.5$ is more 
-encountered in the optimization literature, procuring larger steps but with more 
-risks of collision.
+chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
+more encountered in the optimization literature, procuring larger steps but with 
+more risks of collision.
 
@@
 \begin{algorithmic}%[1] % line numerotation
-\Procedure{Optimize}{$\xx_0$}
+\INPUT{Path to optimize $\xx_0$}
+\OUTPUT{Optimized collision-free path $\xx_0$}
 \State $\alpha \gets \alpha_{init}$
@@
 		\If{$(\alpha \neq 1)$}
-				\State \texttt{addCollisionConstraints}()
+			\State \texttt{computeCollisionConstraint}($\xx_1$, $\xx_0$)
+			\State \texttt{solveRedundantConstraint}()
+			\State \texttt{addCollisionConstraint}()
 			\State $\alpha \gets 1$
@@
 \EndWhile
-	\Return $\xx_0$
-\EndProcedure
 \end{algorithmic}
-\caption{Gradient-based (GB) path-optimization.} \label{algo:gradient}
+\caption{Gradient-based (GB) algorithm for path-optimization.} \label{algo:gradient}
 \end{algorithm}
@@
 the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
-smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found.
+smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found with a small step (i.e. $\alpha \neq 1$).
 
@@
 	\includegraphics[width=7.8cm]{contact_points6.png}
-	%\noindent\rule{8cm}{0.8pt}
 	\includegraphics[width=8cm]{contact_points2potential.png}
@@
 	represent intermediate iterations, red and black dots resp. $\po_1$ and $\po_2$ 
-	points, red segments $\mathbf{u}$ and blue circles collision-free configurations 
+	points, red segments $\mathbf{u}$ and blue circles collision-free 
+	configurations 
 	of the backtracked paths 
@@
 
-Figure~\ref{local_box_optim} illustrates a \textit{very long} path example that RS or PRS will not manage to 
+Figure~\ref{local_box_optim} illustrates a \textit{very long} path example that RS 
+or PRS will not manage to 
 optimize in an affordable time, because of probabilistically failing to sample 
@@
 
-In the included video, we present multiple situations where the GB algorithm has 
+In the included video and Figures~\ref{pr2_final},~\ref{fiad} and
+ ~\ref{fig:trajectories}, we present multiple situations where the GB algorithm has 
 been tested and compared to RS~\cite{randomShortcutHPP}
@@
 (corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
-For PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{n+1}$.
-Then, for each DOF, configurations are shot as in Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by an interpolation on only the current DOF and the previous subpath for other DOF.
-If this path is collision-free, it is replaced as in RS.
-This process is also stopped when 5 iterations of non-improvement are reached.
+For our implementation of PRS, first a shortcut is tried on each DOF between 
+$\conf_0$ and $\conf_{n+1}$. 
+Then, for each DOF, configurations are shot as in 
+Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by 
+an interpolation on only 
+the current DOF and the previous subpath for other DOF. If this path is collision-
+free, it is replaced as in RS. This process is also stopped when 5 iterations of 
+non-improvement are reached.
 
@@
 
-% analysis UR2:
-On the double-arm problem (see Figure~\ref{trajectories} top-right), one arm has to get around an obstacle while the 
-other is relatively far from obstacles. As expected, the initial path given by 
+\vspace{0.4cm}
+% freeflyer puzzle
+Before entering the manipulators examples, the GB algorithm is tested and compared 
+on a freeflyer puzzle example (see Figure~\ref{fig:trajectories} top right), very 
+encountered in the motion planning literature. The puzzle has to cross down the 
+obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
+contains detours above and below the obstacle, as well as small motion in the hole. 
+Lengths of initial, GB optimized, and RS optimized paths are 
+respectively 23.4, 8.53 and 10.2.
+PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole.
+This behaviour was also noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, acting on a group of DOF.
+The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.
+
+\vspace{0.4cm}
+
+% double-arm:
+On the double-arm problem (see Figure~\ref{fig:trajectories} top-left), one arm has 
+to get around a cylinder obstacle while the 
+other arm is relatively far from obstacles. As expected, the initial path given by 
 RRT-connect activates both arms to solve the problem. Contrary to RS, 
 GB manages to cancel the rotation of the free-arm while optimizing the 
-first arm motion. Initial path length 5.17 has been decreased to 4.55 by RS and to 3.20 by GB.
+first arm motion, even if the second arm is still moving because of the translation 
+motion. The initial path length 5.17 is decreased to 4.55 by RS, to 4.06 by PRS and 
+to 2.77 by GB using 3 collision-constraints with the cylinder obstacle.
 
@@
 	\centering
-	\includegraphics[width=6cm]{fiad_qinit.png}
-	\includegraphics[width=6cm]{fiad_qfinal.png}\\
-	\includegraphics[width=6cm]{fiad_real.png}
-	\includegraphics[width=6cm]{fiad-trajectorires.png}
-
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qinit.png}
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qfinal.png}\\
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_real.png}
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad-trajectorires.png}
 	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
-	the 
-	Factory-in-a-Day project~\cite{factory-day-video}. A similar environment have 
-	been created (top) to 
-	illustrate that our method can comply with an industrial problem, where 
-	initial and final configurations are constrained in boxes. Trajectories results 
-	are presented (bottom right).}
+	the Factory-in-a-Day project~\cite{factory-day-video}. A similar environment 	
+	has been created (top) to illustrate that our method can comply with an 
+	industrial problem, where initial and final configurations of the UR5 are 
+	constrained in boxes. Trajectories results 
+	are presented (bottom right): initial (RRT) path in red, RS optimization in 
+	blue and GB optimization in green.}
 	\label{fiad}
@@
 
-% analysis UR5:
-For the following example with a 6-axis manipulator arm (see Figure~\ref{trajectories} bottom-left)
-%(Figure~\ref{ur5})
-in a  cluttered environment, path reduction only manages to reduce some but not 
-all detours. This behavior can be explained by a path over-constraining due to the 
-important proximity of the robot to the obstacles during the whole path, and 
-because there are not so many DOF to play on to easily avoid collisions and 
-return a small-cost solution.
-We advance this explanation because on the two following high-DOF examples of the 
-video (and Figure~\ref{pr2_final}), results are better 
-in terms of path length and quality.
+% UR5-spheres:
+Some problems were 
+involving an a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
+gripper.
+In a relatively free environment (Figure~\ref{fig:trajectories} middle right), 
+results from GB and RS are similar (path length are respectively 2.03 and 1.60). 
+Note also that the end-effector trajectory has completely changed from the initial 
+one, the robot is easily passing between the spheres, keeping its end-effector 
+above.
+
+% UR5-fiad
+For an UR5 working in a cluttered environment inspired from an industrial issue 
+Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
+to the result of RS and to what can be observed in reality~\cite{factory-day-video}. 
+The initial path length 6.65 is decreased to 2.08 by RS, to 
+1.98 by GB and unchanged by PRS.
+
+% UR5-wall  OPTIONAL PART imo
+%\begin{wrapfigure}[16]{r}{7cm}
+%\centering
+%\includegraphics[width=7cm]{ur5-wall-traj-rank-loss.png}
+%\caption{UR5}
+%\label{ur5-wall}
+%\end{wrapfigure}
+%In a much more cluttered environment Figure~\ref{ur5-wall} where the robot moves 
+%throught a hole in a wall and between two cylinders, path reduction only manages to 
+%reduce some but not 
+%all detours. This behavior can be explained by a path over-constraining due to the 
+%important proximity of the robot to the obstacles during the whole path, and 
+%because there are not so many DOF to play on to easily avoid collisions and 
+%return a small-cost solution.
+%We advance this explanation because on the two following high-DOF examples of the 
+%video (and Figure~\ref{pr2_final}) based on PR2, results are better 
+%in terms of path length and quality.
+
+% baxter:
+An problem involving a Baxter-like\footnote{A torso rotation was added.} robot manipulating in an office environment is shown Figure~\ref{fig:trajectories} bottom right.
+The robot starts with its end-effectors above the computer and has to turn and reach the shelf.
+According to the left-hand trajectories quality (see the video for the full motion) and path lengths, 9.10 (initial RRT-connect), 7.52 (RS), 8.41 (PRS) and 2.69 (GB with $\alpha_{init}=0.1$), the GB optimizer is the most successfull in this case.
+
+\vspace{0.4cm}
+
+On the three following high-DOF examples involving the PR2 robot, we argue that 
+results are better in terms of path quality, in the fact that parasite DOF are 
+removed.
 
@@
 40-DOF PR2 simply has to cross its arms from 
-the left arm up position to the right arm up one. Lengths of the initial path, the 
-RS optimized path and the GB optimized path are 
-respectively 14.5, 6.34 and 3.62. The RRT-connect planner 
+the left arm up position to the right arm up one, without any assumption on the DOF or group of DOF to choose to plan or optimize (i.e. no DOF is locked). 
+The RRT-connect planner 
 returns detours and activates non-useful DOF such as the head, the torso lift 
-and the translation on 
-the ground. Once again, RS will hardly optimize the mobile base 
-translation of the robot and other unnecessary DOF uses, unlike our 
-optimized-path which mainly results in moving the arms as expected 
-(Figure~\ref{pr2_final} right).
-%The removal 
-%of the translation is also a consequence of our cost ponderation as described in 
-%the previous section.
-% NON ???
+and the translation on the ground. Such behavior induces a high initial path length 
+30.6. RS hardly optimizes the mobile base 
+translation (Figure~\ref{pr2_final} middle) of the robot and other unnecessary DOF 
+uses, resulting in a path length of 4.08.
+Whereas the GB optimized-path mainly 
+results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating 2 collision-constraints between the arms. The PRS 
+result, only presented in the video, cancels the mobile base motion, but not all 
+unnecessary DOF. Moreover, the arms motion is very wide and not as optimized as in 
+the GB trajectory.
+
+%The removal of the translation is also a consequence of our cost ponderation as described in the previous section.  NON ???
 
@@
 We obtain similar results on the PR2 performing a manipulation task 
-in a kitchen environment (see the video). The robot has to move 
+in a kitchen environment (see the video and Figure~\ref{fig:trajectories} middle 
+left to visualize the trajectory of the right gripper). The robot has to move 
 his hands from the top to the bottom of a table.
-Our optimizer manages to reduce the initial length 15.8 from the visibility-PRM 
+Our optimizer manages to reduce the initial length 12.7 from the visibility-PRM 
 planner and improves the path quality 
-just adding transformation constraints between the table and the 
+just adding orthogonality constraints between the table and the 
 robot's arms and grippers. Thus, the robot just slightly moves 
 backward and uses its arm DOF to avoid the table, instead of 
-processing a large motion to get away from the table. With GB, the path length is downed to 4.59, against 8.24 for RS.
+processing a large motion to get away from the table. With GB, the path length is 
+downed to 3.15, against 5.47 for RS.
+Another example of PR2 working in the kitchen, going from the set to the fridge 
+door, is presented Figure~\ref{fig:trajectories} bottom left with the mobile 
+trajectories. This time, GB path length is higher that RS one, but the resulting 
+path remains acceptable.
+
+\vspace{0.4cm}
 
 Optimization computation time and path length averages are presented for 30 runs of 
-the PR2-crossing-arms in the Table~\ref{tab:results}. In the same way, time 
-computation averages for the PR2-in-kitchen example have been 
-calculated: $59.5\,\text{s}$ (GB) and $76.2\,\text{s}$ (RS).
-Thus our method presents 
+each example in the Table~\ref{tab:results}.
+It appears that our method presents 
 similar computation times and path lengths for mobile manipulation tasks. However, 
-the disappearance of the mobile base motion in the gradient-based optimized path is 
-shown in the right column of Table~\ref{tab:results}. Thus the problem addressed 
-Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
+the complete disappearance of the mobile base motion in the PRS and GB optimized 
+paths is highlighted in the right column of Table~\ref{tab:results}. Thus the 
+problem addressed Figure~\ref{decoupled_DOF_optimization} is solved considering this 
+DOF.
 
-\begin{figure}[t]
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
     \centering
-	\includegraphics[width=7.0cm,height=6.8cm]{pr2_orth_constr_results.png}
-	\includegraphics[width=7.0cm,height=6.8cm]{ur2_orth_constr_results.png}\\
-	\vspace{0.05cm}
-	\includegraphics[width=7.0cm,height=6.3cm]{ur5sph_orth_constr_results.png}
-	\includegraphics[width=7.0cm,height=6.3cm]{puzzle_orth_constr_results.png}
-	\caption{Trajectories of wrists, end-effectors or center allong different paths: the initial path is represented in red, the RS optimized path in blue and the GB optimized path ($\alpha_{init}=0.1$) in green.}
-	\label{trajectories}
+    \includegraphics[width=6.5cm,height=5.5cm]{ur2_orth_constr_results.png}
+    \caption*{Double-arm robot}
+    \vspace{1ex}
+  \end{minipage}%%
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=6.5cm,height=5.5cm]{puzzle_orth_constr_results.png}
+    \caption*{Freeflyer-puzzle robot}
+    \vspace{1ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=6.5cm,height=5.5cm]{pr2-kitchen1_orth_constr_results.png}
+    \caption*{PR2-in-kitchen-1}
+    \vspace{1ex}
+  \end{minipage}%% 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=6.5cm,height=5.5cm]{ur5-spheres_orth_constr_results.png}
+    \caption*{UR5-with-spheres}
+    \vspace{1ex}
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=6.5cm,height=5.5cm]{pr2_kitchen2_orth_constr_results.png}
+    \caption*{PR2-in-kitchen-2}
+  \end{minipage}%%
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=6.5cm,height=5.5cm]{baxter_orth_constr_results.png}
+    \caption*{Baxter-in-office}
+  \end{minipage}
+  \caption{Trajectories of end-effectors, mobile bases or centers allong the 
+  different paths: the initial path is represented in red, the RS optimized path in 
+  blue and the GB optimized path in green. The full robots 
+  motions can also be visualized in the joined video. The trajectories comparison 
+  highlights the optimization sucess of our method, which manages to deliver a 
+  shorter or similar path compared to RS output}
+  \label{fig:trajectories}
 \end{figure}
-
-Trajectories results of the robots end-effectors or center are presented 
-Figure\ref{trajectories}.
-
-DESCRIBE TRAJECTORIES
 
@@
     \midrule
-    & Initial &  & $14.3$  &\\
-    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $  &\\
-    \textbf{puzzle} & PRS & $6.53\,\text{ms}$ & $14.3$  &\\
-    & RS & $68.3\,\text{ms}$ & $6.05$  &\\
-    \midrule
-    & Initial &  & $11.3$  &\\
-    \textbf{Double-arm} & GB & $318\,\text{ms}$ & $4.95$  &\\
-    \textbf{5-DOF} & PRS & $8.42\,\text{ms}$ & $8.73$  &\\
-    & RS & $61.9\,\text{ms}$ & $6.96$  &\\
-    \midrule
-    & Initial &  & $10.8$  & $8.03$\,\text{m}\\
-    \textbf{PR2-crossing-} & GB & $24.8\,\text{s}$ & $1.90$  & $0\,\text{m}$\\
-    \textbf{arms} & PRS & $41.1\,\text{s}$ & $ 4.73$ & $0\,\text{m}$\\
-    & RS & $31.5\,\text{s}$ & $1.60$ & $2.70\,\text{m}$\\
-    \midrule
-    & Initial &  & $15.0$ &\\
-    \textbf{PR2-in-} & GB & $288\,\text{s}$ & $4.40$ &\\
-    \textbf{kitchen} & PRS & $431\,\text{s}$ & $12.1$ &\\
-    & RS & $95.6\,\text{s}$ & $5.33$ &\\
+    & Initial &  & $21.9$  &\\
+    \textbf{2D slidding-robot} & GB & $9.29\,\text{ms}$ & $17.0$  &\\
+    \textbf{(Fig.~\ref{2D_long} right)} & PRS & $0.336\,\text{ms}$ & $21.7$  &\\
+    & RS & $1.89\,\text{ms}$ & $6.05$  &\\
     \midrule
     & Initial &  & $ $  &\\
-    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ &\\
+    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $  &\\
+    \textbf{puzzle} & PRS & $\,\text{ms}$ & $ $  &\\
+    & RS & $\,\text{ms}$ & $ $  &\\
+    \midrule
+    & Initial &  & $10.8$  &\\
+    \textbf{Double-arm} & GB & $10.3\,\text{ms}$ & $4.79$  &\\
+    \textbf{5-DOF} & PRS & $2.17\,\text{ms}$ & $8.08$  &\\
+    & RS & $5.33\,\text{ms}$ & $6.24$  &\\
+    \midrule
+    & Initial &  & $12.6$  & $9.57$\,\text{m}\\
+    \textbf{PR2-crossing-} & GB & $1.28\,\text{s}$ & $2.57$  & $0\,\text{m}$\\
+    \textbf{arms} & PRS & $2.63\,\text{s}$ & $ 5.31$ & $0\,\text{m}$\\
+    & RS & $1.92\,\text{s}$ & $3.59$ & $2.63\,\text{m}$\\
+    \midrule
+    & Initial &  & $15.3$ &\\
+    \textbf{PR2-in-} & GB & $19.2\,\text{s}$ & $4.53$ &\\
+    \textbf{kitchen-1} & PRS & $23.5\,\text{s}$ & $12.2$ &\\
+    & RS & $9.88\,\text{s}$ & $5.75$ &\\
+    \midrule
+    & Initial &  & $ $ &\\
+    \textbf{UR5-} & GB & $\,\text{s}$ & $ $ &\\
     \textbf{with-spheres} & PRS & $\,\text{s}$ & $ $ &\\
@@
     & Initial &  & $ $ &\\
-    \textbf{6-DOF-arm-} & GB & $\,\text{s}$ & $ $ &\\
-    \textbf{through-window} & PRS & $\,\text{s}$ & $  $ &\\
+    \textbf{UR5-} & GB & $\,\text{s}$ & $ $ &\\
+    \textbf{industrial-example} & PRS & $\,\text{s}$ & $  $ &\\
     & RS &  $\,\text{s}$ & $ $ &\\
     \midrule
-    & Initial &  & $19.0$ &\\
-    \textbf{Baxter-in-} & GB & $2449\,\text{s}$ & $5.85$ &\\
-    \textbf{office} & PRS & $638\,\text{s}$ & $17.5$ &\\
-    & RS & $139\,\text{s}$ & $13.3$ &\\
+    & Initial &  & $ $ &\\
+    \textbf{Baxter-in-} & GB & $\,\text{s}$ & $ $ &\\
+    \textbf{office} & PRS & $\,\text{s}$ & $ $ &\\
+    & RS & $\,\text{s}$ & $ $ &\\
     \bottomrule
@@
 the video. For each run, a solution path is planned by Visibility-PRM as initial 
-guess for each optimizer. $\alpha_{init}$ is set to 0.1.
+guess for each optimizer.
 For the PR2-crossing-arms example only, the right column represents the 
@@
 
-FUTURE WORK
+In a future work, we plan to study a combination of a random shortcut and the gradient-based optimizers since they seem to share a complementarity. .....
 

### 2015-10-21 paper.tex
Add missing results, commit table, modify figures

@@
 \usepackage{mathptmx} % assumes new font selection scheme installed
-%\usepackage{times} % assumes new font selection scheme installed
 \usepackage{amsmath} % assumes amsmath package installed
@@
 \makeatletter
-%\let\NAT@parse\undefined % allow hyperref to work with ieeeconf.cls for bib numbers
 \makeatother
 \usepackage{hyperref}
-%\usepackage{cite}
 
@@
 
-\jvol{00} \jnum{00} \jyear{2015} \jmonth{October}
+\jvol{30} \jnum{15} \jyear{2016} \jmonth{August}
 
@@
 
-\author{Myl\`{e}ne Campana$^{a,b}$ $^{\ast}$ \thanks{$^\ast$Corresponding author. Email: mcampana@laas.fr}, Florent Lamiraux$^{a,b}$ and Jean-Paul Laumond$^{a,b}$
+\author{Myl\`{e}ne Campana$^{\ast}$ \thanks{$^\ast$Corresponding author. Email: mcampana@laas.fr}, Florent Lamiraux and Jean-Paul Laumond
 \\\vspace{6pt}
-$^{a}${\em{LAAS-CNRS, Toulouse, France}};
-$^{b}${\em{ This work has been supported by the project ERC Advanced Grant 340050 Actanthrope and by the FP~7 project Factory in a Day under grant agreement n°~609206 }}
+{\em{LAAS-CNRS, 7 av. Colonel Roche, 31400 Toulouse, FRANCE}}
 }
@@
 Figure~\ref{decoupled_DOF_optimization} problem.
-
-%Unlike our method however, the iterations are composed of 
-%Cubic-B-splines. The benefit of this method is that the result is a 
-%differentiable path.
 
@@
 
-%Let $\translation^{*}\in \real^3$ denotes the position between frames of $
-%\body_1$  and $\body_2$ in configuration $\xx_i(\tcolli)$. Let $\translation \in 
-%\real^3$ (resp. $R\in SO(3)$) denote the same position (resp. rotation) in 
-%configuration $\xx (\tcolli)$ where $\xx$ is any path in $\CS^{wp}$.
-
-%Let $M^{*}\in SE(3)$ denote the homogeneous transformation between frames of $
-%\body_1$  and $\body_2$ in configuration 
-%$\xx_i(\tcolli)$. Let $M\in SE(3)$ denote the same transformation in configuration 
-%$\xx (\tcolli)$ where $\xx$ is any path 
-%in $\CS^{wp}$. Let $R\in SO(3)$ and $\translation\in\real^3$ 
-%(resp. $\translation^{*}$) be the rotation 
-%and translation parts of $M$ (resp. $M^{*}$).
-
-%We denote by $F$ the path constraint defined by the mapping from $\CS^{wp}$ to $
-%\real^3$ defined by
-%\begin{equation}\label{eq:constraint}
-%F (\xx) =  \mathbf{u}^T R^T (\translation^{*} - \translation)
-%\end{equation}
-%where the product $R^T (\translation^{*} - \translation)$ is used to express the translations difference in the global frame ${\cal R}_0$.
-
-\vspace{0.2cm}
-
-%Therefore, the Jacobian $J_F$ of constraint~(\ref{eq:constraint}) with respect to 
-%$x$ is
-%$$
-%J_F = \left(\begin{array}{cccccc}
-%0 & \cdots & (1-\beta)\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \
-%\beta\,\mathbf{u}^T \frac{\partial F}{\partial q} (\xx_j(t_{coll\,i})) & \cdots & 0
-%\end{array}
-%\right)
-%$$
-
-\vspace{0.2cm}
+\vspace{0.4cm}
 
@@
 	\centering
-	\includegraphics[width=7.8cm]{contact_points6.png}
-	\includegraphics[width=8cm]{contact_points2potential.png}
+	\includegraphics[width=7.8cm]{contact_points6.pdf}
+	\includegraphics[width=8cm]{contact_points2potential.pdf}
 	\caption{Path-optimization results on 2D-ponctual robots, moving around 
-	green obstacles. Initial paths are in red, optimized ones in black. Grey paths 
-	represent intermediate iterations, red and black dots resp. $\po_1$ and $\po_2$ 
-	points, red segments $\mathbf{u}$ and blue circles collision-free 
-	configurations 
-	of the backtracked paths 
-	where constraints have been computed.}
+	green obstacles. Initial paths are dashed and crosses represent contact points 
+	related to collision-constraints.}
 	\label{2D_long}
@@
 	\includegraphics[width=7.5cm]{local_box_optim.png}
-	\caption{Case of a long initial path (above) containing a small part that can 
+	\caption{Case of a long initial path in red (above) containing a small part 
+	that can 
 	be optimized (below). Random shortcut is unlikely to optimize the part 
@@
 collision-constraints. Unless another value is given, $\alpha_{init}$ is set to 0.2.
-
-%\begin{figure}[b]
-%	\centering
-%	\includegraphics[width=5cm,height=3.5cm]{ur5_v2.png}
-%	\caption{6-DOF robot arm on which our path-optimization has been tested.}
-%	\label{ur5}
-%\end{figure}
 
@@
 
-%But we admit that restrincting the number of random shortcut iterations is 
-%neither a acceptable criterion because of randomness.
-%We also discuss the good properties and limits of our algorithm.
-
 \vspace{0.4cm}
@@
 1.98 by GB and unchanged by PRS.
-
-% UR5-wall  OPTIONAL PART imo
-%\begin{wrapfigure}[16]{r}{7cm}
-%\centering
-%\includegraphics[width=7cm]{ur5-wall-traj-rank-loss.png}
-%\caption{UR5}
-%\label{ur5-wall}
-%\end{wrapfigure}
-%In a much more cluttered environment Figure~\ref{ur5-wall} where the robot moves 
-%throught a hole in a wall and between two cylinders, path reduction only manages to 
-%reduce some but not 
-%all detours. This behavior can be explained by a path over-constraining due to the 
-%important proximity of the robot to the obstacles during the whole path, and 
-%because there are not so many DOF to play on to easily avoid collisions and 
-%return a small-cost solution.
-%We advance this explanation because on the two following high-DOF examples of the 
-%video (and Figure~\ref{pr2_final}) based on PR2, results are better 
-%in terms of path length and quality.
 
@@
 the GB trajectory.
-
-%The removal of the translation is also a consequence of our cost ponderation as described in the previous section.  NON ???
 
@@
 Optimization computation time and path length averages are presented for 30 runs of 
-each example in the Table~\ref{tab:results}.
-It appears that our method presents 
-similar computation times and path lengths for mobile manipulation tasks. However, 
-the complete disappearance of the mobile base motion in the PRS and GB optimized 
-paths is highlighted in the right column of Table~\ref{tab:results}. Thus the 
-problem addressed Figure~\ref{decoupled_DOF_optimization} is solved considering this 
-DOF.
+each example in the Table~\ref{tab:results}. In some cases $\alpha_{init}$ was reduced to comply with very narrow passages, or increased in the opposite case. 
+It appears that, for non-high DOF examples, computation time is higher using the GB optimization.
+The worst case shows up for the puzzle problem where the number of constraints of GB is typically important.
+On the other hand, our method almost always provides the shortest paths, whereas it is not even the best criterion to quantify the parasite motions disappearance.
+Such results are the artefact of the termination conditions of RS and PRS, length reduction or affordable computation time.
+This is particularly visible for PRS which easily gives up in some problems, such as the 2D slidding-robot, the UR5-with-spheres or the puzzle.
+Finally, the complete vanishing of the mobile base motion in the PRS and GB 
+optimized paths is highlighted in the right column of Table~\ref{tab:results}. Thus 
+the problem addressed Figure~\ref{decoupled_DOF_optimization} is solved considering 
+this DOF.
 
@@
 
-\begin{table}
+\begin{table}[ht]
 \centering
@@
     \textbf{(Fig.~\ref{2D_long} right)} & PRS & $0.336\,\text{ms}$ & $21.7$  &\\
-    & RS & $1.89\,\text{ms}$ & $6.05$  &\\
+    & RS & $1.89\,\text{ms}$ & $17.7$  &\\
     \midrule
-    & Initial &  & $ $  &\\
-    \textbf{Freeflyer-} & GB & $\,\text{ms}$ & $ $  &\\
-    \textbf{puzzle} & PRS & $\,\text{ms}$ & $ $  &\\
-    & RS & $\,\text{ms}$ & $ $  &\\
+    & Initial &  & $18.8$  &\\
+    \textbf{Freeflyer-} & GB ($\alpha_{init}=0.05$) & $1.02\,\text{s}$ & $9.65$  &\\
+    \textbf{puzzle} & PRS & $1.13\,\text{ms}$ & $18.8$  &\\
+    & RS & $4.72\,\text{ms}$ & $10.3$  &\\
     \midrule
@@
     & RS & $5.33\,\text{ms}$ & $6.24$  &\\
+    \midrule
+    & Initial (RRT) &  & $4.97$ &\\
+    \textbf{UR5-} & GB ($\alpha_{init}=0.4$) & $2.12\,\text{s}$ & $2.97$ &\\
+    \textbf{with-spheres} & PRS & $0.884\,\text{s}$ & $4.56$ &\\
+    & RS & $2.89\,\text{s}$ & $2.99$ &\\
+    \midrule
+    & Initial &  & $4.83$ &\\
+    \textbf{UR5-} & GB ($\alpha_{init}=0.4$) & $314\,\text{ms}$ & $2.25$ &\\
+    \textbf{industrial-example} & PRS & $30.2\,\text{ms}$ & $4.53$ &\\
+    & RS &  $95.6\,\text{ms}$ & $2.16$ &\\
+    \midrule
+    & Initial &  & $20.2$ &\\
+    \textbf{Baxter-in-} & GB & $22.7\,\text{s}$ & $6.65$ &\\
+    \textbf{office} & PRS & $9.49\,\text{s}$ & $18.9$ &\\
+    & RS & $3.13\,\text{s}$ & $11.9$ &\\
     \midrule
@@
     & RS & $9.88\,\text{s}$ & $5.75$ &\\
-    \midrule
-    & Initial &  & $ $ &\\
-    \textbf{UR5-} & GB & $\,\text{s}$ & $ $ &\\
-    \textbf{with-spheres} & PRS & $\,\text{s}$ & $ $ &\\
-    & RS & $\,\text{s}$ & $ $ &\\
-    \midrule
-    & Initial &  & $ $ &\\
-    \textbf{UR5-} & GB & $\,\text{s}$ & $ $ &\\
-    \textbf{industrial-example} & PRS & $\,\text{s}$ & $  $ &\\
-    & RS &  $\,\text{s}$ & $ $ &\\
-    \midrule
-    & Initial &  & $ $ &\\
-    \textbf{Baxter-in-} & GB & $\,\text{s}$ & $ $ &\\
-    \textbf{office} & PRS & $\,\text{s}$ & $ $ &\\
-    & RS & $\,\text{s}$ & $ $ &\\
     \bottomrule
@@
 \caption{Average results for 30 runs of several examples presented in the paper or 
-the video. For each run, a solution path is planned by Visibility-PRM as initial 
-guess for each optimizer.
+the video. For each run, a solution path is planned by Visibility-PRM (unless `RRT' 
+for RRT-connect is mentioned) as initial guess for each optimizer.
 For the PR2-crossing-arms example only, the right column represents the 
@@
 
-In a future work, we plan to study a combination of a random shortcut and the gradient-based optimizers since they seem to share a complementarity. .....
-
-\bibliographystyle{tADR} % plain, alpha, abbr, unsrt
+In a future work, we plan to study .....
+
+\section*{Funding}
+This work has been supported by the project ERC Advanced Grant 340050 Actanthrope and by the FP~7 project Factory in a Day under grant agreement n°~609206.
+
+\bibliographystyle{tADR}
 \bibliography{paper}

### 2015-10-22 paper.tex
Fix mistakes, subfigures numbers

@@
 \usepackage{caption}
-\usepackage{wrapfig}
 \usepackage{color}
@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\hypersetup{
+    colorlinks,
+    linkcolor={red!80!black},
+    citecolor={blue!80!black},
+    urlcolor={blue!80!black}
+}
 
@@
 \begin{itemize}
-\item requiring only collision checking, and not any time-consuming obstacle distance computation,
+\item requiring only collision checking, and not any time-consuming obstacle distance computation nor geometry simplification,
 \item constraining only part of the configuration variables that may cause a 
 collision, and not the entire configurations, and
-\item reducing parasite motions that are not useful for the problem resolution.
+\item reducing parasite motions that are not useful for the problem resolution without any assumption.
 \end{itemize}
@@
 results include navigation and manipulation tasks, e.g. an robotic arm 
-manipulating throw a window and a PR2 robot working in a kitchen environment, 
-and comparisons with a random shortcut optimizer.
+manipulating throw in boxes and a PR2 robot working in a kitchen environment. 
+Comparisons with a random shortcut optimizer and a partial shortcut are also provided.
 
@@
 than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
-today randomly sample the system configuration space~\cite{ref-spaceconfig} in
-order to find a collision-free path. The main issue using these techniques is
- that the computed path makes unnecessary detours and needs to be post-processed 
- before being executed by a virtual or real robot. Alternative
-strategies exist however.
+randomly sample the system configuration space~\cite{ref-spaceconfig} in
+order to find a collision-free path, linking the initial and final configurations. 
+The main issue using these techniques is that the computed path makes unnecessary 
+detours and needs to be post-processed before being executed by a virtual or real 
+robot. Alternative strategies exist however.
 \begin{itemize}
@@
 nearest obstacle distance. Most of these planners are using non-linear
-optimization~\cite{BettsNonlinopt} under constraints.
-
-Such planners provide close-to-optimality paths and have smaller time computation
+optimization~\cite{BettsNonlinopt} under constraints. Such planners provide close-to-optimality paths and have smaller time computation
 for easy problems, but they are mostly unable to solve narrow passage issues.
@@
 tuning of the termination condition (see 
-Algorithm~\ref{algo:random-shortcut}) and is no efficient due to randomness for long 
+(Algorithm~\ref{algo:random-shortcut})) and is no efficient due to randomness for long 
 trajectories where only a small part needs to be optimized. 
-Figure~\ref{decoupled_DOF_optimization} presents another situation where RS will 
+(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS will 
 always fail to optimize the initial path, since it cannot decouple the 
@@
 	decrease the path length. The optimal path belongs to the $x-y$ plane 
-	containing $q_{init}$ and $q_{end}$.}
+	containing $\conf_{init}$ and $\conf_{final}$.}
 	\label{decoupled_DOF_optimization}
@@
 \section{RELATED WORK}
-
-CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
+Previous work on path optimization has been conducted. CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
 input. It minimizes a time invariant cost function using efficient covariant
@@
 ~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
-generation with local holonimic considerations. But this method remains only locally 
-optimal, and is not addressing high-DOF problems.
+generation with local holonimic considerations. But this method remains only 
+locally optimal, and is not addressing high-DOF problems.
 ~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
 using medial axis retractation for clearance and a PRS (applied only on random DOF) 
-which can address the problem of Figure~\ref{decoupled_DOF_optimization} but is 
+which can address the problem of (Figure~\ref{decoupled_DOF_optimization}) but is 
 relatively slower than RS. Moreover, PRS is not using the information returned by 
@@
 In some way, our method shares similarities with~\cite{PanSmoothSplineShort} 
-since this latter method relies on collision checking and backtracks when an 
+since this latter work relies on collision checking and backtracks when an 
 iteration is detected in collision, instead of trying to constantly satisfy 
-distance constraints. Unlike our method however, interpolating configuration -which 
-can be viewed as constraints- are freezing the whole robot configuration instead of 
-a relevant subpart. Thus, this method cannot solve the 
-Figure~\ref{decoupled_DOF_optimization} problem.
+distance constraints. Unlike our method however, anti-collision constraints are 
+handled by interpolating configurations which, at some points of the trajectory, 
+freeze the whole robot configuration instead of a relevant subpart. Thus, this 
+method cannot solve the (Figure~\ref{decoupled_DOF_optimization}) problem.
 
@@
 in the frame of its parent. In the examples shown in this paper, we consider 4 
-types of joints described in Table~\ref{tab:joints}.
+types of joints described in (Table~\ref{tab:joints}).
 \begin{table}[b]
@@
 constant velocity'' makes sense for the four types of joints defined in 
-Table~\ref{tab:joints}. We refer to~\cite{riemanian-optim2008} Chapter~5 for 
+(Table~\ref{tab:joints}). We refer to~\cite{riemanian-optim2008} Chapter~5 for 
 details about Riemanian geometry.
@@
 corresponds to the unconstrained function \texttt{computeIterate} in 
-Algorithm~\ref{algo:gradient}.
+(Algorithm~\ref{algo:gradient}).
 When a collision is detected, we introduce a constraint and perform a new 
@@
 collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
-paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in Algorithm~\ref{algo:gradient}.
+paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}).
 
@@
 	to be in collision with the obstacle, the first colliding configuration 
-	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The corresponding constraint will be computed in configuration $\xx_{i}(\tcolli)$.}
+	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
+	corresponding constraint will be computed in the backtracked configuration 
+	$\xx_{i}(\tcolli)$.}
 	\label{optim_grad}
@@
 Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
-interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. Let 
+interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
 us denote by $\tcolli$ the abscissa of the first collision detected on path 
-$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free (see 
-Figure~\ref{optim_grad}). Thus in 
+$\xx_{i+1}$, which previous iteration $\xx_i$ was collision-free. Thus in 
 configuration $\xx_{i+1}(\tcolli)$ a collision has been 
@@
 $\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
-linking the contact points in configuration 
-$\xx_{i}(t_{coll\ i})$ (the latest collision-free path), expressed in the local 
-frame of $\body_1$:
+linking the contact points $\po_1$ and 
+$\po_2$ in configuration $\xx_{i}(t_{coll\ i})$, $\xx_{i}$ being the latest collision-free path:
 $$
@@
 $$
-The constraint is defined by
+The constraint is defined for any path $\xx$ by:
 \begin {equation}\label{eq:new-constraint}
@@
 \end {equation}
-As this constraint is non-linear, it is linearized around $\xx_{i}$.
-
-Let $q_{k\,i}$ denote the waypoint $k$ along path $x_i$.
+As this constraint is non-linear, it is linearized around $\xx_{i}$. Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
 The constraint on the position at $t_{coll\,i}$ is a function that depends only
-on two waypoints $q_{k\,i}$, $q_{k+1\,i}$, with $0\leq k\leq n$. There exists 
+on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq n$. There exists 
 $\beta\in[0,1]$ such that for any $j$:
@@
 and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
-Algorithm~\ref{algo:gradient}.
+(Algorithm~\ref{algo:gradient}).
 Then, a line is added in the contraint Jacobian matrix (function 
@@
 programs (QP) (function \texttt{computeIterate} in 
-Algorithm~\ref{algo:gradient}).
+(Algorithm~\ref{algo:gradient})).
 
@@
 \end{equation}
-Let us assume that
+Let us assume that:
 \begin{equation}\label{eq:asumption-1}
@@
 \end{equation}
-We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then
+We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
 $$
@@
 $$
-From~(\ref{eq:decreasing-kernel}), we deduce that
+From~(\ref{eq:decreasing-kernel}), we deduce that:
 $$
@@
 $$
-This expression represents the velocity of point $\po_2$ in the reference frame of Body~1 projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
+This expression represents the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
 \begin{equation}\label{eq:trajectory}
@@
 reached for $t=1$.
-This expression being equal to 0 requires either
+This expression being equal to zero requires either
 \begin {enumerate}
 \item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
-\item the velocity of $\po_2$ in the reference frame of Body~1 is orthogonal to 
+\item the velocity of $\po_2$ in the reference frame of $\body_1$ is orthogonal to 
 $\mathbf{u}$ at contact ($t=1$).
@@
 	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
-	\State $\xx \gets \xx_{Free} + \frac{\alpha}{2}\,\p$
+	\State $\xx \gets \xx_{Free} + 0.5\alpha\,\p$
 	\If{(\texttt{validatePath}($\xx_1$))}
@@
 converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
-body~1 along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
+$\body_1$ along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
 contact. We already handled this case previously, admitting that the velocity norm 
-would not vanish. Therefore, is~\ref{eq:asumption-1} verified.
+would not vanish. Therefore, Equation~(\ref{eq:asumption-1}) is verified.
 
@@
 
-In this part we describe the GB path-optimizer Algorithm~\ref{algo:gradient} 
+In this part we describe the GB path-optimizer (Algorithm~\ref{algo:gradient})
 according to the previous step definitions. The main difficulty here is to handle 
 the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
-be traveled through, and also to return a collision-free path. Typically, we 
+be traveled through. Typically, we 
 chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
@@
 This part gathers optimization results performed on the planning software 
-Humanoid Path Planner~\cite{hpp}. The initial trajectory is obtained with two 
+Humanoid Path Planner~\cite{hpp}. Initial paths are obtained with two 
 kind of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
@@
 	green obstacles. Initial paths are dashed and crosses represent contact points 
-	related to collision-constraints.}
+	$\po_c$ related to collision-constraints. Note that, on the left, the detour 
+	completely disappears.}
 	\label{2D_long}
@@
 
-Figure~\ref{2D_long} shows several iterations of our optimizer on 2D cases. Since 
-in this special case, orthogonal constraints are equivalent to forbidden motion 
-direction, we can verify that our optimized path is applying 
-all computed constraints. We can also understand which collisions have led to 
-the constraints.
-
-Besides, these examples give a better understanding of how the tuning of $\alpha$ 
+(Figure~\ref{2D_long}) shows the result of our optimizer on 2D cases.
+Represented collisions leading to constraints permit to understand how the path is kept out of the obstacles.
+Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle the constraint must be created, as at some parts the robot is very close to obstacles.
+
+Besides, these examples give a better understanding of how the tuning of 
+$\alpha_{init}$ 
 has to balance lot of iterations and relevant collision-constraint addition. For 
@@
 
-Figure~\ref{local_box_optim} illustrates a \textit{very long} path example that RS 
+(Figure~\ref{local_box_optim}) illustrates a \textit{very long} path example that RS 
 or PRS will not manage to 
@@
 configurations in the box. GB succeeds to optimize the 
-path contained in the box, with the following cost coefficients 
+path contained in the box, with the following cost coefficients:
 $$
@@
 minimum as at 
-initial path, represented by the waypoints ($\conf_{i\,0})_{1\leq i\leq n+1}$.
+initial path, represented by the waypoints ($\conf_{i\,0})_{1\leq i\leq wp+1}$.
 Without these coefficients, the path that minimizes the cost corresponds to a 
 straight line with the waypoints equidistantly allocated, which is not adapted for 
-the Figure~\ref{local_box_optim} type of problems with a local passage very
+the (Figure~\ref{local_box_optim}) type of problems with a local passage very
 constrained by obstacles. Note that this cost is also working with all other 
-examples presented in this paper.
+examples presented in this paper, and provides better quality results than the original cost.
 
@@
 	\centering
-	\includegraphics[width=7.5cm]{local_box_optim.png}
-	\caption{Case of a long initial path in red (above) containing a small part 
+	\includegraphics[width=10cm]{local_box_optim.pdf} % MOVE ROBOT TO THE ZOOM !
+	\caption{Case of a long initial path (above) containing a small part 
 	that can 
-	be optimized (below). Random shortcut is unlikely to optimize the part 
-	containing detours in the box, whereas our method 
-	succeeds (in blue). This type of issue is common in navigation problems in 
+	be optimized (below). Random shortcut is unlikely to optimize the initial 
+	dashed part containing detours in the box, whereas our method 
+	succeeds (in blue). This type of issue is common in navigation problems, in 
 	environments with long corridors.}
@@
 
-We also experiment our algorithm on more complex robots, with transformation 
-collision-constraints. Unless another value is given, $\alpha_{init}$ is set to 0.2.
+We also experiment our algorithm on more complex robots and environments. Unless another value is given, $\alpha_{init}$ is set to 0.2.
 
@@
 
-In the included video and Figures~\ref{pr2_final},~\ref{fiad} and
- ~\ref{fig:trajectories}, we present multiple situations where the GB algorithm has 
-been tested and compared to RS~\cite{randomShortcutHPP}
+In the included video and (Figures~\ref{pr2_final},~\ref{fiad} 
+and~\ref{fig:trajectories}), we present multiple situations where the GB algorithm 
+is tested and compared to RS~\cite{randomShortcutHPP}
 (Algorithm~\ref{algo:random-shortcut}) and PRS ~\cite{partialrandomShortcutHPP} 
@@
 The termination condition of RS allows it to try 
-shortening the path until 5 iterations of non-improvement are reached 
-(corresponding to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}).
+shortening the path until five iterations of non-improvement are reached 
+(corresponding to $maxNbFailures$ in (Algorithm~\ref{algo:random-shortcut})).
 For our implementation of PRS, first a shortcut is tried on each DOF between 
-$\conf_0$ and $\conf_{n+1}$. 
+$\conf_0$ and $\conf_{wp+1}$. 
 Then, for each DOF, configurations are shot as in 
-Algorithm~\ref{algo:random-shortcut} and the steering-method returns a path made by 
+(Algorithm~\ref{algo:random-shortcut}) and the steering-method returns a path made by 
 an interpolation on only 
-the current DOF and the previous subpath for other DOF. If this path is collision-
-free, it is replaced as in RS. This process is also stopped when 5 iterations of 
-non-improvement are reached.
+the current DOF and the previous subpath for other DOF. If this path is 
+collision-free, it is replaced as in RS. This process is also stopped when five 
+iterations of non-improvement are reached.
 
@@
 Before entering the manipulators examples, the GB algorithm is tested and compared 
-on a freeflyer puzzle example (see Figure~\ref{fig:trajectories} top right), very 
-encountered in the motion planning literature. The puzzle has to cross down the 
+on a problem very encountered in the motion planning literature: a freeflyer puzzle 
+example, corresponding to (Figure~\ref{fig:trajectories:puzzle}).  The puzzle has to cross down the 
 obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
-contains detours above and below the obstacle, as well as small motion in the hole. 
+contains detours above and below the obstacle, as well as small superfluous
+motions in the hole. 
 Lengths of initial, GB optimized, and RS optimized paths are 
-respectively 23.4, 8.53 and 10.2.
-PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole.
-This behaviour was also noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, acting on a group of DOF.
-The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.
+respectively 23.4, 8.53 and 10.2. PRS fails to decrease the path length because in 
+this problem, all DOF are closely related to cross the hole. This behaviour was also 
+noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, 
+acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory 
+even in the hole, contrary to RS which simply copies the initial path when the 
+trajectory gets near of the obstacle. $\alpha_{init}$ is reduced to 0.05 to slightly modify the very constrained trajectory at each iteration.
 
@@
 % double-arm:
-On the double-arm problem (see Figure~\ref{fig:trajectories} top-left), one arm has 
+On the double-arm problem, (Figure~\ref{fig:trajectories:double_arm}), one arm has 
 to get around a cylinder obstacle while the 
@@
 motion. The initial path length 5.17 is decreased to 4.55 by RS, to 4.06 by PRS and 
-to 2.77 by GB using 3 collision-constraints with the cylinder obstacle.
+to 2.77 by GB using three collision-constraints with the cylinder obstacle.
 
@@
 	constrained in boxes. Trajectories results 
-	are presented (bottom right): initial (RRT) path in red, RS optimization in 
-	blue and GB optimization in green.}
+	are presented (bottom right): initial (RRT-connect) path in red, RS 
+	optimization in blue and GB optimization in green.}
 	\label{fiad}
@@
 % UR5-spheres:
-Some problems were 
+Some problems are 
 involving an a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
 gripper.
-In a relatively free environment (Figure~\ref{fig:trajectories} middle right), 
+In a relatively free environment (Figure~\ref{fig:trajectories:ur5}), 
 results from GB and RS are similar (path length are respectively 2.03 and 1.60). 
@@
 For an UR5 working in a cluttered environment inspired from an industrial issue 
-Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
+(Figure~\ref{fiad}), path optimization efficiently returns a shorter solution, close 
 to the result of RS and to what can be observed in reality~\cite{factory-day-video}. 
@@
 % baxter:
-An problem involving a Baxter-like\footnote{A torso rotation was added.} robot manipulating in an office environment is shown Figure~\ref{fig:trajectories} bottom right.
-The robot starts with its end-effectors above the computer and has to turn and reach the shelf.
-According to the left-hand trajectories quality (see the video for the full motion) and path lengths, 9.10 (initial RRT-connect), 7.52 (RS), 8.41 (PRS) and 2.69 (GB with $\alpha_{init}=0.1$), the GB optimizer is the most successfull in this case.
+An problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
+manipulating in an office environment is shown 
+(Figure~\ref{fig:trajectories:baxter}). The robot starts with its end-effectors 
+above the computer and has to 
+turn and reach the shelf. According to the left-hand trajectories quality (see the 
+video for the full motion) and path lengths, 9.10 (initial RRT-connect), 7.52 (RS), 
+8.41 (PRS) and 2.70 (GB), the GB optimizer is the most successfull in this case.
 
@@
 % analysis PR2 alone:
-In the example presented Figure~\ref{pr2_final} and in the video, the mobile 
+In the example presented (Figure~\ref{pr2_final}) and in the video, the mobile 
 40-DOF PR2 simply has to cross its arms from 
@@
 Whereas the GB optimized-path mainly 
-results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating 2 collision-constraints between the arms. The PRS 
+results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating 
+two collision-constraints between the arms. The PRS 
 result, only presented in the video, cancels the mobile base motion, but not all 
@@
 We obtain similar results on the PR2 performing a manipulation task 
-in a kitchen environment (see the video and Figure~\ref{fig:trajectories} middle 
-left to visualize the trajectory of the right gripper). The robot has to move 
+in a kitchen environment (see the video and (Figure~\ref{fig:trajectories:kitchen1}) 
+to visualize the trajectory of the right gripper). The robot has to move 
 his hands from the top to the bottom of a table.
@@
 robot's arms and grippers. Thus, the robot just slightly moves 
-backward and uses its arm DOF to avoid the table, instead of 
+backward and uses its arms DOF to avoid the table, instead of 
 processing a large motion to get away from the table. With GB, the path length is 
@@
 Another example of PR2 working in the kitchen, going from the set to the fridge 
-door, is presented Figure~\ref{fig:trajectories} bottom left with the mobile 
+door, is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile 
 trajectories. This time, GB path length is higher that RS one, but the resulting 
@@
 Optimization computation time and path length averages are presented for 30 runs of 
-each example in the Table~\ref{tab:results}. In some cases $\alpha_{init}$ was reduced to comply with very narrow passages, or increased in the opposite case. 
-It appears that, for non-high DOF examples, computation time is higher using the GB optimization.
-The worst case shows up for the puzzle problem where the number of constraints of GB is typically important.
-On the other hand, our method almost always provides the shortest paths, whereas it is not even the best criterion to quantify the parasite motions disappearance.
-Such results are the artefact of the termination conditions of RS and PRS, length reduction or affordable computation time.
-This is particularly visible for PRS which easily gives up in some problems, such as the 2D slidding-robot, the UR5-with-spheres or the puzzle.
+each example in the (Table~\ref{tab:results}). In some cases $\alpha_{init}$ was 
+reduced to comply with very narrow passages, or increased in the opposite case. 
+It appears that, for non-high DOF examples, computation time is higher using the GB 
+optimization. The worst case shows up for the puzzle problem where the number of 
+constraints of GB is typically important and $\alpha_{init}$ reduced.
+On the other hand, our method almost always provides the shortest paths, whereas it 
+is not even the best criterion to quantify the parasite motions disappearance.
+Such results are the artefact of the termination conditions of RS and PRS, length 
+reduction or affordable computation time. This is particularly visible for PRS which 
+easily gives up in some problems, such as the 2D slidding-robot, the 
+UR5-with-spheres or the puzzle.
 Finally, the complete vanishing of the mobile base motion in the PRS and GB 
-optimized paths is highlighted in the right column of Table~\ref{tab:results}. Thus 
-the problem addressed Figure~\ref{decoupled_DOF_optimization} is solved considering 
-this DOF.
+optimized paths is highlighted in the right column of (Table~\ref{tab:results}). 
+Thus the problem addressed in the (Figure~\ref{decoupled_DOF_optimization}) is 
+solved considering this DOF.
 
 \begin{figure}[ht]
-  \begin{minipage}[b]{0.5\linewidth}
 	\centering
+	\subfigure[Double-arm]{
 		\includegraphics[width=6.5cm,height=5.5cm]{ur2_orth_constr_results.png}
-    \caption*{Double-arm robot}
-    \vspace{1ex}
-  \end{minipage}%%
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+		\label{fig:trajectories:double_arm}
+	}
+	\subfigure[Freeflyer-puzzle]{
 		\includegraphics[width=6.5cm,height=5.5cm]{puzzle_orth_constr_results.png}
-    \caption*{Freeflyer-puzzle robot}
-    \vspace{1ex}
-  \end{minipage} 
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+		\label{fig:trajectories:puzzle}
+	}
+	\subfigure[PR2-in-kitchen-1]{
 		\includegraphics[width=6.5cm,height=5.5cm]{pr2-kitchen1_orth_constr_results.png}
-    \caption*{PR2-in-kitchen-1}
-    \vspace{1ex}
-  \end{minipage}%% 
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+		\label{fig:trajectories:kitchen1}
+	}
+	\subfigure[UR5-with-spheres]{
 		\includegraphics[width=6.5cm,height=5.5cm]{ur5-spheres_orth_constr_results.png}
-    \caption*{UR5-with-spheres}
-    \vspace{1ex}
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+		\label{fig:trajectories:ur5}
+	}
+	\subfigure[PR2-in-kitchen-2]{
 		\includegraphics[width=6.5cm,height=5.5cm]{pr2_kitchen2_orth_constr_results.png}
-    \caption*{PR2-in-kitchen-2}
-  \end{minipage}%%
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+		\label{fig:trajectories:kitchen2}
+	}
+	\subfigure[Baxter-in-office]{
 		\includegraphics[width=6.5cm,height=5.5cm]{baxter_orth_constr_results.png}
-    \caption*{Baxter-in-office}
-  \end{minipage}
+		\label{fig:trajectories:baxter}
+	}
   \caption{Trajectories of end-effectors, mobile bases or centers allong the 
@@
 \section{CONCLUSIONS}
-We managed to settle a path optimization for decoupled-DOF robots such as mobile 
-manipulators. Our algorithm uses standard numerical tools as collision checking 
-and QP resolution, and correlates them in a 
-simple but effective way, playing on the scalar iteration parameter. Therefore, 
-our method only require few distances computation, so geometry pre-processing or 
-offline optimization are not necessary to remain time competitive. We demonstrate 
-that the optimizer is 
-time-competitive comparing to random shortcut and proposes better quality paths 
-for high-DOF robots, removing unnecessary DOF motions. Our optimizer also manages 
-to reduce a local detour in a long path while random shortcut methods will mostly 
-fail.
-
-In a future work, we plan to study .....
+We managed to settle a path optimization for navigation and manipulation problems, and tested with various robots and environments. Our algorithm uses standard 
+numerical tools as collision checking, linearized one-dimensional constraint
+and QP resolution. It correlates them in a 
+simple but effective way, playing on the scalar iteration parameter. Moreover, 
+our method only requires collision checking, therefore geometry pre-processing or 
+offline optimization are not necessary to remain time-competitive. We demonstrate 
+that the optimizer may be 
+time-competitive compared to random shortcut and proposes better quality paths, reducing the path length and removing unnecessary DOF motions.
+Our optimizer also manages to reduce a local detour in a long path while random shortcut methods will mostly fail.
+
+In a future work, we plan to improve the computation speed of our method.
 

### 2015-10-26 paper.tex
Add Florent and JP corrections and new convergence figure

@@
 
-\jvol{30} \jnum{15} \jyear{2016} \jmonth{August}
-
-\title{{\itshape Advanced Robotics} A simple path optimization method for motion planning }
-
-\author{Myl\`{e}ne Campana$^{\ast}$ \thanks{$^\ast$Corresponding author. Email: mcampana@laas.fr}, Florent Lamiraux and Jean-Paul Laumond
+\title{A simple path optimization method for motion planning$^{1}$ \thanks{$^{1}$Note for the reviewers: This paper is an extended version of a paper submitted to IEEE ICRA 2016~\cite{pathOptimICRA_HAL}.
+The current version includes a problem formulation, a convergence analysis and many experimental results (including industrial examples) which are absent from the conference version.}}
+
+\author{Myl\`{e}ne Campana$^{\ast}$ \thanks{$^\ast$Corresponding author. Email: \href{mailto:mcampana@laas.fr}{mcampana@laas.fr}}, Florent Lamiraux and Jean-Paul Laumond
 \\\vspace{6pt}
@@
 \section{Introduction}
-
 Motion planning for systems in cluttered environments has been addressed for more
-than thirty years~\citep{ref-motionplan},~\citep{ref-pianomover}. Most planners
-randomly sample the system configuration space~\cite{ref-spaceconfig} in
-order to find a collision-free path, linking the initial and final configurations. 
+than thirty years~\cite{ref-motionplan}. 
+To explore the connected components of collision-free configuration spaces, pionnering contributions in the 90’s introduced certain levels of random searches, i.e.
+random walks in~\cite{potentielBarraquandLatombe}, random sampling in
+~\cite{KavrakiLatombePRM} and~\cite{LaValleKuffnerRRT}. Today most motion planners are inspired by these seminal approaches.
 The main issue using these techniques is that the computed path makes unnecessary 
@@
 optimal solution, but computation time is significantly higher than classical
-approaches. Moreover the roadmap is only valid for a given motion planning
-   problem. 
+approaches.
+%Moreover the roadmap is only valid for a given motion planning problem. 
 \end{itemize}
@@
 However, RS requires fine 
-tuning of the termination condition (see 
-(Algorithm~\ref{algo:random-shortcut})) and is no efficient due to randomness for long 
-trajectories where only a small part needs to be optimized. 
-(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS will 
-always fail to optimize the initial path, since it cannot decouple the 
-robot degrees of freedom (DOF) on which the optimization occurs, contrary to our 
-algorithm or partial random shortcut (PRS) methods~\cite{Geraerts04clearancebased} 
-that shortcut only one or a group of DOF.
-
-\begin{figure}[H]
+tuning of the termination condition and is no efficient for long 
+trajectories where only a small part needs to be optimized, 
+as in the (Figure~\ref{local_box_optim}). 
+The (Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS 
+will always fail to optimize the initial path, since it cannot decouple the 
+robot degrees of freedom (DOF) on which the optimization occurs. This problem is 
+addressed by our method.
+
+\begin{figure}
 	\centering
 	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
-	\caption{Example of a path which random shortcut will never manage to 
+	\caption{Example of a path in $\real^3$ which random shortcut will never manage to 
 	optimize: each shortcut tentative will provide a collision or will not 
@@
 On the other hand, numerical
-optimization methods like CHOMP~\cite{RatliffChomp} can be used as a
+optimization methods like CHOMP~\cite{chompIjrr} can be used as a
 post-processing step. They have clear termination conditions, but collision
@@
 robot (and/or environment) model in order to make 
-it simpler:~\cite{RatliffChomp} covers PR2 bodies with spheres, 
+it simpler:~\cite{chompIjrr} covers PR2 bodies with spheres, 
 while~\cite{convexOptimMotplan} needs to decompose objects into convex subsets.
-Moreover, in the applications we address, optimality is not desirable as
-such since the shortest path between two configurations in a cluttered
-environment usually contains contacts with the obstacle,
-or is most of the time
-out of the scope of numerical optimization algorithms~\cite{LaumondOptim}.
+
+Finally it should be noticed that optimality in robot motion is a notion that should be clarified.
+Most of the time motion planners provide an optimized motion, which is not optimal at all, but is the output of a given optimization method.
+When optimal motions exist, numerical
+algorithms mostly fail in accounting for their combinatorial structure.
+Moreover, optimization algorithms bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}.
+In that perspective, a path optimization algorithm has to be evaluated with respect to other existing optimization techniques, from qualitative properties and from computational performance.
 
@@
 
-This paper is based on a similar work sent to a conference, where a different type of collision constraints was studied.
-
 \section{RELATED WORK}
-Previous work on path optimization has been conducted. CHOMP algorithm~\cite{RatliffChomp} optimizes an initial guess provided as
+Previous work on path optimization has been conducted. CHOMP algorithm~\cite{chompIjrr} optimizes an initial guess provided as
 input. It minimizes a time invariant cost function using efficient covariant
@@
 are time-optimal since they comply with acceleration and velocity constraints. 
-~\cite{Guernane2011}, also based on random shooting, is guiding configuration-
-generation with local holonimic considerations. But this method remains only 
+~\cite{Guernane2011}, also based on random shooting, is guiding 
+configuration-generation with local holonomic considerations. 
+But this method remains only 
 locally optimal, and is not addressing high-DOF problems.
@@
 iteration is detected in collision, instead of trying to constantly satisfy 
-distance constraints. Unlike our method however, anti-collision constraints are 
+distance constraints. Unlike our method however, collision constraints are 
 handled by interpolating configurations which, at some points of the trajectory, 
@@
 of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
-in the frame of its parent. In the examples shown in this paper, we consider 4 
+in the frame of its parent. In the examples shown in this paper, we consider four 
 types of joints described in (Table~\ref{tab:joints}).
-\begin{table}[b]
+\begin{table}
 \centerline {
@@
 }
-\caption {Translation and rotation joint position are defined by 1 
+\caption {Translation and rotation joint position are defined by one 
 parameter corresponding respectively to the translation along an axis and a 
 rotation angle around an axis. Unbounded rotation is defined by a point on the 
-unit circle of the plane: 2 parameters corresponding to the cosine and the 
+unit circle of the plane: two parameters corresponding to the cosine and the 
 sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
@@
 
-\paragraph {Operations on configurations and vectors} by analogy with the case 
+\subsubsection{Operations on configurations and vectors}%
+By analogy with the case 
 where the configuration space is a vector space, we define the following 
@@
 
-\subsubsection {Cost}
+\subsubsection{Cost}
 
@@
 where $I_{p_i}$ is the identity matrix of size $p_i$ and $w_i$ is the weight 
-associated to the lever arm of the joint $J_i$. We define the length of the straight interpolation 
+associated to the joint $J_i$. We define the length of the straight interpolation 
 between two configurations as:
@@
 For rotations, the weight is equal to the maximal distance of the robot bodies 
-moved by the joint to the center of the joint.
+moved by the joint to the center of the joint, in a given configuration.
 
@@
 $$
-where the $\lambda_{i-1}$ coefficients benefit will be explained in the results 
-section. For now we can assume that $\forall i,\, \lambda_{i-1}=1$.
+where the $\lambda_{i-1}$ coefficients will be explained in the results 
+section. For now it is assumed that $(\lambda_{i-1})_{1\leq i\leq wp+1}=1$.
 
@@
 
+The gradient of the cost function $\nabla \cost (\xx)$ is computed as follows:
+$$
+\nabla \cost (\xx) = 
+\left( (\lambda_{i}(\conf_{i+1} - \conf_{i})^T - \lambda_{i+1}(\conf_{i+2} - \conf_{i+1})^T) \weight^2 \right)_{i=0\cdots wp-1}
+$$
+
+Let $\mbox{H}$ denote the constant Hessian of the cost function:
+$$
+\mbox{H} = \left(\begin{array}{cccccc}
+(\lambda_{0}+\lambda_{1})\weight^2 & -\lambda_{1}\weight^2 & 0 & \cdots & & 0 \\
+-\lambda_{1}\weight^2 & (\lambda_{1}+\lambda_{2})\weight^2 & -\lambda_{2}\weight^2 & 0 & \cdots & 0 \\
+0 & -\lambda_{2}\weight^2 &  (\lambda_{2}+\lambda_{3})\weight^2 & -\lambda_{3}\weight^2 & 0 & \vdots \\
+\vdots & \ddots & \ddots & \ddots & \ddots & \vdots\\
+\vdots & & \ddots & \ddots & \ddots & \vdots\\
+0 & \cdots  & 0 & -\lambda_{wp-2}\weight^2 & (\lambda_{wp-2}+\lambda_{wp-1})\weight^2 & -\lambda_{wp-1}\weight^2 \\
+0 & \cdots &  \cdots & 0 & -\lambda_{wp-1}\weight^2 & (\lambda_{wp-1}+\lambda_{wp})\weight^2  \\
+\end{array}\right)
+$$
+
 \subsection {Resolution}
-We assume that the direct interpolation between the initial and final configurations contains collisions.
-Let $H$ denotes the constant Hessian of the cost function, an iteration is 
-described as follow:
+We assume that the direct interpolation between the initial and final configurations contains collisions. An iteration is described as follow:
 \begin{equation}\label{eq:iteration-1}
 \begin{split}
-& \pii =  -H^{-1} \nabla c(\xx_i)^{T} \\
+& \pii =  -H^{-1} \nabla \cost(\xx_i)^{T} \\
 & \xx_{i+1} =  \xx_{i} + \alpha_i \pii
@@
 unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
-between $q_0$ and $q_{n+1}$. Since this solution is in collision, we set 
+between $q_0$ and $q_{wp+1}$. Since this solution is in collision, we set 
 $\alpha_i = \alpha$ 
@@
 
-\begin{figure}[t]
+\begin{figure}
 	\centering
@@
 
-\subsection*{New constraint}
-
-Let $T$ be a positive real number such that each path $\xx_i$ is a mapping from 
-interval $[0,T]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(T) = q_{n+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
+\subsubsection{New constraint}
+
+Each path $\xx_i$ is a mapping from 
+interval $[0,1]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(1) = q_{wp+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
 us denote by $\tcolli$ the abscissa of the first collision detected on path 
@@
 \end{enumerate}
-In the rest of this section, we will only consider the first case. Reasoning 
+In the rest of this section, the first case will only be considered. Reasoning 
 about the second case is similar, except that the 
@@
 
-The principle of the method is to compute a constraint on the backtracked free-
-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
+The principle of the method is to compute a constraint on the backtracked 
+free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
 (\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
@@
 In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
-contact point expressed in the global frame ${\cal R}_0$ 
+contact point expressed in the global frame 
 (Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
@@
 \end {equation}
-As this constraint is non-linear, it is linearized around $\xx_{i}$. Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
-The constraint on the position at $t_{coll\,i}$ is a function that depends only
-on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq n$. There exists 
-$\beta\in[0,1]$ such that for any $j$:
-$$
-\xx_j (t_{coll\,i}) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
-$$
-The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
-blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
-and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
-(Algorithm~\ref{algo:gradient}).
-Then, a line is added in the contraint Jacobian matrix (function 
+As this constraint is non-linear, it is linearized around $\xx_{i}$. This step is described in a further section. Then, a line is added in the contraint Jacobian matrix (function 
 \texttt{addCollisionConstraint}), and the constraint 
@@
 
-Then we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
+Finally, we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
 programs (QP) (function \texttt{computeIterate} in 
@@
 
-\subsection*{Convergence Analysis}
-For iteratively constrained problem, the risk of falling in an infinite loop 
+\vspace{0.2cm}
+
+\subsubsection{Linearized constraint computation} \label{sec:lin_constr_compt}
+
+Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
+The constraint on the position at $t_{coll\,i}$ is a function that depends only
+on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq wp$.
+There exists $\beta\in[0,1]$ such that for any $j$:
+$$
+\xx_j (\tcolli) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
+$$
+The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
+blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
+and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
+(Algorithm~\ref{algo:gradient}).
+
+\subsection{Convergence Analysis}
+For iterative algorithms, the risk of falling in an infinite loop 
 that keeps adding a redundant constraint exists.
-In this section, we demonstrate that, given the constraint definition and the 
-function \texttt{solveRedundantConstraint}, our 
-algorithm will converge.
-Under the previous notations, it is straight-forward that
+In this section, we give some asumptions under which our 
+algorithm will converge, given the constraint definition and the 
+function \texttt{solveRedundantConstraint}.
+From the definition of $\Jf_{i}$, it is straightforward that:
 \begin{equation}\label{eq:decreasing-kernel}
@@
 $$
-This results proves that each added constraint is linearly independant from the 
+This result proves that under Asumption~(\ref{eq:asumption-1}), each added constraint 
+is linearly independent from the 
 previous ones, and thus our algorithm terminates in a finite number of iterations.
 
-\subsubsection*{About asumption~(\ref{eq:asumption-1})}%
+\vspace{0.2cm}
+
+\subsubsection{Note: about Asumption~(\ref{eq:asumption-1})}%
 Let us consider the following expression:
@@
 path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
-reached for $t=1$.
+reached when $t=1$.
 This expression being equal to zero requires either
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \begin{algorithmic}%[1] % line numerotation
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
+	\caption{Representation of the trajectory $\conf (t)$ from 
+	Equation~(\ref{eq:trajectory}) in the 
+	configuration space (left). For the same trajectory in the workspace, the 
+	colliding bodies are illustrated. Note that a case with 
+	$\dot{\po_2}(\xx_{Free}).\mathbf{u}=0$ is considered. Therefore, 
+	(Algorithm~\ref{algo:solveRedundant}) will produce one intermediate step
+	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
+	$\dot{\po_2}(\xx).\mathbf{u}\neq 0$ so the redundancy is solved and the
+	constraint will be linearized on $\xx$.}
+	\label{fig:convergence-diagram}
+\end{figure}
+
 Doing so, either the new constraint makes the dimension of the kernel decrease, or 
@@
 case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
-closer and closer. And the continuity of $f$ implies that
+closer and closer (see (Figure~\ref{fig:convergence-diagram}) as an example). And 
+the continuity of $f$ implies that:
 $$
@@
 contact. We already handled this case previously, admitting that the velocity norm 
-would not vanish. Therefore, Equation~(\ref{eq:asumption-1}) is verified.
+would not vanish at $t=1$. Therefore, Asumption~(\ref{eq:asumption-1}) is verified.
 
@@
 In this part we describe the GB path-optimizer (Algorithm~\ref{algo:gradient})
-according to the previous step definitions. The main difficulty here is to handle 
+according to the previous steps definitions. One difficulty here is to handle 
 the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
@@
 smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found with a small step (i.e. $\alpha \neq 1$).
+Note that even if the constraints are linearized, the algorithm converges.
 
@@
 Humanoid Path Planner~\cite{hpp}. Initial paths are obtained with two 
-kind of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
+kinds of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
 RRT-connect~\cite{rrt-connect}.
@@
 	\caption{Path-optimization results on 2D-ponctual robots, moving around 
-	green obstacles. Initial paths are dashed and crosses represent contact points 
+	gray obstacles. Initial paths are dashed and crosses represent contact points 
 	$\po_c$ related to collision-constraints. Note that, on the left, the detour 
@@
 compute the corresponding collisions, constraints will be chosen on a very-not 
-optimal path and will not be pertinent w.r.t. the obstacles we wanted to 
+optimal path and will not be relevant w.r.t. the obstacles we wanted to 
 avoid.
@@
 \lambda_{i-1} = \frac{1}{\sqrt{(\conf_{i\,0}-\conf_{i-1\,0})^T \weight^2 
-(\conf_{i\,0}-\conf_{i-1\,0})}}
+(\conf_{i\,0}-\conf_{i-1\,0})}}, \text{where} \  1\leq i\leq wp+1 
 $$
@@
 
-\begin{figure}[t]
+\begin{figure}
 	\centering
-	\includegraphics[width=10cm]{local_box_optim.pdf} % MOVE ROBOT TO THE ZOOM !
+	\includegraphics[width=10cm]{local_box_optim.pdf}
 	\caption{Case of a long initial path (above) containing a small part 
@@
 
-\begin{figure}[b]
+\begin{figure}
 	\centering
@@
 % freeflyer puzzle
-Before entering the manipulators examples, the GB algorithm is tested and compared 
-on a problem very encountered in the motion planning literature: a freeflyer puzzle 
+Before entering the manipulator examples, the GB algorithm is tested and compared 
+on a problem very popular in the motion planning literature: a freeflyer puzzle 
 example, corresponding to (Figure~\ref{fig:trajectories:puzzle}).  The puzzle has to cross down the 
@@
 Some problems are 
-involving an a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
+involving a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
 gripper.
@@
 % baxter:
-An problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
+A problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
 manipulating in an office environment is shown 
@@
 On the three following high-DOF examples involving the PR2 robot, we argue that 
-results are better in terms of path quality, in the fact that parasite DOF are 
-removed.
+results are better in terms of path quality, in the fact that parasite DOF motions 
+are removed.
 
@@
 to visualize the trajectory of the right gripper). The robot has to move 
-his hands from the top to the bottom of a table.
+its hands from the top to the bottom of a table.
 Our optimizer manages to reduce the initial length 12.7 from the visibility-PRM 
@@
 Another example of PR2 working in the kitchen, going from the set to the fridge 
-door, is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile 
-trajectories. This time, GB path length is higher that RS one, but the resulting 
-path remains acceptable.
+door, is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile base
+trajectories. Here, GB and RS results are similar in terms of length and rendering.
 
@@
 Optimization computation time and path length averages are presented for 30 runs of 
-each example in the (Table~\ref{tab:results}). In some cases $\alpha_{init}$ was 
+each example in (Table~\ref{tab:results}). In some cases $\alpha_{init}$ was 
 reduced to comply with very narrow passages, or increased in the opposite case. 
@@
 constraints of GB is typically important and $\alpha_{init}$ reduced.
+Moreover in this case, robot and environment geometries are simple so computing numerous collision tests for random shortcut is not handicaping.
 On the other hand, our method almost always provides the shortest paths, whereas it 
@@
 reduction or affordable computation time. This is particularly visible for PRS which 
-easily gives up in some problems, such as the 2D slidding-robot, the 
+easily gives up in some problems, such as the 2D sliding-robot, the 
 UR5-with-spheres or the puzzle.
@@
 
-\begin{figure}[ht]
+\begin{figure}
 	\centering
@@
 
-\begin{table}[ht]
+\begin{table}
 \centering
@@
     & Initial &  & $18.8$  &\\
-    \textbf{Freeflyer-} & GB ($\alpha_{init}=0.05$) & $1.02\,\text{s}$ & $9.65$  &\\
+    \textbf{Freeflyer-} & GB ($\alpha_{init}=0.05$) & $1020\,\text{ms}$ & $9.65$  &\\
     \textbf{puzzle} & PRS & $1.13\,\text{ms}$ & $18.8$  &\\
@@
   }%scalebox
-\caption{Average results for 30 runs of several examples presented in the paper or 
-the video. For each run, a solution path is planned by Visibility-PRM (unless `RRT' 
-for RRT-connect is mentioned) as initial guess for each optimizer.
+\caption{Average results for 30 runs of several examples presented in the paper and 
+in the video. For each run, a solution path is planned by Visibility-PRM (unless 
+`RRT' for RRT-connect is mentioned) as initial guess for each optimizer.
 For the PR2-crossing-arms example only, the right column represents the 
@@
 that the optimizer may be 
-time-competitive compared to random shortcut and proposes better quality paths, reducing the path length and removing unnecessary DOF motions.
+time-competitive compared to random shortcut in complex models where collision tests are handicaping, and proposes better quality paths, reducing the path length and removing unnecessary DOF motions.
 Our optimizer also manages to reduce a local detour in a long path while random shortcut methods will mostly fail.
 
-In a future work, we plan to improve the computation speed of our method.
+For future work, we have room for improvement.
 

### 2015-10-29 paper.tex
Add HAL reference, re-write abstract

@@
 \usepackage[numbers]{natbib}
-\algnewcommand\algorithmicinput{\textbf{input:}}
+\algnewcommand\algorithmicinput{\textbf{Input:}}
 \algnewcommand\INPUT{\item[\algorithmicinput]}
-\algnewcommand\algorithmicoutput{\textbf{output:}}
+\algnewcommand\algorithmicoutput{\textbf{Output:}}
 \algnewcommand\OUTPUT{\item[\algorithmicoutput]}
@@
 
-\title{A simple path optimization method for motion planning$^{1}$ \thanks{$^{1}$Note for the reviewers: This paper is an extended version of a paper submitted to IEEE ICRA 2016~\cite{pathOptimICRA_HAL}.
-The current version includes a problem formulation, a convergence analysis and many experimental results (including industrial examples) which are absent from the conference version.}}
+\title{A simple path optimization method for motion planning$^{1}$ \thanks{$^{1}$Note for the reviewers: 
+This paper is an extended version of a paper submitted to IEEE ICRA 2016~\cite{pathOptimICRA_HAL}. The 
+current version includes a problem formulation, a convergence analysis and many experimental results (including an 
+industrial example) which are absent from the conference version.}}
 
@@
 to the randomness of sampling, the paths make detours that need to be optimized.
-The contribution of this paper is to propose a gradient-based (GB) algorithm that
-transforms a polygonal collision-free path into a shorter one, while both:
-\begin{itemize}
-\item requiring only collision checking, and not any time-consuming obstacle distance computation nor geometry simplification,
-\item constraining only part of the configuration variables that may cause a 
-collision, and not the entire configurations, and
-\item reducing parasite motions that are not useful for the problem resolution without any assumption.
-\end{itemize}
-The algorithm is simple and requires few parameter tuning. Experimental
-results include navigation and manipulation tasks, e.g. an robotic arm 
-manipulating throw in boxes and a PR2 robot working in a kitchen environment. 
-Comparisons with a random shortcut optimizer and a partial shortcut are also provided.
+The contribution of this paper is to propose a basic gradient-based (GB) algorithm 
+that transforms a polygonal collision-free path into a shorter one.
+While requiring only collision 
+checking, and not any time-consuming obstacle distance computation nor geometry 
+simplification, we constrain only part of the configuration variables 
+that may cause a collision, and not entire configurations. As a result, parasite motions that are not useful for the problem resolution are reduced without any 
+assumption.
+Experimental results include navigation and manipulation tasks, 
+e.g. a manipulator arm filling boxes and a PR2 robot working in a kitchen 
+environment. Comparisons with a random shortcut optimizer and a partial 
+shortcut have also been studied.
+
+% few parameter tuning REMOVED
 
@@
 than thirty years~\cite{ref-motionplan}. 
-To explore the connected components of collision-free configuration spaces, pionnering contributions in the 90’s introduced certain levels of random searches, i.e.
-random walks in~\cite{potentielBarraquandLatombe}, random sampling in
+To explore the connected components of collision-free configuration spaces, pioneering 
+contributions in the 90’s introduced certain levels of random searches, i.e. random walks in~\cite{potentielBarraquandLatombe}, random sampling in 
 ~\cite{KavrakiLatombePRM} and~\cite{LaValleKuffnerRRT}. Today most motion planners are inspired by these seminal approaches.
@@
 approaches.
-%Moreover the roadmap is only valid for a given motion planning problem. 
 \end{itemize}
@@
 tuning of the termination condition and is no efficient for long 
-trajectories where only a small part needs to be optimized, 
-as in the (Figure~\ref{local_box_optim}). 
-The (Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS 
+trajectories where only a minor part needs to be optimized, 
+as in (Figure~\ref{local_box_optim}).
+(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS 
 will always fail to optimize the initial path, since it cannot decouple the 
@@
 algorithms mostly fail in accounting for their combinatorial structure.
-Moreover, optimization algorithms bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}.
+In addition, optimization algorithms bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}.
 In that perspective, a path optimization algorithm has to be evaluated with respect to other existing optimization techniques, from qualitative properties and from computational performance.
@@
 input. It minimizes a time invariant cost function using efficient covariant
-hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
+Hamiltonian gradient descent. The cost is quantified by non-smooth parts (with
 high velocities) and an obstacle avoidance term, provided by the distance to the 
@@
 random samples. But as for CHOMP, the obstacle cost term requires a voxel map to 
-perform its Euclidian Distance Transforms, and represents the robot bodies with 
+perform its Euclidean Distance Transforms, and represents the robot bodies with 
 overlapping spheres. Such technique provides lots of distance and penetration 
@@
 are time-optimal since they comply with acceleration and velocity constraints. 
-~\cite{Guernane2011}, also based on random shooting, is guiding 
+Also based on random shooting,~\cite{Guernane2011} is guiding 
 configuration-generation with local holonomic considerations. 
-But this method remains only 
+Nevertheless this method remains only 
 locally optimal, and is not addressing high-DOF problems.
-~\cite{Geraerts04clearancebased}, whereas just implemented for freeflyer robots, is 
-using medial axis retractation for clearance and a PRS (applied only on random DOF) 
-which can address the problem of (Figure~\ref{decoupled_DOF_optimization}) but is 
-relatively slower than RS. Moreover, PRS is not using the information returned by 
+~\cite{Geraerts04clearancebased} is 
+using medial axis retraction for clearance and a PRS (applied only on random DOF) 
+which can address the problem of (Figure~\ref{decoupled_DOF_optimization}). However, it is 
+relatively slower than RS and only implemented for freeflyer robots. 
+Furthermore, PRS is not using the information returned by 
 the checker of which links are colliding to guide which DOF or group of DOF are 
@@
 freeze the whole robot configuration instead of a relevant subpart. Thus, this 
-method cannot solve the (Figure~\ref{decoupled_DOF_optimization}) problem.
+method cannot solve (Figure~\ref{decoupled_DOF_optimization}) issue.
 
@@
 \subsubsection{Cost}
-
-let $W\in\real^{p\times p}$ be a diagonal matrix of weights:
+Let $W\in\real^{p\times p}$ be a diagonal matrix of weights:
 $$
@@
 Weights are used to homogenize translations and rotations in the velocity vector. 
-For rotations, the weight is equal to the maximal distance of the robot bodies 
+For a rotation, the weight is equal to the maximal distance of the robot bodies 
 moved by the joint to the center of the joint, in a given configuration.
@@
 collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
-paths and to return the first colliding configuration along a path. This step corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}).
+paths and to return the first colliding configuration along a path. This stage corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}).
 
@@
 
-Let us assume that at iteration $i$, $k$ constraints have been inserted before the current interation. These constraints are stored as lines of a Jacobian matrix
+Let us assume that at iteration $i$, $k$ constraints have been inserted before the current iteration. 
+These constraints are stored as lines of a Jacobian matrix as follows:
 $$
@@
 $$
-where iterate $\p_i$ are constrained to be in the kernel of $\Jf_{i}$
+where the step $\p_i$ is constrained to be in the kernel of $\Jf_{i}$ as follows:
 $$
@@
 $\xx_{i}(t_{coll\ i})$ is collision-free.
+Let $g$ be the function defined from $\CS$ to $\real$ by:
+$$
+g (\conf) = \left(M_{1}^{-1} (\conf) M_2 (\conf) \po_2 - \po_1 | \mathbf{u}\right)
+$$
+where $\po_1$ and $\po_2$ are computed in the configuration $\conf$. The function $g$ 
+represents the projection of $\overrightarrow{\po_1\po_2}$ on $\mathbf{u}$ which is calculated 
+from the last collision-free path at abscissa $\tcolli$.
+
 Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
 $$
-f (\xx) = \left(M_{1}^{-1} (\xx) M_2 (\xx) \po_2 - \po_1 | \mathbf{u}\right)
+f (\xx) = g(\xx (\tcolli))
 $$
@@
 \end {equation}
-As this constraint is non-linear, it is linearized around $\xx_{i}$. This step is described in a further section. Then, a line is added in the contraint Jacobian matrix (function 
+This constraint forces the configuration $\xx (\tcolli)$ to keep the same distance 
+between points $\po_1$ and $\po_2$ that $\xx_i (\tcolli)$. As this constraint is 
+non-linear, we linearize it around $\xx_{i}$. This step is described in a further 
+section. Then, a line is added in the constraint Jacobian matrix (function 
 \texttt{addCollisionConstraint}), and the constraint 
@@
 that keeps adding a redundant constraint exists.
-In this section, we give some asumptions under which our 
+In this section, we give some assumptions under which our 
 algorithm will converge, given the constraint definition and the 
@@
 Let us assume that:
-\begin{equation}\label{eq:asumption-1}
+\begin{equation}\label{eq:assumption-1}
 L_{k+1}\p_i \not= 0.
 \end{equation}
-We will elaborate later on this asumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
-$$
-\kernel J_i \not= \kernel J_{i+1}.
+We will elaborate later on this assumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
+$$
+\kernel J_i \not= \kernel J_{i+1}
 $$
@@
 $$
-\dim (\kernel J_{i+1}) < \dim (\kernel J_i).
-$$
-This result proves that under Asumption~(\ref{eq:asumption-1}), each added constraint 
+\dim (\kernel J_{i+1}) < \dim (\kernel J_i)
+$$
+This result proves that under assumption~(\ref{eq:assumption-1}), each added constraint 
 is linearly independent from the 
@@
 
-\subsubsection{Note: about Asumption~(\ref{eq:asumption-1})}%
-Let us consider the following expression:
-$$
-\frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
-$$
-This expression represents the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot defined by the following trajectory:
+\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
+\noindent
+Let us consider the following trajectory:
 \begin{equation}\label{eq:trajectory}
-\conf (t) = \xx_{i}(t_{coll\ i}) + t\, \alpha_i \p_{i}.
+\conf (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (t_{coll\ i}), \ t \in [0,1]
 \end{equation}
-The motion is defined by taking a constant abscissa $t_{coll\ i}$ and be moving the 
-path along iterate $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is 
-reached when $t=1$.
-This expression being equal to zero requires either
+where the motion is defined by taking a constant abscissa $\tcolli$ and by moving from the 
+path $\xx_i$ along the step $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is reached when $t=1$.
+
+Thus the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
+$$
+\frac{\partial \conf}{\partial t}(t) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_i + t\alpha_i\xx_{i+1})\p_i
+$$
+Therefore the following expressions
+% \frac{\partial \conf}{\partial t}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \text{and} \ \frac{\partial \conf}{\partial t}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
+$$
+\dot{\conf}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \ \text{and} \ \ \dot{\conf}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
+$$
+correspond to $\dot{\po_2}(\xx).\mathbf{u}$, respectively in configurations $\xx_{i}(t_{coll\ i})$ and $\xx_{i+1}(t_{coll\ i})$. The term $\dot{\conf}(1)$ being equal to zero requires either:
 \begin {enumerate}
@@
 except if the user builds such a case on purpose.
-The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$. In this 
-case, we do not directly insert the new constraint. Instead, we process intermediate 
+The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$.
+Then we do not directly insert the new constraint. Instead, we process intermediate 
 iterations until the redundancy vanishes, as explained in function 
-\texttt{solveRedundantConstraint}.
+\texttt{findNewConstraint} and illustrated in (Figure~\ref{fig:convergence-diagram}).
 
@@
 \State $solved \gets false$
-\State $minReached \gets false$
 \While{($\textbf{not}(solved)$)}
-	\State $\p = \texttt{computeIterate}()$
-	\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
+	%\State $\p = \texttt{computeIterate}()$
+	%\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
 	\State $\xx \gets \xx_{Free} + 0.5\alpha\,\p$
-	\If{(\texttt{validatePath}($\xx_1$))}
+	\If{(\texttt{validatePath}($\xx$))}
 		\State $\xx_{Free} \gets \xx$
@@
 \end{algorithmic}
-\caption{\texttt{solveRedundantConstraint}()} \label{algo:solveRedundant}
+\caption{Description of \texttt{findNewConstraint}() which returns a linearly 
+independent constraint w.r.t. previous constraints stacked in $\Jf$.} 
+\label{algo:solveRedundant}
 \end{algorithm}
@@
 case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
-closer and closer (see (Figure~\ref{fig:convergence-diagram}) as an example). And 
-the continuity of $f$ implies that:
-$$
-\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
-$$
+closer and closer. And the continuity of $f$ implies that $\dot{\conf}(0)$ and 
+$\dot{\conf}(1)$
 converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
 $\body_1$ along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
-contact. We already handled this case previously, admitting that the velocity norm 
-would not vanish at $t=1$. Therefore, Asumption~(\ref{eq:asumption-1}) is verified.
+contact. We already handled this situation previously, admitting that the velocity norm 
+would not vanish at $t=1$. Therefore, assumption~(\ref{eq:assumption-1}) is verified.
+%$$
+%\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
+%$$
 
@@
 
-In this part we describe the GB path-optimizer (Algorithm~\ref{algo:gradient})
+In this section, the GB path-optimizer (Algorithm~\ref{algo:gradient}) is described
 according to the previous steps definitions. One difficulty here is to handle 
@@
 			\State \texttt{computeCollisionConstraint}($\xx_1$, $\xx_0$)
-			\State \texttt{solveRedundantConstraint}()
+			\State \texttt{findNewConstraint}()
 			\State \texttt{addCollisionConstraint}()
@@
 kinds of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
-RRT-connect~\cite{rrt-connect}.
+RRT-connect~\cite{rrt-connect}. Unless another value is given, $\alpha_{init}$ 
+is set to 0.2.
 
@@
 	\includegraphics[width=8cm]{contact_points2potential.pdf}
-	\caption{Path-optimization results on 2D-ponctual robots, moving around 
+	\caption{Path-optimization results on 2D robots, moving around 
 	gray obstacles. Initial paths are dashed and crosses represent contact points 
@@
 
-(Figure~\ref{2D_long}) shows the result of our optimizer on 2D cases.
-Represented collisions leading to constraints permit to understand how the path is kept out of the obstacles.
-Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle the constraint must be created, as at some parts the robot is very close to obstacles.
+(Figure~\ref{2D_long}) shows the result of our optimizer on 2D cases. Contact points which have 
+led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
+Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
+the constraint must be created, as at some parts the robot is very close to obstacles.
 
@@
 example, if we alter a lot the initial path with a large gradient step and 
-compute the corresponding collisions, constraints will be chosen on a very-not 
-optimal path and will not be relevant w.r.t. the obstacles we wanted to 
-avoid.
+compute the corresponding collisions, constraints will be chosen on a 
+path closer-to-optimality but which is not relevant
+w.r.t. the obstacles we want to avoid.
 
@@
 optimize in an affordable time, because of probabilistically failing to sample 
-configurations in the box. GB succeeds to optimize the 
+configurations in the box. The GB method succeeds to optimize the 
 path contained in the box, with the following cost coefficients:
@@
 straight line with the waypoints equidistantly allocated, which is not adapted for 
-the (Figure~\ref{local_box_optim}) type of problems with a local passage very
+(Figure~\ref{local_box_optim}) type of problems with a local passage very
 constrained by obstacles. Note that this cost is also working with all other 
@@
 
-We also experiment our algorithm on more complex robots and environments. Unless another value is given, $\alpha_{init}$ is set to 0.2.
+We also experiment our algorithm on more complex robots and environments.
 
 \begin{algorithm}
-\begin{algorithmic}[1] % line numerotation
-\Procedure{randomShortcut}{$\xx$}
+\begin{algorithmic}%[1] % line numerotation
+\INPUT{Path to optimize $\xx$}
+\OUTPUT{Optimized collision-free path $\xx$}
    \State nbFailures $\gets 0$
@@
     \Return $\xx$
-\EndProcedure
 \end{algorithmic}
@@
 explicit indication, any probabilistic motion planner will compute a path that 
-makes the PR2 mobile base purposelessly moving. This is the case of RRT-connect 
+makes the PR2 mobile base purposelessly moving. This is the result of RRT-connect 
 algorithm (left). Path optimization is expected to remove unnecessary motions. 
@@
 is tested and compared to RS~\cite{randomShortcutHPP}
-(Algorithm~\ref{algo:random-shortcut}) and PRS ~\cite{partialrandomShortcutHPP} 
+(Algorithm~\ref{algo:random-shortcut}) and PRS~\cite{partialrandomShortcutHPP} 
 adapted from~\cite{Geraerts04clearancebased}.
@@
 $\conf_0$ and $\conf_{wp+1}$. 
-Then, for each DOF, configurations are shot as in 
-(Algorithm~\ref{algo:random-shortcut}) and the steering-method returns a path made by 
+Then, for each DOF, configurations are sampled as in 
+(Algorithm~\ref{algo:random-shortcut}). The steering-method returns a path made of 
 an interpolation on only 
-the current DOF and the previous subpath for other DOF. If this path is 
-collision-free, it is replaced as in RS. This process is also stopped when five 
+the current DOF and based on the previous subpath for other DOF. If this path is 
+collision-free, we replace it as in RS. This process is also stopped when five 
 iterations of non-improvement are reached.
@@
 respectively 23.4, 8.53 and 10.2. PRS fails to decrease the path length because in 
-this problem, all DOF are closely related to cross the hole. This behaviour was also 
+this problem, all DOF are closely related to cross the hole. This behavior was also 
 noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, 
@@
 other arm is relatively far from obstacles. As expected, the initial path given by 
-RRT-connect activates both arms to solve the problem. Contrary to RS, 
-GB manages to cancel the rotation of the free-arm while optimizing the 
+RRT-connect activates both arms to solve the problem. Unlike RS, 
+the GB optimizer manages to cancel the rotation of the free-arm while optimizing the 
 first arm motion, even if the second arm is still moving because of the translation 
 motion. The initial path length 5.17 is decreased to 4.55 by RS, to 4.06 by PRS and 
-to 2.77 by GB using three collision-constraints with the cylinder obstacle.
+to 2.77 by the GB method using three collision-constraints with the cylinder obstacle.
 
@@
 	are presented (bottom right): initial (RRT-connect) path in red, RS 
-	optimization in blue and GB optimization in green.}
+	optimization in blue and the GB optimization in green.}
 	\label{fiad}
@@
 In a relatively free environment (Figure~\ref{fig:trajectories:ur5}), 
-results from GB and RS are similar (path length are respectively 2.03 and 1.60). 
-Note also that the end-effector trajectory has completely changed from the initial 
-one, the robot is easily passing between the spheres, keeping its end-effector 
+results from our method and RS are similar (path length are respectively 2.03 and 1.60). 
+Note also that the end-effector trajectory is completely different from the initial 
+one: the robot is easily passing between the spheres, keeping its end-effector 
 above.
@@
 % UR5-fiad
-For an UR5 working in a cluttered environment inspired from an industrial issue 
+For an UR5 working in a cluttered environment inspired by an industrial issue 
 (Figure~\ref{fiad}), path optimization efficiently returns a shorter solution, close 
@@
 The initial path length 6.65 is decreased to 2.08 by RS, to 
-1.98 by GB and unchanged by PRS.
+1.98 by the GB optimizer and unchanged by PRS.
 
@@
 A problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
-manipulating in an office environment is shown 
+manipulating in an office environment is presentated 
 (Figure~\ref{fig:trajectories:baxter}). The robot starts with its end-effectors 
@@
 video for the full motion) and path lengths, 9.10 (initial RRT-connect), 7.52 (RS), 
-8.41 (PRS) and 2.70 (GB), the GB optimizer is the most successfull in this case.
+8.41 (PRS) and 2.70 (GB), the GB optimizer is the most successful in this case.
 
@@
 result, only presented in the video, cancels the mobile base motion, but not all 
-unnecessary DOF. Moreover, the arms motion is very wide and not as optimized as in 
-the GB trajectory.
+unnecessary DOF. In addition, the arms motion is very wide and not as optimized as in the GB trajectory.
 
@@
 backward and uses its arms DOF to avoid the table, instead of 
-processing a large motion to get away from the table. With GB, the path length is 
+processing a large motion to get away from the table. With the GB algorithm, the path length is 
 downed to 3.15, against 5.47 for RS.
@@
 reduced to comply with very narrow passages, or increased in the opposite case. 
-It appears that, for non-high DOF examples, computation time is higher using the GB 
-optimization. The worst case shows up for the puzzle problem where the number of 
-constraints of GB is typically important and $\alpha_{init}$ reduced.
-Moreover in this case, robot and environment geometries are simple so computing numerous collision tests for random shortcut is not handicaping.
+On the one hand, for non-high DOF examples, it appears that computation time is higher using the GB 
+optimization. The worst case comes out for the puzzle problem where the number of 
+constraints of GB is significant and $\alpha_{init}$ reduced. 
+In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimization spends time on the QP resolution to avoid collisions.
 On the other hand, our method almost always provides the shortest paths, whereas it 
@@
 optimized paths is highlighted in the right column of (Table~\ref{tab:results}). 
-Thus the problem addressed in the (Figure~\ref{decoupled_DOF_optimization}) is 
+Thus the problem addressed in (Figure~\ref{decoupled_DOF_optimization}) is 
 solved considering this DOF.
@@
 	}
-  \caption{Trajectories of end-effectors, mobile bases or centers allong the 
+  \caption{Trajectories of end-effectors, mobile bases or centers along the 
   different paths: the initial path is represented in red, the RS optimized path in 
@@
   motions can also be visualized in the joined video. The trajectories comparison 
-  highlights the optimization sucess of our method, which manages to deliver a 
+  highlights the optimization success of our method, which manages to deliver a 
   shorter or similar path compared to RS output}
@@
 \section{CONCLUSIONS}
-We managed to settle a path optimization for navigation and manipulation problems, and tested with various robots and environments. Our algorithm uses standard 
+We managed to settle a path optimization for navigation and manipulation problems, and tested it with various robots and environments. Our algorithm uses standard 
 numerical tools as collision checking, linearized one-dimensional constraint
 and QP resolution. It correlates them in a 
-simple but effective way, playing on the scalar iteration parameter. Moreover, 
+simple but effective way, and the algorithm structure is organized so that the convergence is guaranteed. Furthermore, 
 our method only requires collision checking, therefore geometry pre-processing or 
@@
 that the optimizer may be 
-time-competitive compared to random shortcut in complex models where collision tests are handicaping, and proposes better quality paths, reducing the path length and removing unnecessary DOF motions.
-Our optimizer also manages to reduce a local detour in a long path while random shortcut methods will mostly fail.
-
-For future work, we have room for improvement.
+time-competitive compared to random shortcut in complex models where collision tests are time-consuming. 
+It also proposes better quality paths, reducing the path length and removing unnecessary DOF motions. 
+Finally, our optimizer manages to reduce a local detour in a long path while random shortcut methods will mostly fail.
+
+For future work, we have room for improvement: taking advantage of the sparsity of 
+the constraint Jacobian to reduce computation time. We may also adapt the iteration 
+scalar parameter from geometries considerations on the current path, e.g. a lower 
+bound of the distance between nearest objects, since this bound can be returned by 
+the collision checker.
 

### 2015-10-29 paper.tex
A few corrections.

@@
 
-Let $\mbox{H}$ denote the constant Hessian of the cost function:
+From the gradient expression, we notice that the Hessian $\mbox{H}$ is constant:
 $$
@@
 contact point expressed in the global frame 
-(Figure~\ref{contact-points} left). We compute the respective positions $\po_1$ and 
-$\po_2$ of $\po_c$ in 
-the local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
+(Figure~\ref{contact-points} left). We define $\po_1$ and $\po_2$ as the coordinates of $\po_c$ in the  respective local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
 
@@
 $$
-where $\po_1$ and $\po_2$ are computed in the configuration $\conf$. The function $g$ 
+The function $g$
 represents the projection of $\overrightarrow{\po_1\po_2}$ on $\mathbf{u}$ which is calculated 
@@
 $$
-\frac{\partial \conf}{\partial t}(t) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_i + t\alpha_i\xx_{i+1})\p_i
+ \dot{\conf}(t) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_i + t\alpha_i\xx_{i+1})\p_i
 $$
@@
 	$\dot{\po_2}(\xx).\mathbf{u}\neq 0$ so the redundancy is solved and the
-	constraint will be linearized on $\xx$.}
+	constraint will be linearized around $\xx$.}
 	\label{fig:convergence-diagram}
@@
 be traveled through. Typically, we 
-chose $\alpha_{init} = 0.2$ or 0.1 to process small steps. $\alpha_{init} = 0.5$ is 
-more encountered in the optimization literature, procuring larger steps but with 
-more risks of collision.
+chose $\alpha_{init} = 0.2$ or 0.1 to process small steps.
 

### 2015-10-29 paper.tex
Add Justin's feedback, fix typo

@@
 
-The velocity of each joint $J_i$, $1\leq i \leq m$ is defined by a vector of $
-\real^{p_i}$, where $p_i$ is the number of DOF of $J_i$. Note that 
+The velocity of each joint $J_i$, $1\leq i \leq m$,  belongs to the tangent space of 
+the joint configuration space, and is defined by a vector of $\real^{p_i}$, where 
+$p_i$ is the number of DOF of $J_i$. Note that
 the velocity vector does not necessarily have the same dimension as the 
@@
 \begin{split}
-& \pii =  -H^{-1} \nabla \cost(\xx_i)^{T} \\
-& \xx_{i+1} =  \xx_{i} + \alpha_i \pii
+& \p_i =  -H^{-1} \nabla \cost(\xx_i)^{T} \\
+& \xx_{i+1} =  \xx_{i} + \alpha_i \p_i
 \end{split} 
@@
 unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
-between $q_0$ and $q_{wp+1}$. Since this solution is in collision, we set 
-$\alpha_i = \alpha$ 
-where $\alpha$ is a parameter that will be explained in the algorithm section.
+between $\conf_0$ and $\conf_{wp+1}$. Since this solution is in collision, we set 
+$\alpha_i = \alpha_{init}$
+where $\alpha_{init}$ is a parameter in interval $[0,1]$.
 
@@
 the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
-be traveled through. Typically, we 
-chose $\alpha_{init} = 0.2$ or 0.1 to process small steps.
+be traveled along. We choose $\alpha_{init} = 0.2$ or 0.1 to process small steps.
 

### 2015-10-30 paper.tex
Fix first page footnote placement, Hessian typo
final corrections

@@
 simplification, we constrain only part of the configuration variables 
-that may cause a collision, and not entire configurations. As a result, parasite motions that are not useful for the problem resolution are reduced without any 
+that may cause a collision, and not entire configurations. Thus parasite motions that 
+are not useful for the problem resolution are reduced without any 
 assumption.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
-	\caption{Example of a path in $\real^3$ which random shortcut will never manage to 
-	optimize: each shortcut tentative will provide a collision or will not 
-	decrease the path length. The optimal path belongs to the $x-y$ plane 
-	containing $\conf_{init}$ and $\conf_{final}$.}
-	\label{decoupled_DOF_optimization}
-\end{figure}
-
 On the other hand, numerical
@@
 
-Finally it should be noticed that optimality in robot motion is a notion that should be clarified.
-Most of the time motion planners provide an optimized motion, which is not optimal at all, but is the output of a given optimization method.
-When optimal motions exist, numerical
-algorithms mostly fail in accounting for their combinatorial structure.
-In addition, optimization algorithms bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}.
-In that perspective, a path optimization algorithm has to be evaluated with respect to other existing optimization techniques, from qualitative properties and from computational performance.
+\begin{figure}
+	\centering
+	\includegraphics[width=7.5cm]{decoupled_DOF_optimization.pdf}
+	\caption{Example of a path in $\real^3$ which random shortcut will never manage to 
+	optimize: each shortcut attempt will provide a collision or will not 
+	decrease the path length. The optimal path belongs to the $x-y$ plane 
+	containing $\conf_{init}$ and $\conf_{final}$.}
+	\label{decoupled_DOF_optimization}
+\end{figure}
+
+Finally it should be noticed that optimality in robot motion is a notion that should be clarified. Most of the 
+time motion planners provide an optimized motion, which is not optimal at all, but is the output of a given 
+optimization method. When optimal motions exist, numerical
+algorithms mostly fail in accounting for their combinatorial structure. In addition, optimization algorithms 
+bypass (not overcome) the question of the existence of optimal motions~\cite{LaumondOptim}. In that perspective, 
+a path optimization algorithm has to be evaluated with respect to other existing optimization techniques, 
+from qualitative properties and from computational performance.
 
 The idea of our method is to find a good trade-off between
-the simplicity of “blind” methods like shortcut algorithm,
+the simplicity of “blind” methods like shortcut algorithms,
 and the complexity of distance based optimization techniques.
@@
 between the objects detected in collision. Only collisions between objects are 
-evaluated, therefore no pre-processing of the
+evaluated, therefore no pre-processing of either the
 robot or environment models is necessary to increase distance computation speed. 
@@
 straight-line manually or randomly-sampled initialization as 
-TrajOp~\cite{SchulmanConvexOptim}. Its trajectory is iteratively optimized with 
+TrajOp~\cite{SchulmanConvexOptim}. The path is iteratively optimized with 
 sequential convex optimization by minimizing at each step its square length, 
@@
 Some heuristics use random shortcuts on the initial guess combined with a 
-trajectory re-building that returns ${C}^1$ shortcuts made of parabolas and lines 
-(bang-bang control)~\cite{HauserFastSmooth}. These local refined trajectories 
+trajectory re-building.
+For example,~\cite{HauserFastSmooth} returns smooth shortcuts made of parabola and line combinations, relying on the classical bang-bang control approach.
+These local refined trajectories
 are time-optimal since they comply with acceleration and velocity constraints. 
@@
 Nevertheless this method remains only 
-locally optimal, and is not addressing high-DOF problems.
-~\cite{Geraerts04clearancebased} is 
-using medial axis retraction for clearance and a PRS (applied only on random DOF) 
-which can address the problem of (Figure~\ref{decoupled_DOF_optimization}). However, it is 
-relatively slower than RS and only implemented for freeflyer robots. 
-Furthermore, PRS is not using the information returned by 
-the checker of which links are colliding to guide which DOF or group of DOF are 
-relevant to shortcut, contrary to our method.
+locally optimal, and is not addressing high-DOF 
+problems.~\cite{Geraerts04clearancebased} is 
+using medial axis retraction for clearance and a PRS (shortcut is applied only 
+on some random DOF) 
+which can address the problem of (Figure~\ref{decoupled_DOF_optimization}). 
+However it is 
+relatively slower than RS, and~\cite{Geraerts04clearancebased} only investigates 
+it for freeflyer robots. 
+Furthermore PRS is not taking advantage of information returned by 
+the checker of which links are colliding, in order to guide the selection of a relevant group of DOF to shortcut.
 
@@
 handled by interpolating configurations which, at some points of the trajectory, 
-freeze the whole robot configuration instead of a relevant subpart. Thus, this 
+freeze the whole robot configuration instead of a pertinent subpart. Thus, this 
 method cannot solve (Figure~\ref{decoupled_DOF_optimization}) issue.
@@
 of rigid-body motions $SE(3)$. The rigid-body motion is the position of the joint 
-in the frame of its parent. In the examples shown in this paper, we consider four 
+in the frame of its parent. In the examples presented in this paper, we consider four 
 types of joints described in (Table~\ref{tab:joints}).
@@
 }
-\caption {Translation and rotation joint position are defined by one 
+\caption {Translation and rotation joints positions are defined by one 
 parameter corresponding respectively to the translation along an axis and a 
@@
 sine of the rotation angle. $SO(3)$ is defined by a unit quaternion. The 
-velocity of translation and unbounded rotation joints is the derivative of the 
+velocity of translation and bounded rotation joints is the derivative of the 
 configuration variable. The velocity of an unbounded rotation joint corresponds 
@@
 is defined by the concatenation of the joint configurations. The configuration 
-space of the robot is denoted by $\CS\subset\real^n$.
-
-Note that the configuration of the robot belongs to a sub-manifold of $\real^n$.
+space of the robot is denoted by $\CS\subset\real^n$. Note that the configuration of the robot belongs to a sub-manifold of $\real^n$.
 
@@
 The velocity of the robot is defined as the concatenation of the velocities of 
-each joint.
+each joint:
 $$\dot{\conf} = (\underbrace{\dot{q}_{1},\cdots,\dot{q}_{p_1}}_{J_1},
@@
 configuration space of the robot. The above sum corresponds to the exponential 
-map. We do not have enough space in this paper to develop the theory in a more 
-rigorous way. The reader can easily state that ``following a 
+map. The reader can easily state that ``following a 
 constant velocity'' makes sense for the four types of joints defined in 
@@
 
+Each path $\xx$ is a mapping from 
+interval $[0,1]$ into $\CS$: $\xx(0) = q_0$, $\xx(1) = q_{wp+1}$.
+Finally, we use a continuous 
+collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
+paths and to return the first colliding configuration along a path. This stage 
+corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}), which returns $true$ if the given path is collision-free.
+
 \subsubsection{Cost}
@@
 \|\conf_2 - \conf_1\|_{W} \triangleq \sqrt{(\conf_2 - \conf_1)^T W^2 (\conf_2 - 
-\conf_1)}.
+\conf_1)}
 $$
@@
 where the $\lambda_{i-1}$ coefficients will be explained in the results 
-section. For now it is assumed that $(\lambda_{i-1})_{1\leq i\leq wp+1}=1$.
-
-Note that $\cost$ is not exactly the length of the path, but it can be 
+section. For now it is assumed that $(\lambda_{i-1})_{1\leq i\leq wp+1}=1$. Note that 
+$\cost$ is not exactly the length of the path, but it can be 
 established that minimal length paths also minimize $\cost$. This latter cost is 
@@
 \subsection {Resolution}
-We assume that the direct interpolation between the initial and final configurations contains collisions. An iteration is described as follow:
+We assume that the direct interpolation between the initial and final configurations contains collisions. An iteration at stage $i$ is described as follow:
 \begin{equation}\label{eq:iteration-1}
 \begin{split}
-& \p_i =  -H^{-1} \nabla \cost(\xx_i)^{T} \\
+& \p_i =  -\mbox{H}^{-1} \nabla \cost(\xx_i)^{T} \\
 & \xx_{i+1} =  \xx_{i} + \alpha_i \p_i
@@
 When a collision is detected, we introduce a constraint and perform a new 
-iteration from $\xx_i$ as explained in the next section. We use a continuous 
-collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
-paths and to return the first colliding configuration along a path. This stage corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}).
+iteration from $\xx_i$ as explained in the next section.
 
@@
 	appears 
-	to be in collision with the obstacle, the first colliding configuration 
+	to be in collision with the obstacle. The first colliding configuration 
 	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
@@
 
-Each path $\xx_i$ is a mapping from 
-interval $[0,1]$ into $\CS$: $\xx_i(0) = q_0$, $\xx_i(1) = q_{wp+1}$ for all $i$. As illustrated in (Figure~\ref{optim_grad}), let 
+As illustrated in (Figure~\ref{optim_grad}), let 
 us denote by $\tcolli$ the abscissa of the first collision detected on path 
@@
 	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
-	Collision Library~\cite{fcl}, used to detect collisions. The red $(\perp)$ 
-	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
-	direction which is allowed by the constraint.}
+	Collision Library~\cite{fcl}, used to detect collisions. 
+%	The red $(\perp)$ 
+%	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
+%	direction which is allowed by the constraint.
+	}
 	\label{contact-points}
@@
 contact point expressed in the global frame 
-(Figure~\ref{contact-points} left). We define $\po_1$ and $\po_2$ as the coordinates of $\po_c$ in the  respective local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
-
-Let $M_1 \in SE(3)$ (resp. $M_2 \in SE(3)$) denote the homogeneous transformation 
+(Figure~\ref{contact-points} left).
+In any configuration $\conf$, we define $\po_1(\conf)$ and $\po_2(\conf)$ as the coordinates of $\po_c$ in the respective local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
+
+Let $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) denote the homogeneous transformation 
 between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
-$\xx_i(\tcolli)$. Let $\mathbf{u}$ be the unit vector 
-linking the contact points $\po_1$ and 
-$\po_2$ in configuration $\xx_{i}(t_{coll\ i})$, $\xx_{i}$ being the latest collision-free path:
-$$
-\mathbf{u} = \frac{M_{1}^{-1} (\xx_i) M_2 (\xx_i) \po_2 - \po_1}{\|M_{1}^{-1} 
-(\xx_i) M_2 (\xx_i) \po_2 - \po_1\|}
+$\conf$. Let $\mathbf{u}$ be the unit vector 
+linking the contact points $\po_1(\xx_{i}(\tcolli))$ and 
+$\po_2(\xx_{i}(\tcolli))$:
+$$
+\mathbf{u} = \frac{M_{1}^{-1} (\xx_{i}(\tcolli)) M_2 (\xx_{i}(\tcolli)) \po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))}{\|M_{1}^{-1} 
+(\xx_{i}(\tcolli)) M_2 (\xx_{i}(\tcolli)) \po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))\|}
 $$
@@
 $\xx_{i}(t_{coll\ i})$ is collision-free.
-Let $g$ be the function defined from $\CS$ to $\real$ by:
-$$
-g (\conf) = \left(M_{1}^{-1} (\conf) M_2 (\conf) \po_2 - \po_1 | \mathbf{u}\right)
+Let $g$ be the function defined for any $\conf \in \CS$ to $\real$ by:
+$$
+g (\conf) = \left(M_{1}^{-1} (\conf) M_2 (\conf) \po_2(\conf) - \po_1(\conf) | \mathbf{u}\right)
 $$
 The function $g$
-represents the projection of $\overrightarrow{\po_1\po_2}$ on $\mathbf{u}$ which is calculated 
-from the last collision-free path at abscissa $\tcolli$.
+represents the projection of $\overrightarrow{\po_1(\conf)\,\po_2(\conf)}$ 
+on $\mathbf{u}$.
+%which is calculated from the last collision-free path at abscissa $\tcolli$.
 
@@
 \end {equation}
-This constraint forces the configuration $\xx (\tcolli)$ to keep the same distance 
-between points $\po_1$ and $\po_2$ that $\xx_i (\tcolli)$. As this constraint is 
-non-linear, we linearize it around $\xx_{i}$. This step is described in a further 
+This constraint forces the configuration $\xx (\tcolli)$ to keep the same projection on $\mathbf{u}$ as in $\xx_i (\tcolli)$. Since this constraint is 
+non-linear, we linearize it around $\xx_{i}$: 
+$$
+\frac{\partial f}{\partial \xx}(\xx_i)(\xx - \xx_i) = 0
+$$
+The linearization step is described in a further 
 section. Then, a line is added in the constraint Jacobian matrix (function 
@@
 Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
-The constraint on the position at $t_{coll\,i}$ is a function that depends only
-on two waypoints $\conf_{k\,i}$, $\conf_{k+1\,i}$, with $0\leq k\leq wp$.
-There exists $\beta\in[0,1]$ such that for any $j$:
-$$
-\xx_j (\tcolli) = \conf_{k\,j} + \beta (\conf_{k+1\,j} - \conf_{k\,j})
-$$
-The path Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ is thus built by matrix 
-blocks using the previous constraint Jacobian $\frac{\partial f}{\partial \conf}$ 
-and $\beta$, corresponding to \texttt{computeCollisionConstraint} in 
+There exist $\beta\in[0,1]$ and $k$ such that $\xx_i (\tcolli)$ can be written as a combination of two waypoints:
+$$
+\xx_i (\tcolli) = \conf_{k\,i} + \beta (\conf_{k+1\,i} - \conf_{k\,i})
+$$
+
+Thus the linearized constraint Jacobian $\frac{\partial f}{\partial \xx}(\xx_i)$ 
+is built by matrix 
+blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in each of the two waypoints
+and $\beta$. This stage corresponds to \texttt{computeCollisionConstraint} in 
 (Algorithm~\ref{algo:gradient}).
@@
 In this section, we give some assumptions under which our 
-algorithm will converge, given the constraint definition and the 
-function \texttt{solveRedundantConstraint}.
+algorithm will converge.
 From the definition of $\Jf_{i}$, it is straightforward that:
@@
 \begin{equation}\label{eq:assumption-1}
-L_{k+1}\p_i \not= 0.
+L_{k+1}\p_i \not= 0
 \end{equation}
@@
 
-Thus the velocity of point $\po_2$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
+Thus the velocity of point $\po_2(q(t))$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
 $$
@@
 $$
+
 Therefore the following expressions
@@
 $$
-correspond to $\dot{\po_2}(\xx).\mathbf{u}$, respectively in configurations $\xx_{i}(t_{coll\ i})$ and $\xx_{i+1}(t_{coll\ i})$. The term $\dot{\conf}(1)$ being equal to zero requires either:
+correspond to $(\dot{\po_2}|\mathbf{u})$, respectively in configurations $\xx_{i}(t_{coll\ i})$ and $\xx_{i+1}(t_{coll\ i})$. The term $\dot{\conf}(1)$ being equal to zero requires either:
 \begin {enumerate}
@@
 iterations until the redundancy vanishes, as explained in function 
-\texttt{findNewConstraint} and illustrated in (Figure~\ref{fig:convergence-diagram}).
-
-\begin{algorithm}[t]
+\texttt{findNewConstraint} in (Algorithm~\ref{algo:solveRedundant}), and illustrated in (Figure~\ref{fig:convergence-diagram}).
+
+After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ has
+been halved, the obtained path is tested for collisions. If it is collision-free, the 
+constraint is now linearized around it. Otherwise, new contact points are taken 
+on it, and the constraint is linearized around the latest free path.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
+	\caption{Representation of a trajectory $\conf (t)$ from 
+	Equation~(\ref{eq:trajectory}) in the 
+	configuration space (left). For the same trajectory in the workspace, the 
+	colliding bodies are illustrated (right). Note that a case with 
+	$(\dot{\po_2}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
+	with the constraint that should be added if no redundancy was tracked. Therefore, 
+	(Algorithm~\ref{algo:solveRedundant}) will produce one intermediate step
+	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
+	$(\dot{\po_2}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
+	returning the new constraint.}
+	\label{fig:convergence-diagram}
+\end{figure}
+
+\begin{algorithm}[H]
 \begin{algorithmic}%[1] % line numerotation
 \INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
+\OUTPUT{linearized constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$}
 \Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
@@
 
-After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ was 
-halved, the obtained path is tested for collisions. If it is collision-free, the 
-constraint is now linearized around it. Otherwise, new contact points are taken 
-on it, and the constraint is linearized around the latest free path.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
-	\caption{Representation of the trajectory $\conf (t)$ from 
-	Equation~(\ref{eq:trajectory}) in the 
-	configuration space (left). For the same trajectory in the workspace, the 
-	colliding bodies are illustrated. Note that a case with 
-	$\dot{\po_2}(\xx_{Free}).\mathbf{u}=0$ is considered. Therefore, 
-	(Algorithm~\ref{algo:solveRedundant}) will produce one intermediate step
-	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
-	$\dot{\po_2}(\xx).\mathbf{u}\neq 0$ so the redundancy is solved and the
-	constraint will be linearized around $\xx$.}
-	\label{fig:convergence-diagram}
-\end{figure}
-
 Doing so, either the new constraint makes the dimension of the kernel decrease, or 
@@
 converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
-$\body_1$ along trajectory~(\ref{eq:trajectory}) when $\po_2$ and $\po_1$ come to 
+$\body_1$ along trajectory~(\ref{eq:trajectory}), when $\po_2$ and $\po_1$ come to 
 contact. We already handled this situation previously, admitting that the velocity norm 
@@
 
-In this section, the GB path-optimizer (Algorithm~\ref{algo:gradient}) is described
-according to the previous steps definitions. One difficulty here is to handle 
-the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
-be traveled along. We choose $\alpha_{init} = 0.2$ or 0.1 to process small steps.
-
-\begin{algorithm}
+\begin{algorithm}[H]
 \begin{algorithmic}%[1] % line numerotation
@@
 
+In this section, the GB path-optimizer (Algorithm~\ref{algo:gradient}) is described
+according to the previous steps definitions. One difficulty here is to handle 
+the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
+be traveled along. We choose $\alpha_{init} = 0.2$ or 0.05 to process small steps.
+
 Each time a collision-constraint is added, 
 the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
-smaller steps are iteratively applied and tested toward the minimum, and constraints are added each time a collision is found with a small step (i.e. $\alpha \neq 1$).
+smaller steps toward the minimum are iteratively applied and tested, and constraints are added each time a collision is found with a reduced step (i.e. $\alpha \neq 1$).
 Note that even if the constraints are linearized, the algorithm converges.
@@
 kinds of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
-RRT-connect~\cite{rrt-connect}. Unless another value is given, $\alpha_{init}$ 
+RRT-connect~\cite{rrt-connect}. Unless another value is provided, $\alpha_{init}$ 
 is set to 0.2.
@@
 Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
-the constraint must be created, as at some parts the robot is very close to obstacles.
+the constraint must be created, since at some parts the robot is very close to obstacles.
 
@@
 
-(Figure~\ref{local_box_optim}) illustrates a \textit{very long} path example that RS 
+(Figure~\ref{local_box_optim}) illustrates a \textit{very long} path example which RS 
 or PRS will not manage to 
@@
 \lambda_{i-1} = \frac{1}{\sqrt{(\conf_{i\,0}-\conf_{i-1\,0})^T \weight^2 
-(\conf_{i\,0}-\conf_{i-1\,0})}}, \text{where} \  1\leq i\leq wp+1 
+(\conf_{i\,0}-\conf_{i-1\,0})}}, \  i \in [1\,..\, wp+1]  
 $$
@@
 Without these coefficients, the path that minimizes the cost corresponds to a 
-straight line with the waypoints equidistantly allocated, which is not adapted for 
-(Figure~\ref{local_box_optim}) type of problems with a local passage very
+straight line with the waypoints equidistantly allocated. This is not relevant for 
+(Figure~\ref{local_box_optim}) type of problems where a local passage is very
 constrained by obstacles. Note that this cost is also working with all other 
@@
 
-We also experiment our algorithm on more complex robots and environments.
-
 \begin{algorithm}
@@
 \OUTPUT{Optimized collision-free path $\xx$}
-   \State nbFailures $\gets 0$
+   \State $nbFailures$ $\gets 0$
    \While {$nbFailures < maxNbFailures$}
         \State $failure \gets true$
-        \State $T \gets \mbox{upper bound of }\xx\mbox{ definition interval}$
-        \State $t_1 < t_2 \gets \mbox{random numbers in }[0,T]$
+        \State $t_1 < t_2 \gets \mbox{random numbers in }[0,1]$
         \State $lp0 \gets \texttt{steeringMethod} (\xx(0), \xx(t_1))$
         \State $lp1 \gets \texttt{steeringMethod} (\xx(t_1), \xx(t_2))$
-        \State $lp2 \gets \texttt{steeringMethod} (\xx(t_2), \xx(T))$
+        \State $lp2 \gets \texttt{steeringMethod} (\xx(t_2), \xx(1))$
         \State $newPath \gets \mbox{empty path defined on }[0,0]$
-   	\If{$lp0$ is collision-free}
+   	\If{\texttt{validatePath($lp0$)}}
           \State $newPath \gets lp0$; $failure \gets false$
@@
         \EndIf
-   	\If{$lp1$ is collision-free}
+   	\If{\texttt{validatePath($lp1$)}}
           \State $newPath \gets \texttt{concatenate} (newPath, lp1)$
@@
         \EndIf
-   	\If{$lp2$ is collision-free}
+   	\If{\texttt{validatePath($lp2$)}}
           \State $newPath \gets \texttt{concatenate} (newPath, lp2)$
@@
         \Else
-          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_2,T]})$
+          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_2,1]})$
         \EndIf
@@
 
+We also experiment our algorithm on more complex robots and environments.
 In the included video and (Figures~\ref{pr2_final},~\ref{fiad} 
@@
 The termination condition of RS allows it to try 
-shortening the path until five iterations of non-improvement are reached 
-(corresponding to $maxNbFailures$ in (Algorithm~\ref{algo:random-shortcut})).
+shortening the path until five iterations of non-improvement are reached. 
+This corresponds to $maxNbFailures$ in (Algorithm~\ref{algo:random-shortcut}).
 For our implementation of PRS, first a shortcut is tried on each DOF between 
@@
 % freeflyer puzzle
-Before entering the manipulator examples, the GB algorithm is tested and compared 
-on a problem very popular in the motion planning literature: a freeflyer puzzle 
-example, corresponding to (Figure~\ref{fig:trajectories:puzzle}).  The puzzle has to cross down the 
+Before entering the manipulator examples, the GB algorithm is analyzed 
+on a popular problem in the motion planning literature: a freeflyer puzzle, 
+corresponding to (Figure~\ref{fig:trajectories:puzzle}). The 
+puzzle has to cross down the 
 obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
@@
 this problem, all DOF are closely related to cross the hole. This behavior was also 
-noticed and improve in the PRS implementation of~\cite{Geraerts04clearancebased}, 
+noticed and improved in the PRS implementation of~\cite{Geraerts04clearancebased}, 
 acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory 
@@
 % double-arm:
-On the double-arm problem, (Figure~\ref{fig:trajectories:double_arm}), one arm has 
+In the double-arm situation, (Figure~\ref{fig:trajectories:double_arm}), one arm has 
 to get around a cylinder obstacle while the 
@@
 	industrial problem, where initial and final configurations of the UR5 are 
-	constrained in boxes. Trajectories results 
-	are presented (bottom right): initial (RRT-connect) path in red, RS 
+	constrained in boxes. End-effector trajectories 
+	are illustrated (bottom right): initial (RRT-connect) in red, RS 
 	optimization in blue and the GB optimization in green.}
@@
 A problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
-manipulating in an office environment is presentated 
+manipulating in an office environment is presented 
 (Figure~\ref{fig:trajectories:baxter}). The robot starts with its end-effectors 
@@
 
-On the three following high-DOF examples involving the PR2 robot, we argue that 
-results are better in terms of path quality, in the fact that parasite DOF motions 
-are removed.
+In the three following high-DOF examples involving the PR2 robot, note that 
+results are better in terms of path quality, as a result of the parasite 
+DOF motions removal.
 
 % analysis PR2 alone:
-In the example presented (Figure~\ref{pr2_final}) and in the video, the mobile 
+In the example shown (Figure~\ref{pr2_final}) and in the video, the mobile 
 40-DOF PR2 simply has to cross its arms from 
-the left arm up position to the right arm up one, without any assumption on the DOF or group of DOF to choose to plan or optimize (i.e. no DOF is locked). 
+the left arm up position to the right arm up one, without any assumption about the group of DOF to activate (i.e. no DOF is locked). 
 The RRT-connect planner 
@@
 results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating 
-two collision-constraints between the arms. The PRS 
-result, only presented in the video, cancels the mobile base motion, but not all 
-unnecessary DOF. In addition, the arms motion is very wide and not as optimized as in the GB trajectory.
+two collision-constraints between the arms. In the PRS 
+result, only presented in the video, the mobile base motion is canceled.
+However, this is note the case for other unnecessary DOF motions.
+In addition, the arms motion is very wide and not as optimized as in the GB trajectory.
 
 % analysis PR2 in kitchen:
-We obtain similar results on the PR2 performing a manipulation task 
-in a kitchen environment (see the video and (Figure~\ref{fig:trajectories:kitchen1}) 
-to visualize the trajectory of the right gripper). The robot has to move 
-its hands from the top to the bottom of a table.
-Our optimizer manages to reduce the initial length 12.7 from the visibility-PRM 
-planner and improves the path quality 
-just adding orthogonality constraints between the table and the 
-robot's arms and grippers. Thus, the robot just slightly moves 
+We obtain similar results on the PR2 performing manipulation tasks 
+in a kitchen environment.
+Firstly, the robot moves its hands from the top to the bottom of a table.
+The different trajectories of the right gripper are indicated in (Figure~\ref{fig:trajectories:kitchen1}).
+Our optimizer manages to reduce the initial path length 12.7 from visibility-PRM 
+and improves the path quality 
+just adding constraints between the table and the 
+robot's arms. Thus, the robot just slightly moves 
 backward and uses its arms DOF to avoid the table, instead of 
 processing a large motion to get away from the table. With the GB algorithm, the path length is 
-downed to 3.15, against 5.47 for RS.
-Another example of PR2 working in the kitchen, going from the set to the fridge 
-door, is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile base
+reduced to 3.15, against 5.47 for RS.
+Secondly, another example of PR2 going from the set to the fridge 
+door is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile base
 trajectories. Here, GB and RS results are similar in terms of length and rendering.
@@
 Optimization computation time and path length averages are presented for 30 runs of 
-each example in (Table~\ref{tab:results}). In some cases $\alpha_{init}$ was 
+each example in (Table~\ref{tab:results}). In some cases, $\alpha_{init}$ was 
 reduced to comply with very narrow passages, or increased in the opposite case. 
@@
 constraints of GB is significant and $\alpha_{init}$ reduced. 
-In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimization spends time on the QP resolution to avoid collisions.
+In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimizer spends time on the QP resolution to avoid collisions.
 On the other hand, our method almost always provides the shortest paths, whereas it 
 is not even the best criterion to quantify the parasite motions disappearance.
-Such results are the artefact of the termination conditions of RS and PRS, length 
-reduction or affordable computation time. This is particularly visible for PRS which 
+Note that these results are the consequence of the termination conditions of RS and PRS, balancing length reduction or shorter computation time. 
+This is particularly visible for PRS which 
 easily gives up in some problems, such as the 2D sliding-robot, the 
@@
 optimized paths is highlighted in the right column of (Table~\ref{tab:results}). 
-Thus the problem addressed in (Figure~\ref{decoupled_DOF_optimization}) is 
+Thus the issue addressed in (Figure~\ref{decoupled_DOF_optimization}) is 
 solved considering this DOF.
+
+\section{CONCLUSIONS}
+We managed to settle a path optimization for navigation and manipulation problems, and tested it with various robots and environments. Our algorithm uses standard 
+numerical tools as collision checking, linearized one-dimensional constraint
+and QP resolution. It correlates them in a 
+simple but effective way, and the algorithm structure is organized so that its 
+convergence is guaranteed. Furthermore, 
+our method only requires collision checking, therefore neither 
+geometry pre-processing nor 
+offline optimization are necessary to remain time-competitive. We demonstrate 
+that the optimizer may be 
+time-competitive compared to random shortcut in complex models where collision tests 
+are time-consuming. 
+It also proposes better quality paths, reducing the path length and removing 
+unnecessary DOF motions. 
+Finally, our optimizer manages to reduce a local detour in a long path while random 
+shortcut methods will mostly fail.
+
+For future work, we have room for improvement. We can take advantage of the sparsity of 
+the constraint Jacobian to reduce computation time. We may also adapt the iteration 
+scalar parameter from geometries considerations on the current path, e.g. using a lower 
+bound of the distance between nearest objects, since this bound can be returned by 
+the collision checker.
+
+\section*{Funding}
+This work has been supported by the project ERC Advanced Grant 340050 Actanthrope and by the FP~7 project Factory in a Day under grant agreement n°~609206.
 
@@
 
-\section{CONCLUSIONS}
-We managed to settle a path optimization for navigation and manipulation problems, and tested it with various robots and environments. Our algorithm uses standard 
-numerical tools as collision checking, linearized one-dimensional constraint
-and QP resolution. It correlates them in a 
-simple but effective way, and the algorithm structure is organized so that the convergence is guaranteed. Furthermore, 
-our method only requires collision checking, therefore geometry pre-processing or 
-offline optimization are not necessary to remain time-competitive. We demonstrate 
-that the optimizer may be 
-time-competitive compared to random shortcut in complex models where collision tests are time-consuming. 
-It also proposes better quality paths, reducing the path length and removing unnecessary DOF motions. 
-Finally, our optimizer manages to reduce a local detour in a long path while random shortcut methods will mostly fail.
-
-For future work, we have room for improvement: taking advantage of the sparsity of 
-the constraint Jacobian to reduce computation time. We may also adapt the iteration 
-scalar parameter from geometries considerations on the current path, e.g. a lower 
-bound of the distance between nearest objects, since this bound can be returned by 
-the collision checker.
-
-\section*{Funding}
-This work has been supported by the project ERC Advanced Grant 340050 Actanthrope and by the FP~7 project Factory in a Day under grant agreement n°~609206.
-
 \bibliographystyle{tADR}

### 2016-01-07 paper.tex
Remove Figures brackets and fix typo of reviewer 2

@@
 
+% one line version:
+%Most algorithms in probabilistic sampling-based path planning compute collision-free paths made of straight line segments lying in the configuration space.
+Due to the randomness of sampling, the paths make detours that need to be optimized.
+The contribution of this paper is to propose a basic gradient-based (GB) algorithm that transforms a polygonal collision-free path into a shorter one.
+While requiring only collision checking, and not any time-consuming obstacle distance computation nor geometry simplification, we constrain only part of the configuration variables that may cause a collision, and not entire configurations.
+Thus parasite motions that are not useful for the problem resolution are reduced without any assumption.
+Experimental results include navigation and manipulation tasks, e.g.
+a manipulator arm filling boxes and a PR2 robot working in a kitchen environment.
+Comparisons with a random shortcut optimizer and a partial shortcut have also been studied.
+
 % few parameter tuning REMOVED
@@
 trajectories where only a minor part needs to be optimized, 
-as in (Figure~\ref{local_box_optim}).
-(Figure~\ref{decoupled_DOF_optimization}) presents another situation where RS 
+as in Figure~\ref{local_box_optim}.
+Figure~\ref{decoupled_DOF_optimization} presents another situation where RS 
 will always fail to optimize the initial path, since it cannot decouple the 
@@
 on some random DOF) 
-which can address the problem of (Figure~\ref{decoupled_DOF_optimization}). 
+which can address the problem of Figure~\ref{decoupled_DOF_optimization}. 
 However it is 
@@
 freeze the whole robot configuration instead of a pertinent subpart. Thus, this 
-method cannot solve (Figure~\ref{decoupled_DOF_optimization}) issue.
+method cannot solve Figure~\ref{decoupled_DOF_optimization} issue.
 
@@
 
-As illustrated in (Figure~\ref{optim_grad}), let 
+As illustrated in Figure~\ref{optim_grad}, let 
 us denote by $\tcolli$ the abscissa of the first collision detected on path 
@@
 iterations until the redundancy vanishes, as explained in function 
-\texttt{findNewConstraint} in (Algorithm~\ref{algo:solveRedundant}), and illustrated in (Figure~\ref{fig:convergence-diagram}).
+\texttt{findNewConstraint} in (Algorithm~\ref{algo:solveRedundant}), and illustrated in Figure~\ref{fig:convergence-diagram}.
 
@@
 
-(Figure~\ref{2D_long}) shows the result of our optimizer on 2D cases. Contact points which have 
+Figure~\ref{2D_long} shows the result of our optimizer on 2D cases. Contact points which have 
 led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
@@
 
-(Figure~\ref{local_box_optim}) illustrates a \textit{very long} path example which RS 
+Figure~\ref{local_box_optim} illustrates a \textit{very long} path example which RS 
 or PRS will not manage to 
@@
 straight line with the waypoints equidistantly allocated. This is not relevant for 
-(Figure~\ref{local_box_optim}) type of problems where a local passage is very
+Figure~\ref{local_box_optim} type of problems where a local passage is very
 constrained by obstacles. Note that this cost is also working with all other 
@@
 We also experiment our algorithm on more complex robots and environments.
-In the included video and (Figures~\ref{pr2_final},~\ref{fiad} 
-and~\ref{fig:trajectories}), we present multiple situations where the GB algorithm 
+In the included video and Figures~\ref{pr2_final},~\ref{fiad} 
+and~\ref{fig:trajectories}, we present multiple situations where the GB algorithm 
 is tested and compared to RS~\cite{randomShortcutHPP}
@@
 on a popular problem in the motion planning literature: a freeflyer puzzle, 
-corresponding to (Figure~\ref{fig:trajectories:puzzle}). The 
+corresponding to Figure~\ref{fig:trajectories:puzzle}. The 
 puzzle has to cross down the 
@@
 % double-arm:
-In the double-arm situation, (Figure~\ref{fig:trajectories:double_arm}), one arm has 
+In the double-arm situation, Figure~\ref{fig:trajectories:double_arm}, one arm has 
 to get around a cylinder obstacle while the 
@@
 gripper.
-In a relatively free environment (Figure~\ref{fig:trajectories:ur5}), 
+In a relatively free environment represented in Figure~\ref{fig:trajectories:ur5}, 
 results from our method and RS are similar (path length are respectively 2.03 and 1.60). 
@@
 For an UR5 working in a cluttered environment inspired by an industrial issue 
-(Figure~\ref{fiad}), path optimization efficiently returns a shorter solution, close 
+Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
 to the result of RS and to what can be observed in reality~\cite{factory-day-video}. 
@@
 A problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
-manipulating in an office environment is presented 
-(Figure~\ref{fig:trajectories:baxter}). The robot starts with its end-effectors 
+manipulating in an office environment is presented in
+Figure~\ref{fig:trajectories:baxter}. The robot starts with its end-effectors 
 above the computer and has to 
@@
 % analysis PR2 alone:
-In the example shown (Figure~\ref{pr2_final}) and in the video, the mobile 
+In the example shown Figure~\ref{pr2_final} and in the video, the mobile 
 40-DOF PR2 simply has to cross its arms from 
@@
 result, only presented in the video, the mobile base motion is canceled.
-However, this is note the case for other unnecessary DOF motions.
+However, this is not the case for other unnecessary DOF motions.
 In addition, the arms motion is very wide and not as optimized as in the GB trajectory.
@@
 Firstly, the robot moves its hands from the top to the bottom of a table.
-The different trajectories of the right gripper are indicated in (Figure~\ref{fig:trajectories:kitchen1}).
+The different trajectories of the right gripper are indicated in Figure~\ref{fig:trajectories:kitchen1}.
 Our optimizer manages to reduce the initial path length 12.7 from visibility-PRM 
@@
 Secondly, another example of PR2 going from the set to the fridge 
-door is presented (Figure~\ref{fig:trajectories:kitchen2}) with the mobile base
+door is presented Figure~\ref{fig:trajectories:kitchen2} with the mobile base
 trajectories. Here, GB and RS results are similar in terms of length and rendering.
@@
 optimized paths is highlighted in the right column of (Table~\ref{tab:results}). 
-Thus the issue addressed in (Figure~\ref{decoupled_DOF_optimization}) is 
+Thus the issue addressed in Figure~\ref{decoupled_DOF_optimization} is 
 solved considering this DOF.

### 2016-01-07 paper.tex
Change Geraerts reference from ICRA to IJRR

@@
 still very 
-popular~\cite{Sekhavat-Svestka1998,Geraerts04clearancebased,HauserFastSmooth}. 
+popular~\cite{Sekhavat-Svestka1998,GeraertsIJRR07,HauserFastSmooth}. 
 However, RS requires fine 
@@
 locally optimal, and is not addressing high-DOF 
-problems.~\cite{Geraerts04clearancebased} is 
+problems.~\cite{GeraertsIJRR07} is 
 using medial axis retraction for clearance and a PRS (shortcut is applied only 
@@
 However it is 
-relatively slower than RS, and~\cite{Geraerts04clearancebased} only investigates 
+relatively slower than RS, and~\cite{GeraertsIJRR07} only investigates 
 it for freeflyer robots. 
@@
 (Algorithm~\ref{algo:random-shortcut}) and PRS~\cite{partialrandomShortcutHPP} 
-adapted from~\cite{Geraerts04clearancebased}.
+adapted from~\cite{GeraertsIJRR07}.
 The termination condition of RS allows it to try 
@@
 this problem, all DOF are closely related to cross the hole. This behavior was also 
-noticed and improved in the PRS implementation of~\cite{Geraerts04clearancebased}, 
+noticed and improved in the PRS implementation of~\cite{GeraertsIJRR07}, 
 acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory 

### 2016-01-11 paper.tex
Changes according to reviews:
* move Figure 6 (very long path optim) to introduction
* change title
* change index to be consistent (i is only for path iterations and begining of Sec.3)
* explain unclear notations in figure caption (x_Free, x_Coll, p)
* remove reference to [1] (our ICRA paper version)
* start to change the way to declare the algorithm's functions (until now: remove too early
ref to Algorithm)

@@
 \newcommand\translation{\mathbf{t}}
-\newcommand\tcolli{t_{coll\ i}}
+%\newcommand\tcolli{t_{coll\ i}}
+\newcommand\tcolli{\kappa_i}
 \newcommand\po{\mathbf{po}}
@@
 
-\title{A simple path optimization method for motion planning$^{1}$ \thanks{$^{1}$Note for the reviewers: 
-This paper is an extended version of a paper submitted to IEEE ICRA 2016~\cite{pathOptimICRA_HAL}. The 
-current version includes a problem formulation, a convergence analysis and many experimental results (including an 
-industrial example) which are absent from the conference version.}}
+\title{\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$}
 
@@
 
+\textcolor{blue}{Definition of what could be optimized (path length, execution time, number of via points, ...}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=10cm]{local_box_optim2.pdf}
+	\caption{Case of a long initial path (above) containing a small part 
+	that can 
+	be optimized (below). Random shortcut is unlikely to optimize the initial 
+	dashed part containing detours in the box, whereas our method 
+	succeeds (in blue). This type of issue is common in navigation problems, in 
+	environments with long corridors.}
+	\label{local_box_optim}
+\end{figure}
+
 For this shortening purpose, random shortcut (RS) methods are
@@
 trajectories where only a minor part needs to be optimized, 
-as in Figure~\ref{local_box_optim}.
+as in \textcolor{blue}{Figure~\ref{local_box_optim}}.
 Figure~\ref{decoupled_DOF_optimization} presents another situation where RS 
@@
 A robot is defined by a kinematic chain composed of a tree of joints. We denote 
-by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $1\leq i\leq 
-m$ is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
+by $(J_1,\cdots,J_m)$ the ordered list of joints. Each joint $J_i$, $i=1\cdots m$, is represented by a mapping from a sub-manifold of $\real^{n_i}$, where $n_i$ 
 is the dimension of $J_i$ in the configuration space, to the space 
@@
 collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
-paths and to return the first colliding configuration along a path. This stage 
-corresponds to \texttt{validatePath} in (Algorithm~\ref{algo:gradient}), which returns $true$ if the given path is collision-free.
+paths and to return the first colliding configuration 
+\textcolor{blue}{and its abcissa} along a path. \textcolor{blue}{This stage 
+is handled by a function \texttt{validatePath}, which returns $true$ if the given path is collision-free}.
 
@@
 $$
-\cost (\xx) \triangleq \frac{1}{2}\sum_{i=1}^{wp+1} \lambda_{i-1} \|\conf_{i}-\conf_{i-1}\|_{W}
+\cost (\xx) \triangleq \frac{1}{2}\sum_{k=1}^{wp+1} \lambda_{k-1} \|\conf_{k}-\conf_{k-1}\|_{W}
 ^{2}
 $$
-where the $\lambda_{i-1}$ coefficients will be explained in the results 
-section. For now it is assumed that $(\lambda_{i-1})_{1\leq i\leq wp+1}=1$. Note that 
+where the $\lambda_{k-1}$ coefficients will be explained in the results 
+section. For now it is assumed that $(\lambda_{k-1})_{k=1\cdots wp+1}=1$. Note that 
 $\cost$ is not exactly the length of the path, but it can be 
@@
 \nabla \cost (\xx) = 
-\left( (\lambda_{i}(\conf_{i+1} - \conf_{i})^T - \lambda_{i+1}(\conf_{i+2} - \conf_{i+1})^T) \weight^2 \right)_{i=0\cdots wp-1}
+\left( (\lambda_{k}(\conf_{k+1} - \conf_{k})^T - \lambda_{k+1}(\conf_{k+2} - \conf_{k+1})^T) \weight^2 \right)_{k=0\cdots wp-1}
 $$
@@
 \end{equation}
-where $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
+where \textcolor{blue}{$\p_i$ is computed by an unconstrained version of function \texttt{computeIterate}, and} $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
 unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
@@
 
-We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision. This 
-corresponds to the unconstrained function \texttt{computeIterate} in 
-(Algorithm~\ref{algo:gradient}).
+We iterate step~(\ref{eq:iteration-1}) until path $\xx_{i+1}$ is in collision.
 When a collision is detected, we introduce a constraint and perform a new 
@@
 	\includegraphics[width=9cm]{optim_grad.pdf}
-	\caption{Illustration of one iteration of our optimization. $\xx_{i+1}$ 
-	appears 
+	\caption{Illustration of one iteration of our optimization.
+	$\xx_{i+1}$ appears 
 	to be in collision with the obstacle. The first colliding configuration 
-	$\xx_{i+1}(\tcolli)$ is returned by the continuous collision checker. The 
+	$\xx_{i+1}(\tcolli)$ \textcolor{blue}{at abcissa $\tcolli$} is returned by 
+	the continuous collision checker. The 
 	corresponding constraint will be computed in the backtracked configuration 
@@
 
-Let us assume that at iteration $i$, $k$ constraints have been inserted before the current iteration. 
+\textcolor{blue}{The constraint function is noted $f$}. Let us assume that at iteration $i$, $j$ constraints have been inserted before the current iteration. 
 These constraints are stored as lines of a Jacobian matrix as follows:
 $$
-\Jf_{i} = \left(\begin{array}{c}L_1 \\ \vdots \\ L_k\end{array}\right)
+\Jf_{i} = \left(\begin{array}{c}L_1 \\ \vdots \\ L_j\end{array}\right)
 $$
@@
 Notice that $\mathbf{u}$ is well defined since configuration 
-$\xx_{i}(t_{coll\ i})$ is collision-free.
+$\xx_{i}(\tcolli)$ is collision-free.
 Let $g$ be the function defined for any $\conf \in \CS$ to $\real$ by:
@@
 \begin {align*}
-\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{k+1}\end{array}\right)\ \mbox {with}\\
-L_{k+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
+\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{j+1}\end{array}\right)\ \mbox {with}\\
+L_{j+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
 \end{align*}
@@
 Finally, we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
-programs (QP) (function \texttt{computeIterate} in 
-(Algorithm~\ref{algo:gradient})).
+programs (QP). \textcolor{blue}{The QP step computation corresponds to a constrained 
+version of \texttt{computeIterate}}.
 
@@
 
-Let $\conf_{k\,i}$ denote the waypoint $k$ along path $\xx_i$.
+Let $\conf_{k,i}$ denote the waypoint $k$ along path $\xx_i$.
 There exist $\beta\in[0,1]$ and $k$ such that $\xx_i (\tcolli)$ can be written as a combination of two waypoints:
 $$
-\xx_i (\tcolli) = \conf_{k\,i} + \beta (\conf_{k+1\,i} - \conf_{k\,i})
+\xx_i (\tcolli) = \conf_{k,i} + \beta (\conf_{k+1,i} - \conf_{k,i})
 $$
@@
 is built by matrix 
-blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in each of the two waypoints
-and $\beta$. This stage corresponds to \texttt{computeCollisionConstraint} in 
-(Algorithm~\ref{algo:gradient}).
+blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in 
+each of the two waypoints and $\beta$. This stage \textcolor{blue}{is refered as 
+to \texttt{computeCollisionConstraint}}.
 
@@
 \end{equation}
+\textcolor{blue}{In other words, any iteration step $\p_i$ that complies with the 
+collision constraints on path $\xx_{i+1}$, necessarily sastifies $\xx_i$ constraints.}
 Let us assume that:
@@
 \end{equation}
-We will elaborate later on this assumption. This means that $\p_{i}\notin\kernel J_{i+1}$, and as $\p_{i}\in\kernel J_{i}$, then:
-$$
-\kernel J_i \not= \kernel J_{i+1}
+We will elaborate later on this assumption. This means that 
+$\p_{i}\notin\kernel \Jf_{i+1}$, and as $\p_{i}\in\kernel \Jf_{i}$, then:
+$$
+\kernel \Jf_i \not= \kernel \Jf_{i+1}
 $$
@@
 $$
-\dim (\kernel J_{i+1}) < \dim (\kernel J_i)
+\dim (\kernel \Jf_{i+1}) < \dim (\kernel \Jf_i)
 $$
@@
 \begin{equation}\label{eq:trajectory}
-\conf (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (t_{coll\ i}), \ t \in [0,1]
+\conf (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
 \end{equation}
-where the motion is defined by taking a constant abscissa $\tcolli$ and by moving from the 
-path $\xx_i$ along the step $\p_i$. Note that configuration $\xx_{i+1}(t_{coll\ i})$ is reached when $t=1$.
-
-Thus the velocity of point $\po_2(q(t))$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
+\textcolor{blue}{where $\xx_i$ is still collision-free and $\xx_{i+1}$ 
+has its first collision detected at abcissa $\tcolli$.}
+The motion $\conf(t)$ is defined by taking a constant abscissa $\tcolli$ and by 
+moving from the 
+path $\xx_i$ along the step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
+
+Thus the velocity of point $\po_2(\conf(t))$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
 $$
@@
 $$
-correspond to $(\dot{\po_2}|\mathbf{u})$, respectively in configurations $\xx_{i}(t_{coll\ i})$ and $\xx_{i+1}(t_{coll\ i})$. The term $\dot{\conf}(1)$ being equal to zero requires either:
+correspond to $(\dot{\po_2}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$. The term $\dot{\conf}(1)$ being equal to zero requires either:
 \begin {enumerate}
@@
 iterations until the redundancy vanishes, as explained in function 
-\texttt{findNewConstraint} in (Algorithm~\ref{algo:solveRedundant}), and illustrated in Figure~\ref{fig:convergence-diagram}.
+\texttt{findNewConstraint} in Algorithm~\ref{algo:solveRedundant}, and illustrated in Figure~\ref{fig:convergence-diagram}.
 
@@
 	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
-	\caption{Representation of a trajectory $\conf (t)$ from 
-	Equation~(\ref{eq:trajectory}) in the 
-	configuration space (left). For the same trajectory in the workspace, the 
-	colliding bodies are illustrated (right). Note that a case with 
+	\caption{\textcolor{blue}{Considering a trajectory 
+	$\conf (t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
+	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
+	and $\kappa$ resp. collision-free path, in-collision path, iteration step 
+	and collision-abcissa).
+	The trajectory is represented in the configuration space (left) and in the 
+	workspace with the colliding bodies (right).}
+	Note that a case with 
 	$(\dot{\po_2}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
 	with the constraint that should be added if no redundancy was tracked. Therefore, 
-	(Algorithm~\ref{algo:solveRedundant}) will produce one intermediate step
+	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
 	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
@@
 	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
+	\textcolor{blue}{\State $\alpha \gets 0.5\alpha$}
 \EndWhile
@@
 
-In this section, the GB path-optimizer (Algorithm~\ref{algo:gradient}) is described
+In this section, the GB path-optimizer Algorithm~\ref{algo:gradient} is described
 according to the previous steps definitions. One difficulty here is to handle 
@@
 $$
-\lambda_{i-1} = \frac{1}{\sqrt{(\conf_{i\,0}-\conf_{i-1\,0})^T \weight^2 
-(\conf_{i\,0}-\conf_{i-1\,0})}}, \  i \in [1\,..\, wp+1]  
+\lambda_{k-1} = \frac{1}{\sqrt{(\conf_{k,0}-\conf_{k-1,0})^T \weight^2 
+(\conf_{k,0}-\conf_{k-1,0})}}, \  k=1\cdots wp+1  
 $$
@@
 minimum as at 
-initial path, represented by the waypoints ($\conf_{i\,0})_{1\leq i\leq wp+1}$.
+initial path, represented by the waypoints ($\conf_{i,0})_{k=1\cdots wp+1}$.
 Without these coefficients, the path that minimizes the cost corresponds to a 
@@
 examples presented in this paper, and provides better quality results than the original cost.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=10cm]{local_box_optim.pdf}
-	\caption{Case of a long initial path (above) containing a small part 
-	that can 
-	be optimized (below). Random shortcut is unlikely to optimize the initial 
-	dashed part containing detours in the box, whereas our method 
-	succeeds (in blue). This type of issue is common in navigation problems, in 
-	environments with long corridors.}
-	\label{local_box_optim}
-\end{figure}
 
@@
 is tested and compared to RS~\cite{randomShortcutHPP}
-(Algorithm~\ref{algo:random-shortcut}) and PRS~\cite{partialrandomShortcutHPP} 
+Algorithm~\ref{algo:random-shortcut} and PRS~\cite{partialrandomShortcutHPP} 
 adapted from~\cite{GeraertsIJRR07}.
@@
 shortening the path until five iterations of non-improvement are reached. 
-This corresponds to $maxNbFailures$ in (Algorithm~\ref{algo:random-shortcut}).
+This corresponds to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}.
 For our implementation of PRS, first a shortcut is tried on each DOF between 
@@
 Then, for each DOF, configurations are sampled as in 
-(Algorithm~\ref{algo:random-shortcut}). The steering-method returns a path made of 
+Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path made of 
 an interpolation on only 
@@
 result, only presented in the video, the mobile base motion is canceled.
-However, this is not the case for other unnecessary DOF motions.
+However, this is \textcolor{blue}{not} the case for other unnecessary DOF motions.
 In addition, the arms motion is very wide and not as optimized as in the GB trajectory.

### 2016-01-11 answers-to-reviewers.tex
Starting version of answer-to-reviewers letter

\documentclass{scrartcl}
\usepackage[svgnames]{xcolor}
%\usepackage{biblatex}
\usepackage{tikz}
%\include{tikz_commands}
\usetikzlibrary{patterns}

\newcommand{\todo}{\textcolor{red}{TODO}}

%\bibliographystyle{unsrt}
%\bibliographystyle{IEEEtran}
\bibliographystyle{tADR}
\bibliography{paper}
%\bibliography{IEEEabrv,paper}

\begin{document}

\title{A simple path-optimization method for motion planning}
\author{M. Campana, F. Lamiraux and J.-P. Laumond}
\date{Answer to reviewers}
\maketitle

The authors would like to thank the editor and the reviewers for their helpful comments and remarks.
They have been taken into account in this revised version of the manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Editor}

\hrulefill\\

\subsection{Comment}
\hrulefill\\

Please revise the paper based on the comments, esp. the ones related to fair comparisons and other issues.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

The authors took into account all the remarks made by the editor and reviewers.
We improved the clarity of the paper and the consistency of the notation.

We would like to briefly restate here why we think the paper is showing substantial contribution with respect to [10].
When testing [10] on the HRP-2 located in LAAS, Toulouse the performances of the algorithm were really bad: limited to 10 cm/s in the sagittal plan, 
few centimeters/s sideways, rotational velocity were really small, and motion coupling X and Y velocities were not possible.
It is shown in the new companion video. 
The introduction of the dynamical filter is now allowing us to have 30 cm/s forward, achieve motion with strong X-Y velocities.
The improvement is also shown in the new video.

The theoretical contribution in using a nonlinear formulation allow us to easily integrates obstacle in the system without resorting to any preprocessing
which would be necessary when using linear MPC approaches.
We do believe that this formulation might be applicable to other situations and that it will be very helpful for other application such as visual-servoing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer 1}
\label{r2}
\hrulefill\\

\subsection{Comment 1}
\hrulefill\\

Title:
I do not see why the authors have named it “A simple path optimization method” as it is significantly more complex than the classical random sampling based methods. A title with something like “A gradiant based optimization method…” would be much more fitting in my view.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

%The authors agree with the reviewers that this sentence should be modified as it is unclear.
We thank the reviewer for the suggestion. The proposed title is indeed much more precise. We changed the title accordingly.

\subsection{Comment 2}
\hrulefill\\

Introduction: 
It would be very nice with a discussion of what optimal means. Optimal can be with respect to path length (the classical criteria also used in this paper), but could just as well be optimal with respect to execution time, where the number of via points and the clearance (allowing for larger blends) are important factors. Optimal could also depend on some task specific objectives.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We have developed the meaning of optimality, and we refer to a recent paper ACM-Com by one of the author.

Concerning the choice to optimize the path length, we... by choosing the length we get a quadratic criterion, which is easier blablabla...

\subsection{Comment 3}
\hrulefill\\

The introduction contains a reference to figure 6, which is in the results section. This long reference forces the reader to flip many pages back and forth. Please either just explain the content or put the figure earlier.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

As suggested, the figure was moved to the introduction.

\subsection{Comment 4}
\hrulefill\\

Related work: 
The work in [11] is also described in an IJRR paper of the same authors. This paper could be cited as well.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

As suggested, the reference has been changed to cite the journal paper.

\subsection{Comment 5}
\hrulefill\\

Section 3:
The paper starts describing the details of the algorithm and reference to algorithm 2 placed several pages later. It would make the paper more readable if the overall algorithm was presented earlier making the reader aware of where the components would be used before describing them. This would also mean that the reader does not have to flip back and forth all the time.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We understand that starting to define the algorithm's functions and reference to it too early is confusing. To enhance it, we choose to decouple the algorithm from the problem definition, and define the functions in the algorithm section. This method may seem redundant but avoid the reader to flip back and forth to find the functions definitions.
Furthermore, we prefer to not introduce the algorithm earlier because .... \todo

\subsection{Comment 6}
\hrulefill\\

Section 3.5.1:
This section does not read a nicely as the previous. Simplifying the notation by using sub/super scripts to define the reference frame of points rather than having the long expression for u, may help a bit.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 7}
\hrulefill\\

Results:
The authors only compares their approach with random shortcut and partial random shortcut, but omit the standard path pruning technique also described in [11]. This technique works excellent for the special case in Figure 6.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER
It is true that in the case of Figure 6, Prune is complementary to random shortcut, but this method, used alone, has also it's counter-examples, as the one presented in Figure 1 (and the PR2-crossing-arms Figure \todo).

\subsection{Comment 8}
\hrulefill\\

In the paper the authors state that it can be hard to select the stopping criteria of the random short cutting algorithms, for which the authors have selected five iterations with non-improvement. Looking at the time it takes before the optimization algorithms terminates (Table 2) it appears that this criterion is too harsh as the PRS and RS in many cases terminates significantly before the GB based method. To show if their method is indeed more efficient they should try to let the RS and PRS run for the same amount of time as the GB method. It would also be nice if Table 2 would include the number of iterations and potentially selected plots showing how the path length decreases as a function of time/iterations.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 9}
\hrulefill\\

Related to my previous comment about path pruning, then I would generally recommend people to use a combination of the deterministic path pruning followed by random short cut algorithm, as this in my experience performs better (in terms of how much it can optimize in a given time) then using just short cutting.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer 2}

\subsection{Comment}
\hrulefill\\

It looks that the conference version of this paper [1] is not cited in the main text though it appears in the Reference section.  Please add the citation in an appropriate place.

[1] Campana M, Lamiraux F, Laumond JP. A simple path optimization method for motion planning. https:
//hal.archives-ouvertes.fr/hal-01220048. 2015.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

We removed the reference.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reviewer 3}

\subsection{Comment 1}
\hrulefill\\

1) Notation and clarity of the presentation:

When introducing the Jacobian of f, at no point has f been introduced as the constraint function. I think in general the problem should be introduced as an optimization problem, what is the objective, what are the constraints, what simplifications will be made.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 2}
\hrulefill\\

In that regard it would be nice to comment the Newton method, for example there exists many methods to select alpha. Having a constant alpha the most naïve approach.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 3}
\hrulefill\\

I wonder if the discussion concerning the configuration space, which is well done by the way, is of importance here, it is a perquisite in the robotics community.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 4}
\hrulefill\\

The iteration index notation and the time index use the same letter (i), making it difficult to follow. The p variable is sometimes called back with mention of what it is.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 5}
\hrulefill\\

The kernel discussion, and velocity discussion are a bit breve and unclear for the reader.

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER
%The authors agree with the reviewers that this part should be modified as it is unclear.

\subsection{Comment 6}
\hrulefill\\

2) Too few and unfair comparisons

The experimental section leaves the impression that the comparisons with other methods are either giving the advantage to the state-of-the-art or, simply unfair.

Tuning the shortcut method by changing the distribution of samples along the path drastically changes the converging behavior, so choosing '5 non improving iterations' as the stopping criterion is very arbitrary. And it is the case in the results that the GB method beats the other methods using 1000 times the budget allocated to the other methods... I think reducing the comparison to one number in this case does not allow the reader to get a fair estimate of the convergence rate. A common criterion to be used is simply a max iteration count, which would already be fairer, but in order to not have these criteria influence the comparison, why not plot the averaged convergence curves?

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\subsection{Comment 7}
\hrulefill\\

Otherwise it would be interesting to see the alpha parameter being discussed at that point in the paper (with experimental results to back up the discussion).

\rule{\linewidth}{.1pt}
\textbf{Answer}:

ANSWER

\end{document}

### 2016-01-13 answers-to-reviewers.tex
Rewrite paper, fix 1st answer-to-reviewer, fix Figures typo. In paper:
* add criterion-optimality discussion
* change mathematical notations of contact-points frames
* start rewrite the algorithm text section

@@
 The authors took into account all the remarks made by the editor and reviewers.
-We improved the clarity of the paper and the consistency of the notation.
-
-We would like to briefly restate here why we think the paper is showing substantial contribution with respect to [10].
-When testing [10] on the HRP-2 located in LAAS, Toulouse the performances of the algorithm were really bad: limited to 10 cm/s in the sagittal plan, 
-few centimeters/s sideways, rotational velocity were really small, and motion coupling X and Y velocities were not possible.
-It is shown in the new companion video. 
-The introduction of the dynamical filter is now allowing us to have 30 cm/s forward, achieve motion with strong X-Y velocities.
-The improvement is also shown in the new video.
-
-The theoretical contribution in using a nonlinear formulation allow us to easily integrates obstacle in the system without resorting to any preprocessing
-which would be necessary when using linear MPC approaches.
-We do believe that this formulation might be applicable to other situations and that it will be very helpful for other application such as visual-servoing.
+We improved the clarity of the paper, the consistency of the notation and the relevance of the comparison.
 

### 2016-01-13 paper.tex
Rewrite paper, fix 1st answer-to-reviewer, fix Figures typo. In paper:
* add criterion-optimality discussion
* change mathematical notations of contact-points frames
* start rewrite the algorithm text section

@@
 \newcommand\p{\mathbf{p}}
-\newcommand\pii{\mathbf{p_{i,i+1}}}
-\newcommand\pij{\mathbf{p_{i,j}}}
 \newcommand\gT{\tilde{g}^T}
@@
 \newcommand\tcolli{\kappa_i}
-\newcommand\po{\mathbf{po}}
+\newcommand\po{\mathbf{P}}
 \newcommand\Jf{{J_f}}
 \newcommand\kernel{\mbox{Ker }}
+\newcommand\U{\mathbf{u}}
 
@@
 
+\textcolor{blue}{Optimization is always regarding to one or several criteria.
+The most common in motion planning are the path length, which penalize detours, the obstacle clearance for safety and the execution time, which is influenced by the number of via points, introducing velocity discontinuities.
+One can also mention the minimization of the acceleration or jerk peaks to implement safely the motion on a robot.}
+
 In this paper, we propose a method aimed at shortening path length after a path
@@
 
-\textcolor{blue}{Definition of what could be optimized (path length, execution time, number of via points, ...}
-
 \begin{figure}
@@
 	\includegraphics[width=10cm]{local_box_optim2.pdf}
-	\caption{Case of a long initial path (above) containing a small part 
+	\caption{Case of a long initial path from $\conf_{init}$ to $\conf_{final}$ (above) containing a small part 
 	that can 
@@
 
+\textcolor{blue}{This section describes the settling of the Gradient-based optimizer.
+The method works as a classical Quadratic Program (QP), reducing the path length with a cost function and avoid collisions with constraints.
+Details of the QP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}
+
 \subsection {Kinematic chain}
@@
 $$
-\Jf_{i} \p_i = 0
+\Jf_{i} \, \p_i = 0
 $$
@@
 (Figure~\ref{contact-points} left).
-In any configuration $\conf$, we define $\po_1(\conf)$ and $\po_2(\conf)$ as the coordinates of $\po_c$ in the respective local frames of $\body_1$ and $\body_2$ (Figure~\ref{contact-points} right).
-
+\textcolor{blue}{
+In any configuration $\conf$, let 
+$\po_j^l(\conf)$ be the coordinates in the frame of $\body_l$ of $\po_c$ expressed the frame of $\body_j$.
 Let $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) denote the homogeneous transformation 
-between frames of $\body_1$ (resp. $\body_2$) and the global frame in configuration 
-$\conf$. Let $\mathbf{u}$ be the unit vector 
-linking the contact points $\po_1(\xx_{i}(\tcolli))$ and 
-$\po_2(\xx_{i}(\tcolli))$:
-$$
-\mathbf{u} = \frac{M_{1}^{-1} (\xx_{i}(\tcolli)) M_2 (\xx_{i}(\tcolli)) \po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))}{\|M_{1}^{-1} 
-(\xx_{i}(\tcolli)) M_2 (\xx_{i}(\tcolli)) \po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))\|}
-$$
+between frames of $\body_1$ (resp. $\body_2$) and the global frame.
+Thus, $\po_2^1(\conf) = M_{1}^{-1} (\conf) M_2 (\conf)$.
+}
+
+\textcolor{blue}{
+In configuration $\xx_{i}(\tcolli)$, let $\mathbf{u}$ be the unit vector 
+linking the points $\po_1$ and $\po_2$ (Figure~\ref{contact-points} right):
+$$
+\U = \frac{\po_2^1(\xx_{i}(\tcolli)) - \po_1^1(\xx_{i}(\tcolli))}{\|\po_2^1(\xx_{i}(\tcolli)) - \po_1^1(\xx_{i}(\tcolli))\|}
+$$
+}
 
@@
 $\xx_{i}(\tcolli)$ is collision-free.
-Let $g$ be the function defined for any $\conf \in \CS$ to $\real$ by:
-$$
-g (\conf) = \left(M_{1}^{-1} (\conf) M_2 (\conf) \po_2(\conf) - \po_1(\conf) | \mathbf{u}\right)
-$$
-The function $g$
-represents the projection of $\overrightarrow{\po_1(\conf)\,\po_2(\conf)}$ 
-on $\mathbf{u}$.
+\textcolor{blue}{
+Let $g$ be the function representing the projection of $\overrightarrow{\po_1(\conf)\,\po_2(\conf)}$ on $\mathbf{u}$. $g$ is defined for any $\conf \in \CS$ to $\real$ by:
+$$
+g (\conf) = \left(\po_2^1(\conf) - \po_1^1(\conf) | \mathbf{u}\right)
+$$
+}
 %which is calculated from the last collision-free path at abscissa $\tcolli$.
@@
 \begin{algorithmic}%[1] % line numerotation
-\INPUT{Path to optimize $\xx_0$}
-\OUTPUT{Optimized collision-free path $\xx_0$}
+\INPUT{path to optimize $\xx_0$}
+\OUTPUT{optimized collision-free path $\xx_0$}
 \State $\alpha \gets \alpha_{init}$
@@
 In this section, the GB path-optimizer Algorithm~\ref{algo:gradient} is described
-according to the previous steps definitions. One difficulty here is to handle 
-the scalar parameter $\alpha$ determining how much of the computed step $\p$ will 
-be traveled along. We choose $\alpha_{init} = 0.2$ or 0.05 to process small steps.
-
-Each time a collision-constraint is added, 
-the solution of the current QP is tested (i.e. $\alpha = 1$). If this path is collision-free, the algorithm returns it as the solution. Otherwise, 
-smaller steps toward the minimum are iteratively applied and tested, and constraints are added each time a collision is found with a reduced step (i.e. $\alpha \neq 1$).
+according to the previous steps definitions.
+\textcolor{blue}{PARTIR A FINIR!!}
+\textcolor{blue}{Remind that, the QP optimal step $\p$ computed by \texttt{computeIterate} is known.
+The idea of the algorithm is to process ....
+So the difficulty here is to handle 
+the scalar parameter $\alpha$ determining how much of the computed step  will be traveled along. 
+As presented in Algorithm~\ref{algo:gradient}, $\alpha$ takes two values, $\alpha_{init}<1$ to process smaller steps or $1$ to go directly to the optimum.
+This latter case is interesting since, if this optimal path is collision-free, the algorithm has converged and returns the path as the solution.
+Choosing to travel small steps from a valid path decrease the chances of being in collision. Besides if one occurs, the collision-constraint is computed on the last valid path.
+$\alpha_{init}$ could be computed from geometrical considerations.
+...
+smaller steps toward the minimum are iteratively applied and tested, and constraints are added each time a collision is found with a reduced step (i.e. $\alpha \neq 1$).}
+
+The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.
+
 Note that even if the constraints are linearized, the algorithm converges.

### 2016-01-14 paper.tex
Fix Figures typo, add algorithm description, update RS and PRS

@@
 according to the previous steps definitions.
-\textcolor{blue}{PARTIR A FINIR!!}
 \textcolor{blue}{Remind that, the QP optimal step $\p$ computed by \texttt{computeIterate} is known.
-The idea of the algorithm is to process ....
-So the difficulty here is to handle 
+The idea of the algorithm is to process iterations that reduce the path length according to the QP cost.
+If collision-constraints have been added to the QP, further iterations will comply with them.
+The algorithm stops whether the QP minimum is reached and collision-free.
+The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.
+}
+
+\textcolor{blue}{
+The main difficulty is to handle 
 the scalar parameter $\alpha$ determining how much of the computed step  will be traveled along. 
@@
 This latter case is interesting since, if this optimal path is collision-free, the algorithm has converged and returns the path as the solution.
-Choosing to travel small steps from a valid path decrease the chances of being in collision. Besides if one occurs, the collision-constraint is computed on the last valid path.
-$\alpha_{init}$ could be computed from geometrical considerations.
-...
-smaller steps toward the minimum are iteratively applied and tested, and constraints are added each time a collision is found with a reduced step (i.e. $\alpha \neq 1$).}
-
-The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.
+Choosing to travel small steps from a valid path decrease the chances of being in collision.
+Besides if one occurs, the collision-constraint is computed on the last valid path, which is not too much deformed compared to the path that has collisions.
+As a result, collision-constraints are only computed (\texttt{computeCollisionConstraint}) and added (\texttt{addCollisionConstraint}) when performing a reduced iteration (i.e.
+$\alpha = \alpha_{init}$).
+}
 
@@
 RRT-connect~\cite{rrt-connect}. Unless another value is provided, $\alpha_{init}$ 
-is set to 0.2.
+is set to 0.2. \textcolor{blue}{A further section provides a discussion on the $\alpha_{init}$ value setting.}
 
@@
 
+\subsubsection{Analysis of $\alpha_{init}$ influence}
+\textcolor{blue}{Results without comparison to RS and PRS here
+$\alpha_{init}$ could be computed from geometrical considerations.
+}
+TODO !!
+
+\subsubsection{Comparison to random shortcut algorithms}
 \begin{algorithm}
 \begin{algorithmic}%[1] % line numerotation
-\INPUT{Path to optimize $\xx$}
+\INPUT{Path to optimize $\xx$, time limit $t_{lim}$}
 \OUTPUT{Optimized collision-free path $\xx$}
-   \State $nbFailures$ $\gets 0$
-   \While {$nbFailures < maxNbFailures$}
+   \State \textcolor{blue}{$t_{start} \gets \texttt{currentTime()}; \ t \gets 0$}
+   \While {\textcolor{blue}{$t < t_{lim}$}}
         \State $failure \gets true$
@@
    	\If{\texttt{validatePath($lp0$)}}
-          \State $newPath \gets lp0$; $failure \gets false$
+          \State $newPath \gets lp0$;
         \Else
@@
           \State $newPath \gets \texttt{concatenate} (newPath, lp1)$
-          \State $failure \gets false$
         \Else
@@
           \State $newPath \gets \texttt{concatenate} (newPath, lp2)$
-          \State $failure \gets false$
         \Else
@@
         \State $\xx \gets newPath$
-        \If {$failure$} $nbFailures \gets nbFailures + 1$
-        \EndIf
+        \State \textcolor{blue}{$t \gets \texttt{currentTime()} - t_{start}$}
       \EndWhile
@@
 two configurations. $\xx_{|I}$ denotes path $\xx$ restricted to interval $I$. 
-$maxNbFailures$ is a parameter that affects time of computation and quality of 
-the result.}\label{algo:random-shortcut}
+\textcolor{blue}{$t_{lim}$ represents the duration limitation of the algorithm.}}\label{algo:random-shortcut}
 \end{algorithm}
@@
 adapted from~\cite{GeraertsIJRR07}.
+\textcolor{blue}{
 The termination condition of RS allows it to try 
-shortening the path until five iterations of non-improvement are reached. 
-This corresponds to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}.
-For our implementation of PRS, first a shortcut is tried on each DOF between 
-$\conf_0$ and $\conf_{wp+1}$. 
-Then, for each DOF, configurations are sampled as in 
+shortening the path until the optimization duration overflows a time limit. 
+This corresponds to $t_{lim}$ in Algorithm~\ref{algo:random-shortcut} and is typically equal to the Gradient-Based optimization duration.
+%For our implementation of PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{wp+1}$. % NOT ANYMORE
+The implementation of PRS is identical to~\cite{GeraertsIJRR07}:
+for a random DOF, configurations are sampled as in 
 Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path made of 
@@
 the current DOF and based on the previous subpath for other DOF. If this path is 
-collision-free, we replace it as in RS. This process is also stopped when five 
-iterations of non-improvement are reached.
+collision-free, we replace it as in RS. As RS, the process is stopped when the duration exceeds $t_{lim}$.
+}
 

### 2016-01-18 paper.tex
Start adding new images-results, change Jacobian notation

@@
 \newcommand\po{\mathbf{P}}
-\newcommand\Jf{{J_f}}
+\newcommand\Jf{\Phi}
 \newcommand\kernel{\mbox{Ker }}
@@
 
-\textcolor{blue}{The constraint function is noted $f$}. Let us assume that at iteration $i$, $j$ constraints have been inserted before the current iteration. 
-These constraints are stored as lines of a Jacobian matrix as follows:
+Let us assume that at iteration $i$, $j$ \textcolor{blue}{linear} constraints have been inserted before the current iteration. 
+These constraints are stored as lines of a matrix as follows:
 $$
@@
 $$
+\textcolor{blue}{These linear constraints are built from the linearization of a differentiable function $f$, which will be detailed in the following section.}
 
@@
 $\po_j^l(\conf)$ be the coordinates in the frame of $\body_l$ of $\po_c$ expressed the frame of $\body_j$.
-Let $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) denote the homogeneous transformation 
-between frames of $\body_1$ (resp. $\body_2$) and the global frame.
+Let $M_1(\conf) \in SE(3)$ (respectively $M_2(\conf) \in SE(3)$) denote the homogeneous transformation 
+between frames of $\body_1$ (respectively $\body_2$) and the global frame.
 Thus, $\po_2^1(\conf) = M_{1}^{-1} (\conf) M_2 (\conf)$.
@@
 $$
-The linearization step is described in a further 
-section. Then, a line is added in the constraint Jacobian matrix (function 
-\texttt{addCollisionConstraint}), and the constraint 
-Jacobian becomes:
+The linearization step is described in the following 
+section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:
 \begin {align*}
@@
 \end{align*}
+\textcolor{blue}{This step is denoted by the function \texttt{addCollisionConstraint}}.
 
@@
 	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
-	and $\kappa$ resp. collision-free path, in-collision path, iteration step 
+	and $\kappa$ respectively collision-free path, in-collision path, iteration step 
 	and collision-abcissa).
@@
 
-\subsection{From 2D basic examples...}
+\subsection{From 2D basic examples}
 
@@
 TODO !!
+
+\begin{figure}[H]
+	\centering
+	\subfigure[Double-arm]{
+		\includegraphics[width=6.5cm,height=5.5cm]{2Darm_comparison_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:double_arm}
+	}
+	\subfigure[Double-arm]{
+		\includegraphics[width=6.5cm,height=5.5cm]{2Darm_comparison_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:double_arm}
+	}
+  \caption{Influence of $\alpha_{init}$.}
+  \label{fig:remainingLengthAlphaInit}
+\end{figure}
 
@@
 
+\begin{figure}[H]
+	\centering
+	\subfigure[Double-arm]{
+		\includegraphics[width=6.5cm,height=5.5cm]{ur2_comparison_RS_PRS_GB_sd.png}
+		\label{fig:remainingLengthComp:double_arm}
+	}
+	\subfigure[PR2-crossing-arms]{
+		\includegraphics[width=6.5cm,height=5.5cm]{PR2alone_comparison_RS_PRS_GB_sd.png}
+		\label{fig:remainingLengthComp:pr2alone}
+	}
+  \caption{Performance comparison between the three optimizers.}
+  \label{fig:remainingLength}
+\end{figure}
+
 Optimization computation time and path length averages are presented for 30 runs of 
@@
   highlights the optimization success of our method, which manages to deliver a 
-  shorter or similar path compared to RS output}
+  shorter or similar path compared to RS output.}
   \label{fig:trajectories}

### 2016-01-18 paper.tex
Use Linearly Constrained Quadratic Problem instead of Quadratic Program
  QP -> LCQP

@@
 preserves that a solution can still be found, e.g. for narrow passages as holes or 
-grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a quadratic program.
+grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{linearly constrained} quadratic program.
 
@@
 \textcolor{blue}{This section describes the settling of the Gradient-based optimizer.
-The method works as a classical Quadratic Program (QP), reducing the path length with a cost function and avoid collisions with constraints.
-Details of the QP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}
+The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length with a cost function and avoid collisions with constraints.
+Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}
 
@@
 Finally, we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
-programs (QP). \textcolor{blue}{The QP step computation corresponds to a constrained 
+programs (\textcolor{blue}{LC}QP). \textcolor{blue}{The step computation corresponds to a constrained 
 version of \texttt{computeIterate}}.
@@
 according to the previous steps definitions.
-\textcolor{blue}{Remind that, the QP optimal step $\p$ computed by \texttt{computeIterate} is known.
-The idea of the algorithm is to process iterations that reduce the path length according to the QP cost.
-If collision-constraints have been added to the QP, further iterations will comply with them.
-The algorithm stops whether the QP minimum is reached and collision-free.
+\textcolor{blue}{Remind that, the LCQP optimal step $\p$ computed by \texttt{computeIterate} is known.
+The idea of the algorithm is to process iterations that reduce the path length according to the LCQP cost.
+If collision-constraints have been added to the LCQP, further iterations will comply with them.
+The algorithm stops whether the LCQP minimum is reached and collision-free.
 The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.
@@
 constraints of GB is significant and $\alpha_{init}$ reduced. 
-In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimizer spends time on the QP resolution to avoid collisions.
+In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimizer spends time on the \textcolor{blue}{LC}QP resolution to avoid collisions.
 On the other hand, our method almost always provides the shortest paths, whereas it 
@@
 numerical tools as collision checking, linearized one-dimensional constraint
-and QP resolution. It correlates them in a 
+and \textcolor{blue}{LC}QP resolution. It correlates them in a 
 simple but effective way, and the algorithm structure is organized so that its 

### 2016-01-18 paper.tex
State that we optimize on Riemanian manifolds in a proper way.

@@
 grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{linearly constrained} quadratic program.
+
+\textcolor{blue}{Another important feature of our contribution is that we optimize paths on the robot configuration space in a proper mathematical way.
+Most other optimization algorithms represent $SO3$ rotations by a vector directed along the rotation axis and the norm of which is the rotation angle, also known as the exponential map of $so3$, or even worse by Euler angles.}
 

### 2016-01-20 paper.tex
Add figure explaining the algorithm.

@@
 \subsection{Convergence Analysis}
-For iterative algorithms, the risk of falling in an infinite loop 
-that keeps adding a redundant constraint exists.
-In this section, we give some assumptions under which our 
-algorithm will converge.
+
+\begin {figure}
+  \centerline {
+    \def\svgwidth {.7\linewidth}
+    \graphicspath{{./images/}}
+    \input {images/convergence.pdf_tex}
+  }
+  \caption {Illustration of linearized constraints: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2$ under the non-linear constraint $f(\xx)\leq 0$.
+$\alpha_{init}=\frac{1}{4}$.
+The algorithm starts from $\xx_0$.
+The first iterate is $\xx_1$ which satisfies the constraint.
+The second iterate is $\xx_2$ that does not satisfy the constraint.
+The algorithm backtracks to $\xx_1$ and inserts linear constraint $L_1: \frac{\partial f}{\partial \xx} (\xx_1)(\xx - \xx_1) = 0$.
+$\xx_3$ is the global minimum under $L_1$.
+As $\xx_3$ does not satisfy the constraint, the algorithm moves to $\xx_4$ that satisfies the constraints, and then to $\xx_5$ that does not.
+The algorithms backtracks to $\xx_4$, inserts constraint $L_2: \frac{\partial f}{\partial \xx} (\xx_2)(\xx - \xx_2) = 0$, and returns $\xx_4$ as a solution since the dimension of the search space is 0.
+Notice that the same non-linear constraint may give rise to several linear constraints.
+The convergence of the algorithm relies on the fact that the kernel of constraint $L2$ is not contained in the kernel of the current constraints ($L_1$ only here).
+The convergence analysis can be roughly summarized as follows.
+$L2$ to be linearly dependent of $L_1$ requires that $f$ is stationary along $L_1$ at $\xx_4$.
+This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$.
+If $L_2$ was not linearly independent from $L_1$, Algorithm~\ref{algo:solveRedundant} would keep searching new iterates between $\xx_5$ and $\xx_4$.
+If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$.
+By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary.
+In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom left picture).
+This is possible but unlikely, unless the problem has been defined as such on purpose.}
+  \label {fig:convergence}
+\end{figure}
+
+Although linearized constraints may differ from the initial geometrically relevant non-linear constraint when the iterate goes away from the linearization path, we will show in this section that  our algorithm converges under some reasonable assumptions.
+The underlying idea of the proof is sketched in Figure~\ref{fig:convergence}.
+
 From the definition of $\Jf_{i}$, it is straightforward that:

### 2016-01-20 paper.tex
Modify definition of constraints.

@@
 \newcommand\U{\mathbf{u}}
+\newcommand\rotation{R}
 
@@
 
-\subsection{Constraints}
+\subsection{Linear Constraints}
 
@@
 
-The principle of the method is to compute a constraint on the backtracked 
+The principle of the method is to compute a \textcolor{blue}{linear} constraint on the initialized on the 
 free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
@@
 (Figure~\ref{contact-points} left).
+\textcolor{red}{We denote by
+  \begin{itemize}
+  \item $\po_1$ (resp. $\po_2$) the point belonging to body $\body_1$ (resp. $\body_2$) that coincides with $\po_c$ in configuration $\xx_{i+1}(\tcolli)$,
+  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
+  \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
+  \end{itemize}
+  The position of $\po_1$ (resp. $\po_2$) in the global reference frame depends on the configuration $\conf$ of the robot. We therefore denote $\po_1(\conf)$ (resp. $\po_2(\conf)$) this position.
+We define $\U$ as the unit vector attached to $\body_1$, linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$ (Figure~\ref{contact-points} right):
+$$
+\U (\conf) = \rotation_1(\conf)\rotation_1^{T}(\xx_{i}(\tcolli))\frac{\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))}{\|\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))\|}
+$$
+Note that
+\begin{itemize}
+\item $\U$ depends on the configuration of the robot $\conf$, since $\U$ moves with $\body_1$,
+\item $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
+\end{itemize}
+}
 \textcolor{blue}{
-In any configuration $\conf$, let 
-$\po_j^l(\conf)$ be the coordinates in the frame of $\body_l$ of $\po_c$ expressed the frame of $\body_j$.
-Let $M_1(\conf) \in SE(3)$ (respectively $M_2(\conf) \in SE(3)$) denote the homogeneous transformation 
-between frames of $\body_1$ (respectively $\body_2$) and the global frame.
-Thus, $\po_2^1(\conf) = M_{1}^{-1} (\conf) M_2 (\conf)$.
-}
-
-\textcolor{blue}{
-In configuration $\xx_{i}(\tcolli)$, let $\mathbf{u}$ be the unit vector 
-linking the points $\po_1$ and $\po_2$ (Figure~\ref{contact-points} right):
-$$
-\U = \frac{\po_2^1(\xx_{i}(\tcolli)) - \po_1^1(\xx_{i}(\tcolli))}{\|\po_2^1(\xx_{i}(\tcolli)) - \po_1^1(\xx_{i}(\tcolli))\|}
-$$
-}
-
-Notice that $\mathbf{u}$ is well defined since configuration 
-$\xx_{i}(\tcolli)$ is collision-free.
-\textcolor{blue}{
-Let $g$ be the function representing the projection of $\overrightarrow{\po_1(\conf)\,\po_2(\conf)}$ on $\mathbf{u}$. $g$ is defined for any $\conf \in \CS$ to $\real$ by:
-$$
-g (\conf) = \left(\po_2^1(\conf) - \po_1^1(\conf) | \mathbf{u}\right)
+Let $g$ be the function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
+$$
+g (\conf) = \left(\po_2(\conf) - \po_1(\conf) | \mathbf{u}\right)
 $$
@@
 $$
-The constraint is defined for any path $\xx$ by:
+The constraint defined for any path $\xx$ by:
 \begin {equation}\label{eq:new-constraint}
@@
 \end {equation}
-This constraint forces the configuration $\xx (\tcolli)$ to keep the same projection on $\mathbf{u}$ as in $\xx_i (\tcolli)$. Since this constraint is 
-non-linear, we linearize it around $\xx_{i}$: 
+aims at keeping colliding points $P_1$ and $P_2$ in configuration $\xx_{i+1} (\tcolli)$ apart in direction $\U$ defined by those points in configuration $\xx_{i} (\tcolli)$.
+We linearize the constraint around $\xx_{i}$: 
 $$
@@
 $$
-The linearization step is described in the following 
+The computation of the linearized constraint is described in the next
 section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:

### 2016-01-25 answers-to-reviewers.tex
Update answer letter

@@
 \title{A simple path-optimization method for motion planning}
-\author{M. Campana, F. Lamiraux and J.-P. Laumond}
+\author{Myl\`{e}ne Campana, Florent Lamiraux and Jean-Paul Laumond}
 \date{Answer to reviewers}
@@
 
-We have developed the meaning of optimality, and we refer to a recent paper ACM-Com by one of the author.
+It is true that other optimization criteria can be different of the path length. 
+Thus, relying on the reviewer comment, we add the mention of others. Concerning 
+the choice of the path length, it allows us to obtain a quadratic cost, better 
+conditioned for optimization techniques.
 
-Concerning the choice to optimize the path length, we... by choosing the length we get a quadratic criterion, which is easier blablabla...
+Finally, we have developed the meaning of optimality, and we refer to a 
+recent paper ACM-Com by one of the author.
 
@@
 
-As suggested, the reference has been changed to cite the journal paper.
+We thank the reviewer for the reference, [11] has been changed to cite the journal paper instead of the conference.
 
@@
 
-We understand that starting to define the algorithm's functions and reference to it too early is confusing.
-To enhance it, we choose to decouple the algorithm from the problem definition, and define the functions in the algorithm section.
-This method may seem redundant but avoid the reader to flip back and forth to find the functions definitions.
-Furthermore, we prefer to not introduce the algorithm earlier because .... \todo
+We understand that starting to define the algorithm's functions and reference 
+to it too early is confusing.
+We prefer not to introduce the algorithm earlier because its structure ensue 
+from the problem's construction.
+Instead, we prefer to remove the reference to the algorithm. We also precise at 
+the beginning of the part 3 that some of the problem steps correspond to 
+functions that will be called in the algorithm, given at the end. Thus we do not need to redefine completely the functions in the algorithm section.
 
@@
 
-ANSWER
+We have rewritten the notations of this part to clarify it.
 
@@
 
-ANSWER
-It is true that in the case of Figure 6, Prune is complementary to random shortcut, but this method, used alone, has also it's counter-examples, as the one presented in Figure 1 (and the PR2-crossing-arms Figure \todo).
+It is true that in the case of Figure 6, Prune is complementary to random shortcut.
+But this method, used alone, has also its counter-examples, as the one presented in Figure 1 (and the PR2-crossing-arms Figure 9).
 
@@
 
-ANSWER
+We admit that the termination conditions of the algorithms were not fair.
+Instead, as suggested reviewers 1 and 3, we choose to allow the same amount of optimization duration for the three optimizers.
+Since the Gradient-Based optimizer is deterministic, its duration has been chosen.
+
+To provide information about the ``path length decreases as a function of time'', we have added Figure 13 some graphs comparing the three optimizers.
 
@@
 
-ANSWER
+We thank the reviewer for the suggestion of using Prune as a preliminary optimization step for each optimizer.
+Regarding to our algorithm, Prune has pros and cons: its removes waypoints i.e.
+optimization variables so the computation time reduces.
+However, less waypoints results in less possible actions to blend the path.
+In fact, we noted that we obtained shorter paths without the pruning stage.
+
+\todo : voir si ajout d'une partie "comparaison avec et sans prune" ....
+globalement, prune ne change pas grand chose a la comparaison avec les autres optim, si ce n'est qu'on a en moyenne des chemins initiaux similaires (peu de waypoints)
 
@@
 
-[1] Campana M, Lamiraux F, Laumond JP. A simple path optimization method for motion planning. https:
-//hal.archives-ouvertes.fr/hal-01220048. 2015.
+[1] Campana M, Lamiraux F, Laumond JP. A simple path optimization method for 
+motion planning. https://hal.archives-ouvertes.fr/hal-01220048. 2015.
 
@@
 
-We removed the reference.
+We thank the reviewer for the comment. We removed the reference from the paper.
 
@@
 
-ANSWER
+We thank the reviewer for the remark. We have improved the notations of the problem, avoiding to introduce unreferenced objects at the beginning.
+
+\todo : adapter / etoffer en fonction de ce que Florent a re-modifie
 
@@
 
-ANSWER
+We think that the reviewer is referring to line-search algorithms to 
+evaluate alpha. However these algorithms are not relevant for a QP problem since 
+the cost minimum is known.
+We just intend to process \textit{small} steps toward it, and to note collision-constraints.
+We admit that the alpha value could be chosen in a smarter way, e.g.
+due to geometrical considerations, but we mentioned it in the conclusion as a future work.
+
+\todo : verifier pertinence
 
@@
 
-ANSWER
+We want to highlight the fact that we optimize paths on the robot 
+configuration space in a proper mathematical way, specially regarding to the $SO3$ rotations. Thus we prefer to briefly define notations and operators.
 
@@
 
-ANSWER
+We apologize for the unclear/unreferenced notations and have improved it.
 
@@
 
-ANSWER
-%The authors agree with the reviewers that this part should be modified as it is unclear.
+The authors agree with the reviewer that this part should be modified as it is unclear. We add the Figure 5 to help understanding the interest of the 3.6 part.
+
+\todo : adapter en fonction de ce que Florent a re-modifie
 
@@
 
-ANSWER
+The unfair comparison, also raised by the reviewer 1, has been improved: we allocate the Gradient-Based duration as the computation times for the random optimizers.
+Thus we avoid arbitrary termination condition.
+
+The drawback of this method is that we ignore if the computation time of our method is too important.
+Thus, as suggested, we have represented the convergence curves for some benchmarks, to see if RS and PRS are not converging far rapidly than GB.
 
@@
 
-ANSWER
+As suggested, we have added at the end of the results some graphs for different values of $\alpha_{init}$.
 

### 2016-01-25 paper.tex
Rewrite sections 3.5, 3.6.

@@
 \newcommand\rotation{R}
-
+\newcommand\traj{\Gamma}
+\newcommand\velocity{\mathbf{v}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection{New constraint}
+\subsubsection{New constraint}\label{subsubsec:new-constraint}
 
@@
 	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
-	Collision Library~\cite{fcl}, used to detect collisions. 
-%	The red $(\perp)$ 
-%	hyperplan is orthogonal to ($\po_2$,$\mathbf{u}$) and represents the motion 
-%	direction which is allowed by the constraint.
+	Collision Library~\cite{fcl}, used to detect collisions. Constraint $f$ defined by~(\ref{eq:f}) aims at keeping $\po_2$ in plane $\Pi$ fixed to $\body_1$.
 	}
@@
   \begin{itemize}
-  \item $\po_1$ (resp. $\po_2$) the point belonging to body $\body_1$ (resp. $\body_2$) that coincides with $\po_c$ in configuration $\xx_{i+1}(\tcolli)$,
+  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
   \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
-  \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
+  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
+  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
+  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
   \end{itemize}
-  The position of $\po_1$ (resp. $\po_2$) in the global reference frame depends on the configuration $\conf$ of the robot. We therefore denote $\po_1(\conf)$ (resp. $\po_2(\conf)$) this position.
-We define $\U$ as the unit vector attached to $\body_1$, linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$ (Figure~\ref{contact-points} right):
-$$
-\U (\conf) = \rotation_1(\conf)\rotation_1^{T}(\xx_{i}(\tcolli))\frac{\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))}{\|\po_2(\xx_{i}(\tcolli)) - \po_1(\xx_{i}(\tcolli))\|}
-$$
-Note that
-\begin{itemize}
-\item $\U$ depends on the configuration of the robot $\conf$, since $\U$ moves with $\body_1$,
-\item $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
-\end{itemize}
+We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
+$$
+\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
+$$
+Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
 }
 \textcolor{blue}{
-Let $g$ be the function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
-$$
-g (\conf) = \left(\po_2(\conf) - \po_1(\conf) | \mathbf{u}\right)
-$$
+  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
+\begin{equation}\label{eq:g}
+g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
+\end{equation}
 }
@@
 Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
-$$
+\begin{equation}\label{eq:f}
 f (\xx) = g(\xx (\tcolli))
-$$
-The constraint defined for any path $\xx$ by:
+\end{equation}
+\textcolor{red} {The constraint defined for any path $\xx$ by:
 \begin {equation}\label{eq:new-constraint}
@@
 \end {equation}
-aims at keeping colliding points $P_1$ and $P_2$ in configuration $\xx_{i+1} (\tcolli)$ apart in direction $\U$ defined by those points in configuration $\xx_{i} (\tcolli)$.
+aims at keeping point $\po_2$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
+}
 We linearize the constraint around $\xx_{i}$: 
@@
 \end{align*}
-\textcolor{blue}{This step is denoted by the function \texttt{addCollisionConstraint}}.
+\textcolor{blue}{This step is performed by function \texttt{addCollisionConstraint}}.
 
@@
 blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in 
-each of the two waypoints and $\beta$. This stage \textcolor{blue}{is refered as 
-to \texttt{computeCollisionConstraint}}.
-
-\subsection{Convergence Analysis}
+each of the two waypoints and $\beta$. This stage \textcolor{blue}{is performed by \texttt{computeCollisionConstraint}}.
+
+\subsection{Convergence Analysis and algorithm refinment}
 
@@
   }
-  \caption {Illustration of linearized constraints: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2$ under the non-linear constraint $f(\xx)\leq 0$.
+  \caption {\textcolor{red}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$.
 $\alpha_{init}=\frac{1}{4}$.
@@
 This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$.
-If $L_2$ was not linearly independent from $L_1$, Algorithm~\ref{algo:solveRedundant} would keep searching new iterates between $\xx_5$ and $\xx_4$.
+If $L_2$ was not linearly independent from $L_1$, The algorithm would keep searching new iterates between $\xx_5$ and $\xx_4$.
 If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$.
 By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary.
-In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom left picture).
-This is possible but unlikely, unless the problem has been defined as such on purpose.}
+In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom right picture).
+This is possible but unlikely, unless the problem has been defined as such on purpose.}}
   \label {fig:convergence}
@@
 \end{equation}
-\textcolor{blue}{In other words, any iteration step $\p_i$ that complies with the 
-collision constraints on path $\xx_{i+1}$, necessarily sastifies $\xx_i$ constraints.}
 Let us assume that:
 \begin{equation}\label{eq:assumption-1}
-L_{k+1}\p_i \not= 0
+L_{j+1}\p_i \not= 0
 \end{equation}
@@
 $$
-This result proves that under assumption~(\ref{eq:assumption-1}), each added constraint 
-is linearly independent from the 
+This result proves that under assumption~(\ref{eq:assumption-1}), each additional constraint is linearly independent from the 
 previous ones, and thus our algorithm terminates in a finite number of iterations.
@@
 \noindent
-Let us consider the following trajectory:
+\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
+According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
+\begin{equation}\label{eq:f-along-pi}
+f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
+\end{equation}
+The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
 \begin{equation}\label{eq:trajectory}
-\conf (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
+\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
 \end{equation}
-\textcolor{blue}{where $\xx_i$ is still collision-free and $\xx_{i+1}$ 
-has its first collision detected at abcissa $\tcolli$.}
-The motion $\conf(t)$ is defined by taking a constant abscissa $\tcolli$ and by 
-moving from the 
-path $\xx_i$ along the step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
-
-Thus the velocity of point $\po_2(\conf(t))$ in the reference frame of $\body_1$ projected onto $\mathbf{u}$ along the motion of the robot is given by:
-$$
- \dot{\conf}(t) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_i + t\alpha_i\xx_{i+1})\p_i
-$$
-
+The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
+moving from 
+path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
+Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
+\begin{align}
+f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
+\label{eq:df/dx}
+\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
+\frac{d}{dt}g (\traj (t))
+\end{align}
+From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
+}
+
+\textcolor{red}{
 Therefore the following expressions
-% \frac{\partial \conf}{\partial t}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \text{and} \ \frac{\partial \conf}{\partial t}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
-$$
-\dot{\conf}(0) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i})\p_i \ \ \text{and} \ \ \dot{\conf}(1) = \alpha_i \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
-$$
-correspond to $(\dot{\po_2}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$. The term $\dot{\conf}(1)$ being equal to zero requires either:
-\begin {enumerate}
-\item that $\po_1$ and $\po_2$ come to contact at zero velocity, or
-\item the velocity of $\po_2$ in the reference frame of $\body_1$ is orthogonal to 
-$\mathbf{u}$ at contact ($t=1$).
-\end {enumerate}
-The first case is very unlikely and almost never encountered in real applications, 
-except if the user builds such a case on purpose.
-The second case may appear and can be noted by a rank loss of $\Jf_{i+1}$.
-Then we do not directly insert the new constraint. Instead, we process intermediate 
-iterations until the redundancy vanishes, as explained in function 
-\texttt{findNewConstraint} in Algorithm~\ref{algo:solveRedundant}, and illustrated in Figure~\ref{fig:convergence-diagram}.
-
-After producing an intermediate iteration $\xx_{i+2}$ where $\alpha_i$ has
-been halved, the obtained path is tested for collisions. If it is collision-free, the 
-constraint is now linearized around it. Otherwise, new contact points are taken 
-on it, and the constraint is linearized around the latest free path.
+\begin{align*}
+\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
+\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
+\end{align*}
+correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$.
+Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0.
+According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$.
+Although very unlikely, this case might appear for some new constraint.
+In this case, inserting constraint $L_{j+1}$ is useless since
+$$
+\kernel \Jf_i = \kernel \Jf_{i+1}.
+$$
+}
+
+\subsubsection {Algorithm refinment}
+
+\textcolor{red}{
+When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent.
+This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}.
+The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy.
+A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop.
+New iterations are chosen in the middle of this pair.
+\begin{itemize}
+\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
+\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
+\end{itemize}
+In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
+\begin{enumerate}
+\item at some point in the loop the new constraint is linearly independent from the previous constraints.
+The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
+\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
+\end{enumerate}
+In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$.
+$\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$.
+Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
+}
 
@@
 	Note that a case with 
-	$(\dot{\po_2}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
+	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
 	with the constraint that should be added if no redundancy was tracked. Therefore, 
@@
 	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
-	$(\dot{\po_2}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
+	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
 	returning the new constraint.}

### 2016-01-25 paper.tex
Update results

--- 2016-01-20 paper.tex
@@
 }
+\usepackage{minibox}
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
 
@@
 \textcolor{blue}{
-The main difficulty is to handle 
+One main difficulty is to handle 
 the scalar parameter $\alpha$ determining how much of the computed step  will be traveled along. 
@@
 \subsection{To 3D complex problems}
-
-\subsubsection{Analysis of $\alpha_{init}$ influence}
-\textcolor{blue}{Results without comparison to RS and PRS here
-$\alpha_{init}$ could be computed from geometrical considerations.
-}
-TODO !!
-
-\begin{figure}[H]
-	\centering
-	\subfigure[Double-arm]{
-		\includegraphics[width=6.5cm,height=5.5cm]{2Darm_comparison_alphaInit.png}
-		\label{fig:remainingLengthAlphaInit:double_arm}
-	}
-	\subfigure[Double-arm]{
-		\includegraphics[width=6.5cm,height=5.5cm]{2Darm_comparison_alphaInit.png}
-		\label{fig:remainingLengthAlphaInit:double_arm}
-	}
-  \caption{Influence of $\alpha_{init}$.}
-  \label{fig:remainingLengthAlphaInit}
-\end{figure}
 
@@
 
+Our algorithm is also experimented on more complex robots and environments.
+In the included video and Figures~\ref{pr2_final},~\ref{fiad} 
+and~\ref{fig:trajectories}, we present multiple situations where the GB algorithm 
+is tested and compared to RS  and PRS.
+\textcolor{blue}{
+After describing the random optimizers, we will present each benchmark and its qualitative path results. Then quantitative averages and graphs will be given and discussed.
+}
+
+\textcolor{blue}{
+The RS (see Algorithm~\ref{algo:random-shortcut}) shortens the path by randomly sampling configurations along it and trying to link them with collision-free interpolations.
+The termination condition of RS is a duration time limit $t_{lim}$, and is typically set as the GB optimization duration.
+%For our implementation of PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{wp+1}$. % NOT ANYMORE
+The implementation of PRS is identical to~\cite{GeraertsIJRR07}:
+for a random DOF, configurations are sampled along the initial path as in 
+Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path 
+made of an interpolation on only 
+the current DOF and based on the previous subpath for other DOF. If this path is 
+collision-free, it is added to the final path as in RS. The process is stopped when the duration exceeds $t_{lim}$.
+}
+
+\vspace{0.4cm}
+% freeflyer puzzle
+Before entering the manipulator examples, the GB algorithm is analyzed 
+on a popular problem in the motion planning literature: a freeflyer puzzle, 
+corresponding to Figure~\ref{fig:trajectories:puzzle}. The 
+puzzle has to cross down the 
+obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
+contains detours above and below the obstacle, as well as small superfluous
+motions in the hole. 
+The GB optimizer manages to modify the puzzle trajectory even in the hole, but remains less efficient than RS and PRS. We explain this behavior by the fact that, ........
+%PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole.
+This behavior was also noticed and improved in the PRS implementation of~\cite{GeraertsIJRR07}, acting on a group of DOF.
+The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.
+
+\vspace{0.4cm}
+
+% double-arm:
+In the double-arm benchmark (4-DOF), 
+Figure~\ref{fig:trajectories:double_arm}, one arm has 
+to get around a cylinder obstacle while the 
+other arm stay in the same configuration. As expected, the initial path given by 
+RRT-connect activates both arms to solve the problem. Unlike RS, 
+the GB optimizer manages to cancel the rotation of the free-arm while optimizing 
+the first arm motion, creating collision-constraints with the cylinder obstacle.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qinit.png}
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qfinal.png}\\
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_real.png}
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad-trajectorires.png}
+	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
+	the Factory-in-a-Day project~\cite{factory-day-video}. A similar environment 	
+	has been created (top) to illustrate that our method can comply with an 
+	industrial problem, where initial and final configurations of the UR5 are 
+	constrained in boxes. End-effector trajectories 
+	are illustrated (bottom right): initial (RRT-connect) in red, RS 
+	optimization in blue and the GB optimization in green.}
+	\label{fiad}
+\end{figure}
+
+% UR5-spheres and fiad:
+Some problems are 
+involving a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
+gripper.
+In a relatively free environment represented in Figure~\ref{fig:trajectories:ur5}, 
+results from our method and RS are similar. 
+Note also that the end-effector trajectory is completely different from the initial 
+one: the robot is easily passing between the spheres, keeping its end-effector 
+above.
+For an UR5 working in a cluttered environment inspired by an industrial issue 
+Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
+to the result of RS and to what can be observed in 
+reality~\cite{factory-day-video}.
+
+% baxter:
+A problem involving a Baxter-like\footnote{A torso rotation was added and 
+the grippers were removed.} robot 
+manipulating in an office environment is presented in
+Figure~\ref{fig:trajectoriesbis:baxter}. The robot starts with its end-effectors 
+above the computer and has to 
+turn and reach the shelf. According to the left-hand trajectories quality (see the 
+video for the full motion) and path lengths, the GB optimizer provides 
+the smoothest motion.
+
+\vspace{0.4cm}
+
+In the three following high-DOF examples involving the PR2 robot, note that 
+results are better in terms of path quality, as a result of the parasite 
+DOF motions removal.
+
 \begin{figure}
@@
 
-We also experiment our algorithm on more complex robots and environments.
-In the included video and Figures~\ref{pr2_final},~\ref{fiad} 
-and~\ref{fig:trajectories}, we present multiple situations where the GB algorithm 
-is tested and compared to RS~\cite{randomShortcutHPP}
-Algorithm~\ref{algo:random-shortcut} and PRS~\cite{partialrandomShortcutHPP} 
-adapted from~\cite{GeraertsIJRR07}.
-\textcolor{blue}{
-The termination condition of RS allows it to try 
-shortening the path until the optimization duration overflows a time limit. 
-This corresponds to $t_{lim}$ in Algorithm~\ref{algo:random-shortcut} and is typically equal to the Gradient-Based optimization duration.
-%For our implementation of PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{wp+1}$. % NOT ANYMORE
-The implementation of PRS is identical to~\cite{GeraertsIJRR07}:
-for a random DOF, configurations are sampled as in 
-Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path made of 
-an interpolation on only 
-the current DOF and based on the previous subpath for other DOF. If this path is 
-collision-free, we replace it as in RS. As RS, the process is stopped when the duration exceeds $t_{lim}$.
-}
-
-\vspace{0.4cm}
-% freeflyer puzzle
-Before entering the manipulator examples, the GB algorithm is analyzed 
-on a popular problem in the motion planning literature: a freeflyer puzzle, 
-corresponding to Figure~\ref{fig:trajectories:puzzle}. The 
-puzzle has to cross down the 
-obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
-contains detours above and below the obstacle, as well as small superfluous
-motions in the hole. 
-Lengths of initial, GB optimized, and RS optimized paths are 
-respectively 23.4, 8.53 and 10.2. PRS fails to decrease the path length because in 
-this problem, all DOF are closely related to cross the hole. This behavior was also 
-noticed and improved in the PRS implementation of~\cite{GeraertsIJRR07}, 
-acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory 
-even in the hole, contrary to RS which simply copies the initial path when the 
-trajectory gets near of the obstacle. $\alpha_{init}$ is reduced to 0.05 to slightly modify the very constrained trajectory at each iteration.
-
-\vspace{0.4cm}
-
-% double-arm:
-In the double-arm situation, Figure~\ref{fig:trajectories:double_arm}, one arm has 
-to get around a cylinder obstacle while the 
-other arm is relatively far from obstacles. As expected, the initial path given by 
-RRT-connect activates both arms to solve the problem. Unlike RS, 
-the GB optimizer manages to cancel the rotation of the free-arm while optimizing the 
-first arm motion, even if the second arm is still moving because of the translation 
-motion. The initial path length 5.17 is decreased to 4.55 by RS, to 4.06 by PRS and 
-to 2.77 by the GB method using three collision-constraints with the cylinder obstacle.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qinit.png}
-	\includegraphics[width=5.8cm,height=4.7cm]{fiad_qfinal.png}\\
-	\includegraphics[width=5.8cm,height=4.7cm]{fiad_real.png}
-	\includegraphics[width=5.8cm,height=4.7cm]{fiad-trajectorires.png}
-	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
-	the Factory-in-a-Day project~\cite{factory-day-video}. A similar environment 	
-	has been created (top) to illustrate that our method can comply with an 
-	industrial problem, where initial and final configurations of the UR5 are 
-	constrained in boxes. End-effector trajectories 
-	are illustrated (bottom right): initial (RRT-connect) in red, RS 
-	optimization in blue and the GB optimization in green.}
-	\label{fiad}
-\end{figure}
-
-% UR5-spheres:
-Some problems are 
-involving a 6-axis manipulator arm, also called UR5, equipped with a bar or a 
-gripper.
-In a relatively free environment represented in Figure~\ref{fig:trajectories:ur5}, 
-results from our method and RS are similar (path length are respectively 2.03 and 1.60). 
-Note also that the end-effector trajectory is completely different from the initial 
-one: the robot is easily passing between the spheres, keeping its end-effector 
-above.
-
-% UR5-fiad
-For an UR5 working in a cluttered environment inspired by an industrial issue 
-Figure~\ref{fiad}, path optimization efficiently returns a shorter solution, close 
-to the result of RS and to what can be observed in reality~\cite{factory-day-video}. 
-The initial path length 6.65 is decreased to 2.08 by RS, to 
-1.98 by the GB optimizer and unchanged by PRS.
-
-% baxter:
-A problem involving a Baxter-like\footnote{A torso rotation was added.} robot 
-manipulating in an office environment is presented in
-Figure~\ref{fig:trajectories:baxter}. The robot starts with its end-effectors 
-above the computer and has to 
-turn and reach the shelf. According to the left-hand trajectories quality (see the 
-video for the full motion) and path lengths, 9.10 (initial RRT-connect), 7.52 (RS), 
-8.41 (PRS) and 2.70 (GB), the GB optimizer is the most successful in this case.
-
-\vspace{0.4cm}
-
-In the three following high-DOF examples involving the PR2 robot, note that 
-results are better in terms of path quality, as a result of the parasite 
-DOF motions removal.
-
 % analysis PR2 alone:
@@
 returns detours and activates non-useful DOF such as the head, the torso lift 
-and the translation on the ground. Such behavior induces a high initial path length 
-30.6. RS hardly optimizes the mobile base 
+and the translation on the ground. Such behavior induces a high initial path length. RS hardly optimizes the mobile base 
 translation (Figure~\ref{pr2_final} middle) of the robot and other unnecessary DOF 
-uses, resulting in a path length of 4.08.
-Whereas the GB optimized-path mainly 
-results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating 
-two collision-constraints between the arms. In the PRS 
-result, only presented in the video, the mobile base motion is canceled.
-However, this is \textcolor{blue}{not} the case for other unnecessary DOF motions.
-In addition, the arms motion is very wide and not as optimized as in the GB trajectory.
+uses. Whereas the GB optimized-path mainly 
+results in moving the arms as expected (Figure~\ref{pr2_final} right), just creating two collision-constraints between the arms. 
+\textcolor{blue}{
+In the PRS result, only presented in the video, the motion is less optimized than with RS: the arms are moving widely and the mobile base remains activated.
+One solution to remove such unnecessary DOF activation can be to try applying a partial shortcut on each DOF between the initial and final configuration.
+It appears that this step is more costly in terms of computation time than the GB duration, therefore it cannot be afforded by our PRS implementation.
+However, it could be applied as a preliminary optimization for each optimizer.
+}
 
 % analysis PR2 in kitchen:
-We obtain similar results on the PR2 performing manipulation tasks 
+Similar results are obtained on the PR2 performing manipulation tasks 
 in a kitchen environment.
 Firstly, the robot moves its hands from the top to the bottom of a table.
-The different trajectories of the right gripper are indicated in Figure~\ref{fig:trajectories:kitchen1}.
-Our optimizer manages to reduce the initial path length 12.7 from visibility-PRM 
+The different trajectories of the right gripper are indicated in Figure~\ref{fig:trajectoriesbis:kitchen1}.
+Our optimizer manages to reduce the initial path length from visibility-PRM 
 and improves the path quality 
@@
 backward and uses its arms DOF to avoid the table, instead of 
-processing a large motion to get away from the table. With the GB algorithm, the path length is 
-reduced to 3.15, against 5.47 for RS.
+processing a large motion to get away from the table.
 Secondly, another example of PR2 going from the set to the fridge 
-door is presented Figure~\ref{fig:trajectories:kitchen2} with the mobile base
+door is presented Figure~\ref{fig:trajectoriesbis:kitchen2} with the mobile base
 trajectories. Here, GB and RS results are similar in terms of length and rendering.
@@
 \vspace{0.4cm}
+
+\textcolor{blue}{
+For some of the presented benchmarks, convergence graphs of the path 
+length reduction are 
+given Figure~\ref{fig:remainingLength}. The chosen initial paths are unchanged, 
+e.g. correspond to Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis} . 
+Each graph illustrates the percent ratio of the optimized path length over the initial path length, during the optimization time.
+For the random optimizers curves, averages and standard deviations on 50 
+launches are given. It is not a surprise that GB is globally slower than RS which converges faster. 
+}
+
+\begin{table}
+\renewcommand{\arraystretch}{1.3}
+\centering
+\scalebox{0.9}{
+  \begin{tabular}{cccC{1.3cm}C{1.3cm}C{1.3cm}}
+  \toprule
+  \multirow{2}{*}{\textbf{Problem}} &  & \multirow{2}{*}{Computation time} & \multicolumn{3}{c}{\minibox{Relative remaining length (\%)}} \\
+   & & &  GB  &  RS  &  PRS  \\
+  \midrule
+  \midrule
+  %\textbf{2D slidding-robot (Fig.~\ref{2D_long} right)} & & 11.9 ms & 78.3 & 68.0 & 81.8 \\
+  \textbf{Freeflyer-puzzle} & ($\alpha_{init}=0.1$) & 0 ms & 0 & 0 & 0 \\
+  \textbf{Double-arm 4-DOF} & & 39.6 ms & 46.8 & 55.5 & 63.5 \\
+  \textbf{UR5-with-spheres} & (RRT, $\alpha_{init}=0.3$) & 3.18 s & 53.8 & 47.5 & 67.8 \\
+  \textbf{UR5-industrial-example} & (RRT) & 0 ms & 0 & 0 & 0 \\
+  \textbf{Baxter-in-office} & & 0 ms & 0 & 0 & 0 \\
+  \textbf{PR2-crossing-arms} & & 882 ms & 19.9 & 43.2 & 95.2 \\
+  \textbf{PR2-in-kitchen-1} & & 0 ms & 0 & 0 & 0 \\
+  \bottomrule
+  \end{tabular}  }%scalebox
+\caption{Average results for 50 runs of several examples presented in the 
+paper. For each run, a solution path is planned by Visibility-PRM 
+(unless `RRT' for RRT-connect is mentioned) as initial guess for the 
+three optimizers. The RS and PRS results correspond to averages of 50 
+launches of the random optimizers on each initial guess. The GB 
+computation time is the work duration allowed for the random optimizers. 
+$\alpha_{init}=0.2$ unless another value is specified.
+%For the PR2-crossing-arms example only, the right column represents the distance that is traveled by the robot mobile base during the different paths.
+}
+\label{tab:results}
+\end{table}
+
+\textcolor{blue}{
+Since the GB optimizer results depend on the shape of the initial guess, e.g.
+the number of waypoints and the proximity to obstacles, results averages for 50 initial paths of each benchmark are presented in Table~\ref{tab:results}.
+As mentioned, the paths are obtained from Visibility-PRM or RRT-connect.
+Due to their nature, these motion planners provide different types of path: the output of Visibility-PRM contains less waypoints and does not tend to be close to the contact, behavior induced by the extension process of RRT.
+}
+
+In some cases, $\alpha_{init}$ was 
+reduced to comply with very narrow passages, or increased in the opposite case.
+\textcolor{blue}{
+The following section deals with the $\alpha_{init}$ influence in depth.
+}
+Except the low-DOF problem of the puzzle, our method almost always provides the 
+shortest paths, whereas it is not even the best criterion to quantify the parasite 
+motions disappearance.
+TO COMPLETE!
+%Finally, the complete vanishing of the mobile base motion in the GB optimized path is noted. 
+%Thus the issue addressed in Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
+
+\subsubsection{Analysis of $\alpha_{init}$ influence}
+\textcolor{blue}{Results without comparison to RS and PRS here
+$\alpha_{init}$ could be computed from geometrical considerations.
+}
+TODO !!
+Contrary to a first idea that $\alpha_{init}$ is reduced to 0.05 to slightly modify the very constrained trajectory at each iteration.
 
@@
 	\centering
+	\subfigure[2D slidding-robot (Fig.~\ref{2D_long} right)]{
+		\includegraphics[width=6.8cm,height=5.5cm]{potential2D_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:2Drobot}
+	}
 	\subfigure[Double-arm]{
-		\includegraphics[width=6.5cm,height=5.5cm]{ur2_comparison_RS_PRS_GB_sd.png}
-		\label{fig:remainingLengthComp:double_arm}
-	}
-	\subfigure[PR2-crossing-arms]{
-		\includegraphics[width=6.5cm,height=5.5cm]{PR2alone_comparison_RS_PRS_GB_sd.png}
-		\label{fig:remainingLengthComp:pr2alone}
-	}
-  \caption{Performance comparison between the three optimizers.}
-  \label{fig:remainingLength}
+		\includegraphics[width=6.8cm,height=5.5cm]{ur2_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:double_arm}
+	}
+  	\subfigure[Freeflyer-puzzle]{
+		\includegraphics[width=6.8cm,height=5.5cm]{puzzle_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:puzzle}
+	}
+	\subfigure[Baxter-in-office]{
+		\includegraphics[width=6.8cm,height=5.5cm]{baxter_alphaInit.png}
+		\label{fig:remainingLengthAlphaInit:baxter}
+	}
+  \caption{Influence of $\alpha_{init}$ on the convergence graphs of the GB optimizer.}
+  \label{fig:remainingLengthAlphaInit}
 \end{figure}
-
-Optimization computation time and path length averages are presented for 30 runs of 
-each example in (Table~\ref{tab:results}). In some cases, $\alpha_{init}$ was 
-reduced to comply with very narrow passages, or increased in the opposite case. 
-On the one hand, for non-high DOF examples, it appears that computation time is higher using the GB 
-optimization. The worst case comes out for the puzzle problem where the number of 
-constraints of GB is significant and $\alpha_{init}$ reduced. 
-In addition, collision checking is rapidly processed in such basic problem geometry. This favors RS while the GB optimizer spends time on the \textcolor{blue}{LC}QP resolution to avoid collisions.
-On the other hand, our method almost always provides the shortest paths, whereas it 
-is not even the best criterion to quantify the parasite motions disappearance.
-Note that these results are the consequence of the termination conditions of RS and PRS, balancing length reduction or shorter computation time. 
-This is particularly visible for PRS which 
-easily gives up in some problems, such as the 2D sliding-robot, the 
-UR5-with-spheres or the puzzle.
-Finally, the complete vanishing of the mobile base motion in the PRS and GB 
-optimized paths is highlighted in the right column of (Table~\ref{tab:results}). 
-Thus the issue addressed in Figure~\ref{decoupled_DOF_optimization} is 
-solved considering this DOF.
 
@@
 	\subfigure[Double-arm]{
-		\includegraphics[width=6.5cm,height=5.5cm]{ur2_orth_constr_results.png}
+		\includegraphics[width=2.9cm,height=5.5cm]{ur2_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{ur2_traj.png}
 		\label{fig:trajectories:double_arm}
@@
 	\subfigure[Freeflyer-puzzle]{
-		\includegraphics[width=6.5cm,height=5.5cm]{puzzle_orth_constr_results.png}
+		\includegraphics[width=2.9cm,height=5.5cm]{puzzle_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{puzzle_traj.png}
 		\label{fig:trajectories:puzzle}
 	}
-	\subfigure[PR2-in-kitchen-1]{
-		\includegraphics[width=6.5cm,height=5.5cm]{pr2-kitchen1_orth_constr_results.png}
-		\label{fig:trajectories:kitchen1}
-	}
 	\subfigure[UR5-with-spheres]{
-		\includegraphics[width=6.5cm,height=5.5cm]{ur5-spheres_orth_constr_results.png}
+		\includegraphics[width=2.9cm,height=5.5cm]{ur5spheres_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{ur5spheres_traj.png}
 		\label{fig:trajectories:ur5}
 	}
-	\subfigure[PR2-in-kitchen-2]{
-		\includegraphics[width=6.5cm,height=5.5cm]{pr2_kitchen2_orth_constr_results.png}
-		\label{fig:trajectories:kitchen2}
-	}
-	\subfigure[Baxter-in-office]{
-		\includegraphics[width=6.5cm,height=5.5cm]{baxter_orth_constr_results.png}
-		\label{fig:trajectories:baxter}
-	}
-  \caption{Trajectories of end-effectors, mobile bases or centers along the 
-  different paths: the initial path is represented in red, the RS optimized path in 
-  blue and the GB optimized path in green. The full robots 
-  motions can also be visualized in the joined video. The trajectories comparison 
-  highlights the optimization success of our method, which manages to deliver a 
-  shorter or similar path compared to RS output.}
+  \caption{(Left) Initial and final configurations.
+  (Right) Trajectories of end-effectors or centers along the 
+  different paths: the initial path is represented in red, the RS 
+  output in blue, the PRS output in cyan (top only) and the GB optimized path 
+  in green. The full robots 
+  motions can also be visualized in the joined video. The 
+  trajectories comparison highlights the optimization success of our method, 
+  which manages to deliver a shorter or similar path compared to the RS output. 
+  GB also cancels unecessary DOF activation contrary to RS and PRS (see the 
+  lower arm in top-right Figure).}
   \label{fig:trajectories}
@@
 
-\begin{table}
+\begin{figure}
 	\centering
-\scalebox{0.9}{
-  \begin{tabular}{ccccc}
-  \toprule
-    \textbf{Problem} & Path obtained from & Computation time  & Path length & Base traveled distance\\
-    \midrule
-    \midrule
-    & Initial &  & $21.9$  &\\
-    \textbf{2D slidding-robot} & GB & $9.29\,\text{ms}$ & $17.0$  &\\
-    \textbf{(Fig.~\ref{2D_long} right)} & PRS & $0.336\,\text{ms}$ & $21.7$  &\\
-    & RS & $1.89\,\text{ms}$ & $17.7$  &\\
-    \midrule
-    & Initial &  & $18.8$  &\\
-    \textbf{Freeflyer-} & GB ($\alpha_{init}=0.05$) & $1020\,\text{ms}$ & $9.65$  &\\
-    \textbf{puzzle} & PRS & $1.13\,\text{ms}$ & $18.8$  &\\
-    & RS & $4.72\,\text{ms}$ & $10.3$  &\\
-    \midrule
-    & Initial &  & $10.8$  &\\
-    \textbf{Double-arm} & GB & $10.3\,\text{ms}$ & $4.79$  &\\
-    \textbf{5-DOF} & PRS & $2.17\,\text{ms}$ & $8.08$  &\\
-    & RS & $5.33\,\text{ms}$ & $6.24$  &\\
-    \midrule
-    & Initial (RRT) &  & $4.97$ &\\
-    \textbf{UR5-} & GB ($\alpha_{init}=0.4$) & $2.12\,\text{s}$ & $2.97$ &\\
-    \textbf{with-spheres} & PRS & $0.884\,\text{s}$ & $4.56$ &\\
-    & RS & $2.89\,\text{s}$ & $2.99$ &\\
-    \midrule
-    & Initial &  & $4.83$ &\\
-    \textbf{UR5-} & GB ($\alpha_{init}=0.4$) & $314\,\text{ms}$ & $2.25$ &\\
-    \textbf{industrial-example} & PRS & $30.2\,\text{ms}$ & $4.53$ &\\
-    & RS &  $95.6\,\text{ms}$ & $2.16$ &\\
-    \midrule
-    & Initial &  & $20.2$ &\\
-    \textbf{Baxter-in-} & GB & $22.7\,\text{s}$ & $6.65$ &\\
-    \textbf{office} & PRS & $9.49\,\text{s}$ & $18.9$ &\\
-    & RS & $3.13\,\text{s}$ & $11.9$ &\\
-    \midrule
-    & Initial &  & $12.6$  & $9.57$\,\text{m}\\
-    \textbf{PR2-crossing-} & GB & $1.28\,\text{s}$ & $2.57$  & $0\,\text{m}$\\
-    \textbf{arms} & PRS & $2.63\,\text{s}$ & $ 5.31$ & $0\,\text{m}$\\
-    & RS & $1.92\,\text{s}$ & $3.59$ & $2.63\,\text{m}$\\
-    \midrule
-    & Initial &  & $15.3$ &\\
-    \textbf{PR2-in-} & GB & $19.2\,\text{s}$ & $4.53$ &\\
-    \textbf{kitchen-1} & PRS & $23.5\,\text{s}$ & $12.2$ &\\
-    & RS & $9.88\,\text{s}$ & $5.75$ &\\
-    \bottomrule
-  \end{tabular}
-  }%scalebox
-\caption{Average results for 30 runs of several examples presented in the paper and 
-in the video. For each run, a solution path is planned by Visibility-PRM (unless 
-`RRT' for RRT-connect is mentioned) as initial guess for each optimizer.
-For the PR2-crossing-arms example only, the right column represents the 
-distance that is traveled by the robot mobile base during the different paths.
-}
-\label{tab:results}
-\end{table}
+	\subfigure[PR2-in-kitchen-1]{
+		\includegraphics[width=2.9cm,height=5.5cm]{pr2kitchen1_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{pr2kitchen1_traj.png}
+		\label{fig:trajectoriesbis:kitchen1}
+	}
+	\subfigure[PR2-in-kitchen-2]{
+		\includegraphics[width=2.9cm,height=5.5cm]{pr2kitchen2_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{pr2kitchen2_traj.png}
+		\label{fig:trajectoriesbis:kitchen2}
+	}
+	\subfigure[Baxter-in-office]{
+		\includegraphics[width=2.9cm,height=5.5cm]{baxter_initfinal.png}
+		\includegraphics[width=6.5cm,height=5.5cm]{baxter_traj.png}
+		\label{fig:trajectoriesbis:baxter}
+	}
+  \caption{Other trajectories comparisons of end-effectors or mobile bases (initial path in red, RS output in blue and GB output in green).}
+  \label{fig:trajectoriesbis}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\subfigure[2D sliding robot (Fig.~\ref{2D_long})]{
+		\includegraphics[width=6.5cm,height=5.5cm]{potential2D_comp.png}
+		\label{fig:remainingLengthComp:potential2D}
+	}
+	\subfigure[Freeflyer-puzzle]{
+		\includegraphics[width=6.5cm,height=5.5cm]{puzzle_comp.png}
+		\label{fig:remainingLengthComp:puzzle}
+	}
+	\subfigure[UR5-with-spheres]{
+		\includegraphics[width=6.5cm,height=5.5cm]{ur5spheres_comp.png}
+		\label{fig:remainingLengthComp:ur5sphere}
+	}
+	\subfigure[PR2-crossing-arms]{
+		\includegraphics[width=6.5cm,height=5.5cm]{pr2crossing_comp_prune.png}
+		\label{fig:remainingLengthComp:pr2alone}
+	}
+	\subfigure[PR2-in-kitchen-1]{
+		\includegraphics[width=6.5cm,height=5.5cm]{pr2kitchen1_comp.png}
+		\label{fig:remainingLengthComp:pr2kitchen1}
+	}
+	\subfigure[Baxter-in-office]{
+		\includegraphics[width=6.5cm,height=5.5cm]{baxter_comp_prune.png}
+		\label{fig:remainingLengthComp:baxter}
+	}
+  \caption{Convergence graphs of the three optimizers during the optimization process.
+For each benchmark, the considered intial paths correspond to the ones of Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis}.
+The remaining path length relative to the initial one is represented.
+The dashed blue line is the final result of GB.
+For RS and PRS, the averages and standard deviations (in grey) over 50 launches are plotted.}
+  \label{fig:remainingLength}
+\end{figure}
 

### 2016-01-27 paper.tex
Finish Section 3.6

--- 2016-01-25 paper.tex
@@
 \end{enumerate}
-In the rest of this section, the first case will only be considered. Reasoning 
+In the rest of this section, the first case only will be considered. Reasoning 
 about the second case is similar, except that the 
@@
 
-The principle of the method is to compute a \textcolor{blue}{linear} constraint on the initialized on the 
-free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
+The principle of the method is to compute a \textcolor{blue}{linear} constraint initialized on the 
+free configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
 (\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
@@
   \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
-  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
+  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame, in configuration $\conf$,
   \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
@@
 
-\vspace{0.2cm}
-
 \subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
@@
 \end{align}
+\begin{property}\label{prop:geometric-reasoning}
 From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
+\end{property}
 }
@@
 Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
+$\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision.
+Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
+}
+
+\textcolor{red}{
+At each iteration, the new linear constraint
+$$
+L_{j+1} = \frac{\partial f}{\partial\xx} (\xx_{Free})
+$$
+is tested. As iterations $\xx_{Free}$ and $\xx_{Coll}$ tend toward $\bar{\xx}$, it can be established by a geometric reasoning analogous to Property~\ref{prop:geometric-reasoning}, that
+$L_{j+1} \p_i$ tends to the norm of the velocity of point $\bar{\po}$ belonging to $\body_2$ in the frame of $\body_1$. The following property summarizes the result of this section.
+}
+
+\textcolor{red}{
+\begin{property}\label{prop:convergence}
+As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
+\end{property}
+The above property means that the gradient-based path length optimization algorithm described in this paper converges in all cases, except for ill-defined problems.
 }

### 2016-01-27 paper.tex
Finish results part, fix Figures

--- 2016-01-25 paper.tex
@@
 \newcommand\velocity{\mathbf{v}}
+\newcommand\todo{\textcolor{red}{NOT FINISHED}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 addressed by our method.
+\textcolor{blue}{Processing a path 
+pruning~\cite{GeraertsIJRR07}, in order to remove redundant nodes from the initial 
+path, is a classical preliminary step for path length shortening. 
+A pruning will efficiently solve the Figure~\ref{local_box_optim} issue, however it 
+will fail tackling the Figure~\ref{decoupled_DOF_optimization} one, as RS.}
 
@@
 preserves that a solution can still be found, e.g. for narrow passages as holes or 
-grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{linearly constrained} quadratic program.
+grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{Linearly Constrained Quadratic Program (LCQP)}.
 
 \textcolor{blue}{Another important feature of our contribution is that we optimize paths on the robot configuration space in a proper mathematical way.
-Most other optimization algorithms represent $SO3$ rotations by a vector directed along the rotation axis and the norm of which is the rotation angle, also known as the exponential map of $so3$, or even worse by Euler angles.}
+Most other optimization algorithms represent $SO(3)$ rotations by a vector directed along the rotation axis and the norm of which is the rotation angle, also known as the exponential map of $SO(3)$, or even worse by Euler angles.}
 
@@
 
-\textcolor{blue}{This section describes the settling of the Gradient-based optimizer.
-The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length with a cost function and avoid collisions with constraints.
+\textcolor{blue}{This section describes the establishment of the Gradient-based optimizer.
+The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length expressed as a cost function and avoid collisions with linearized constraints.
 Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}
@@
 paths and to return the first colliding configuration 
-\textcolor{blue}{and its abcissa} along a path. \textcolor{blue}{This stage 
-is handled by a function \texttt{validatePath}, which returns $true$ if the given path is collision-free}.
+\textcolor{blue}{and its abcissa} along a path.
 
@@
 (Figure~\ref{contact-points} left).
-\textcolor{red}{We denote by
+\textcolor{blue}{We denote by
   \begin{itemize}
@@
 $$
-\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
+\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}\|}
 $$
@@
 \begin{equation}\label{eq:g}
-g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
+g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - \po_1^{loc} | \U\right)
 \end{equation}
@@
 \end{equation}
-\textcolor{red} {The constraint defined for any path $\xx$ by:
+\textcolor{blue} {The constraint defined for any path $\xx$ by:
 \begin {equation}\label{eq:new-constraint}
@@
 \end{align*}
-\textcolor{blue}{This step is performed by function \texttt{addCollisionConstraint}}.
+\textcolor{blue}{This stage is performed by a function \texttt{addCollisionConstraint}}.
 
@@
   }
-  \caption {\textcolor{red}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$.
+  \caption {\textcolor{blue}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$.
 $\alpha_{init}=\frac{1}{4}$.
@@
 \noindent
-\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
+\textcolor{blue}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
 According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
@@
 \end{align}
-From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
-}
-
-\textcolor{red}{
+From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $\po_2$ in reference frame $\body_1$ projected on vector $\U$.
+}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
+	\caption{\textcolor{blue}{Considering a trajectory 
+	$\traj(t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
+	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
+	and $\kappa$ are respectively collision-free path, in-collision path, 
+	iteration step and collision-abcissa).
+	The trajectory is represented in the configuration space (left) and in the 
+	workspace with the colliding bodies (right).}
+	Note that a case with 
+	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, 
+	so $\p$ already complies 
+	with the constraint that should be added if no redundancy was 
+	tracked. Therefore, 
+	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
+	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
+	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
+	returning the new constraint.}
+	\label{fig:convergence-diagram}
+\end{figure}
+
+\textcolor{blue}{
 Therefore the following expressions
@@
 \end{align*}
-correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$.
+correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $\po_2$ in reference frame $\body_1$.
 Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0.
@@
 
+\begin{algorithm}
+\begin{algorithmic}%[1] % line numerotation
+\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
+\OUTPUT{linearized constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$}
+\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
+\State $solved \gets false$
+\While{($\textbf{not}(solved)$)}
+	%\State $\p = \texttt{computeIterate}()$
+	\State $\alpha \gets 0.5\alpha$
+	\State $\xx \gets \xx_{Free} + \alpha\,\p$
+	\If{(\texttt{validatePath}($\xx$))}
+		\State $\xx_{Free} \gets \xx$
+	\Else
+		\State $\xx_{Coll} \gets \xx$
+	\EndIf
+	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
+	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
+\EndWhile
+\end{algorithmic}
+\caption{Description of \texttt{findNewConstraint}() which returns a linearly 
+independent constraint w.r.t. previous constraints stacked in $\Jf$.} 
+\label{algo:solveRedundant}
+\end{algorithm}
+
 \subsubsection {Algorithm refinment}
 
-\textcolor{red}{
+\textcolor{blue}{
 When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
-	\caption{\textcolor{blue}{Considering a trajectory 
-	$\conf (t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
-	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
-	and $\kappa$ respectively collision-free path, in-collision path, iteration step 
-	and collision-abcissa).
-	The trajectory is represented in the configuration space (left) and in the 
-	workspace with the colliding bodies (right).}
-	Note that a case with 
-	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
-	with the constraint that should be added if no redundancy was tracked. Therefore, 
-	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
-	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
-	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
-	returning the new constraint.}
-	\label{fig:convergence-diagram}
-\end{figure}
-
-\begin{algorithm}[H]
-\begin{algorithmic}%[1] % line numerotation
-\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
-\OUTPUT{linearized constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$}
-\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
-\State $solved \gets false$
-\While{($\textbf{not}(solved)$)}
-	%\State $\p = \texttt{computeIterate}()$
-	%\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
-	\State $\xx \gets \xx_{Free} + 0.5\alpha\,\p$
-	\If{(\texttt{validatePath}($\xx$))}
-		\State $\xx_{Free} \gets \xx$
-	\Else
-		\State $\xx_{Coll} \gets \xx$
-	\EndIf
-	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
-	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
-	\textcolor{blue}{\State $\alpha \gets 0.5\alpha$}
-\EndWhile
-\end{algorithmic}
-\caption{Description of \texttt{findNewConstraint}() which returns a linearly 
-independent constraint w.r.t. previous constraints stacked in $\Jf$.} 
-\label{algo:solveRedundant}
-\end{algorithm}
+\textcolor{red}{Partie en cours de modif par Florent}
 
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \begin{algorithmic}%[1] % line numerotation
@@
 
+Figure~\ref{2D_long} shows the result of our optimizer on 2D cases. Contact points which have 
+led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
+Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
+the constraint must be created, since at some parts the robot is very close to obstacles.
+
+Besides, these examples give a better understanding of how the tuning of 
+$\alpha_{init}$ 
+has to balance lot of iterations and relevant collision-constraint addition. For 
+example, if we alter a lot the initial path with a large gradient step and 
+compute the corresponding collisions, constraints will be chosen on a 
+path closer-to-optimality but which is not relevant
+w.r.t. the obstacles we want to avoid.
+
+Figure~\ref{local_box_optim} illustrates a \textit{very long} path example which RS 
+or PRS will not manage to 
+optimize in an affordable time, because of probabilistically failing to sample 
+configurations in the box. The GB method succeeds to optimize the 
+path contained in the box, with the following cost coefficients:
+$$
+\lambda_{k-1} = \frac{1}{\sqrt{(\conf_{k,0}-\conf_{k-1,0})^T \weight^2 
+(\conf_{k,0}-\conf_{k-1,0})}}, \  k=1\cdots wp+1  
+$$
+aiming at keeping the same ratio between path segment lengths at 
+minimum as at 
+initial path, represented by the waypoints ($\conf_{i,0})_{k=1\cdots wp+1}$.
+Without these coefficients, the path that minimizes the cost corresponds to a 
+straight line with the waypoints equidistantly allocated. This is not relevant for 
+Figure~\ref{local_box_optim} type of problems where a local passage is very
+constrained by obstacles. Note that this cost is also working with all other 
+examples presented in this paper, and provides better quality results than the original cost.
+
 \begin{figure}
@@
 \end{figure}
-
-Figure~\ref{2D_long} shows the result of our optimizer on 2D cases. Contact points which have 
-led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
-Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
-the constraint must be created, since at some parts the robot is very close to obstacles.
-
-Besides, these examples give a better understanding of how the tuning of 
-$\alpha_{init}$ 
-has to balance lot of iterations and relevant collision-constraint addition. For 
-example, if we alter a lot the initial path with a large gradient step and 
-compute the corresponding collisions, constraints will be chosen on a 
-path closer-to-optimality but which is not relevant
-w.r.t. the obstacles we want to avoid.
-
-Figure~\ref{local_box_optim} illustrates a \textit{very long} path example which RS 
-or PRS will not manage to 
-optimize in an affordable time, because of probabilistically failing to sample 
-configurations in the box. The GB method succeeds to optimize the 
-path contained in the box, with the following cost coefficients:
-$$
-\lambda_{k-1} = \frac{1}{\sqrt{(\conf_{k,0}-\conf_{k-1,0})^T \weight^2 
-(\conf_{k,0}-\conf_{k-1,0})}}, \  k=1\cdots wp+1  
-$$
-aiming at keeping the same ratio between path segment lengths at 
-minimum as at 
-initial path, represented by the waypoints ($\conf_{i,0})_{k=1\cdots wp+1}$.
-Without these coefficients, the path that minimizes the cost corresponds to a 
-straight line with the waypoints equidistantly allocated. This is not relevant for 
-Figure~\ref{local_box_optim} type of problems where a local passage is very
-constrained by obstacles. Note that this cost is also working with all other 
-examples presented in this paper, and provides better quality results than the original cost.
 
@@
 motions in the hole. 
-The GB optimizer manages to modify the puzzle trajectory even in the hole, but remains less efficient than RS and PRS. We explain this behavior by the fact that, ........
+The GB optimizer manages to modify the puzzle trajectory even in the hole, but remains less efficient than RS and PRS to shorten trajectories in the upper and lower parts. 
+\textcolor{blue}{
+This can be the result of adding collision-constraints on these parts of the trajectories, penalizing the global path length.
+One idea could be to 	arbitrarily cancel constraints in upper and lower parts, and keeping the ones in the hole.
+However we want the present GB algorithm to remain general and basic, such constraints relaxation are part of the future possible work.
+}
 %PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole.
@@
 	\includegraphics[width=5.8cm,height=4.7cm]{fiad_real.png}
-	\includegraphics[width=5.8cm,height=4.7cm]{fiad-trajectorires.png}
+	\includegraphics[width=5.8cm,height=4.7cm]{fiad_traj_prune.png}
 	\caption{(Bottom left) An industrial use-case example proposed by Philips for 	
@@
 	constrained in boxes. End-effector trajectories 
-	are illustrated (bottom right): initial (RRT-connect) in red, RS 
-	optimization in blue and the GB optimization in green.}
+	are illustrated (bottom right): initial (RRT-connect) in red, a RS 
+	optimization in blue, a PRS one in cyan and the GB optimization in green.}
 	\label{fiad}
@@
 
-\vspace{0.4cm}
-
-\textcolor{blue}{
-For some of the presented benchmarks, convergence graphs of the path 
-length reduction are 
-given Figure~\ref{fig:remainingLength}. The chosen initial paths are unchanged, 
-e.g. correspond to Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis} . 
-Each graph illustrates the percent ratio of the optimized path length over the initial path length, during the optimization time.
-For the random optimizers curves, averages and standard deviations on 50 
-launches are given. It is not a surprise that GB is globally slower than RS which converges faster. 
-}
+\vspace{0.3cm}
 
@@
   %\textbf{2D slidding-robot (Fig.~\ref{2D_long} right)} & & 11.9 ms & 78.3 & 68.0 & 81.8 \\
-  \textbf{Freeflyer-puzzle} & ($\alpha_{init}=0.1$) & 0 ms & 0 & 0 & 0 \\
+  \textbf{Freeflyer-puzzle} & \todo ($\alpha_{init}=0.1$) & 0 ms & 0 & 0 & 0 \\
   \textbf{Double-arm 4-DOF} & & 39.6 ms & 46.8 & 55.5 & 63.5 \\
   \textbf{UR5-with-spheres} & (RRT, $\alpha_{init}=0.3$) & 3.18 s & 53.8 & 47.5 & 67.8 \\
-  \textbf{UR5-industrial-example} & (RRT) & 0 ms & 0 & 0 & 0 \\
-  \textbf{Baxter-in-office} & & 0 ms & 0 & 0 & 0 \\
+  \textbf{UR5-industrial-example} & \todo (RRT) & 5.50 s & 55.2 & 41.9 & 65.1 \\
+  \textbf{Baxter-in-office} & \todo & 125 s & 26.0 & 36.8 & 74.7 \\
   \textbf{PR2-crossing-arms} & & 882 ms & 19.9 & 43.2 & 95.2 \\
-  \textbf{PR2-in-kitchen-1} & & 0 ms & 0 & 0 & 0 \\
+  \textbf{PR2-in-kitchen-1} & \todo & 12.8 s & 30.6 & 45.3 & 90.9 \\
   \bottomrule
   \end{tabular}  }%scalebox
-\caption{Average results for 50 runs of several examples presented in the 
+\caption{\textcolor{blue}{Average results for 50 runs of several examples presented in the 
 paper. For each run, a solution path is planned by Visibility-PRM 
@@
 %For the PR2-crossing-arms example only, the right column represents the distance that is traveled by the robot mobile base during the different paths.
-}
+}}
 \label{tab:results}
 \end{table}
+
+% convergence graphs:
+\textcolor{blue}{
+For some of the presented benchmarks, convergence graphs of the path 
+length reduction are 
+given Figure~\ref{fig:remainingLength}. The chosen initial paths are unchanged, 
+e.g. correspond to Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis}. 
+Each graph illustrates the percent ratio of the optimized path length over the initial path length, during the optimization time.
+For the random optimizers curves, averages and standard deviations on 50 
+launches are given.
+It is not a surprise that GB is globally slower than RS due to the difference of the computations complexity during the optimization.
+Thus RS converges faster.
+However, it seems that GB catches up and overcomes RS before ending (see Figures~\ref{fig:remainingLengthComp:pr2alone} and~\ref{fig:remainingLengthComp:pr2kitchen1}), thanks to the optimization of the mobile base motion.
+Therefore, it could be interesting to investigate the performance of a composed optimizer, starting by a RS stage until convergence and finishing by a GB stage to improve the path length reduction.
+As previously stated in the freeflyer-example (see Figures~\ref{fig:trajectories:puzzle} and~\ref{fig:remainingLengthComp:puzzle}), GB is not only less efficient than RS and PRS but converges slower. 
+This can be partly explained by the fact that collision checking is rapidly performed in such basic geometry problem. This favors the random shortcut tries while GB spends time on the LCQP resolution.
+}
+
+\vspace{0.3cm}
 
@@
 }
-
 In some cases, $\alpha_{init}$ was 
@@
 }
-Except the low-DOF problem of the puzzle, our method almost always provides the 
-shortest paths, whereas it is not even the best criterion to quantify the parasite 
-motions disappearance.
-TO COMPLETE!
-%Finally, the complete vanishing of the mobile base motion in the GB optimized path is noted. 
-%Thus the issue addressed in Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
+
+\textcolor{blue}{
+The results seems to be consistent with the trajectories analysis and convergence graphs. Except the low-DOF problem of the puzzle, our method provides 
+shorter or similar results compared to RS. Results seems even to be better when the number of DOF increase, as the baxter and PR2 examples.
+Note that the path length does not necessary account for the parasite motions disappearance.
+}
+%Finally, the complete vanishing of the mobile base motion in the GB optimized path is noted. Thus the issue addressed in Figure~\ref{decoupled_DOF_optimization} is solved considering this DOF.
 
 \subsubsection{Analysis of $\alpha_{init}$ influence}
-\textcolor{blue}{Results without comparison to RS and PRS here
-$\alpha_{init}$ could be computed from geometrical considerations.
-}
-TODO !!
-Contrary to a first idea that $\alpha_{init}$ is reduced to 0.05 to slightly modify the very constrained trajectory at each iteration.
-
-\begin{figure}[H]
+
+\begin{figure}
 	\centering
@@
 	}
-  \caption{Influence of $\alpha_{init}$ on the convergence graphs of the GB optimizer.}
+  \caption{\textcolor{blue}{Influence of $\alpha_{init}$ on the convergence graphs of the GB optimizer.
+For each benchmark, the considered intial paths correspond to the ones of Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis}.}}
   \label{fig:remainingLengthAlphaInit}
+\end{figure}
+
+\textcolor{blue}{
+This section deals with the influence of the parameter $\alpha_{init}$ on the 
+GB convergence. Reducing $\alpha_{init}$ makes Algorithm~\ref{algo:gradient} 
+process smaller iterations. Some expected behaviors are visible in 
+Figure~\ref{fig:remainingLengthAlphaInit}. 
+For instance, Figure~\ref{fig:remainingLengthAlphaInit:2Drobot} illustrates an 
+expected influence of a $\alpha_{init}$ reduction on the final path lengths. Besides, commonly to Figures~\ref{fig:remainingLengthAlphaInit:2Drobot}
+,~\ref{fig:remainingLengthAlphaInit:double_arm} 
+and~\ref{fig:remainingLengthAlphaInit:baxter}, $\alpha_{init}=0.05$ has the higher convergence time.
+}
+
+\textcolor{blue}{
+However, due to the strong non-linearity of the constraints, reduced iterations 
+do not necessary lead to a slower but refined solution. GB can stops earlier 
+in local minimum, which may also have a better path length reduction. This is 
+the case of Figure~\ref{fig:remainingLengthAlphaInit:double_arm} where $\alpha_{init}=0.2$ results in a shorter path than $\alpha_{init}=0.05$.
+More surprisingly in Figure~\ref{fig:remainingLengthAlphaInit:double_arm}, $\alpha_{init}=0.05$ leads to the worse reduction.
+}
+
+\textcolor{blue}{
+Instead of investigating a way to find a $\alpha_{init}$ conditioned by the problem and the initial path, we plan to adapt $\alpha$ from Algorithm~\ref{algo:gradient} due to geometrical considerations, as the continuous collision checker.
+}
+
+\subsubsection{Influence of a pruning preliminary step}
+\textcolor{blue}{For the ur5-industrial benchmark, we compared the optimization convergence graphs with and without a pruning step.
+Prune was implemented following~\cite{GeraertsIJRR07}, to remove redundant nodes in the initial path by creating valid shortcuts between the waypoints.
+RRT-connect has been chosen as motion planner because it usually produces more waypoints that Visibility-PRM, so the impact of pruning is more accountable.
+Results are given Figure~\ref{fig:pruningInfluence}.
+The path length reduction of the three optimizers are still compared, but the notable information is the computation time of GB that is 16 times higher without pruning.
+Such lower computation time prevents the random optimizers to converge, so the GB result appears as better.
+}
+
+\textcolor{blue}{Note that, if applying Prune always reduces the GB optimization time, it sometimes spoils the path length reduction.
+In fact, there can be cases where multiple waypoints are useful to bypass an obstacle, rather than a long straight line in the configuration space.
+}
+
+\begin{figure}[H]
+	\centering
+	\subfigure[Without pruning]{
+		\includegraphics[width=6.8cm,height=5.5cm]{fiadRRT_comp.png}
+		\label{fig:pruningInfluence:comp}
+	}
+	\subfigure[With pruning]{
+		\includegraphics[width=6.8cm,height=5.5cm]{fiadRRT_comp_prune.png}
+		\label{fig:pruningInfluence:comp_prune}
+	}
+  \caption{\textcolor{blue}{Influence of a pruning step on the optimization processes.
+RRT-connect provides an initial path with 62 waypoints, which is downed to two waypoints by Prune.
+Concerning the path length, it is only reduced of 6.2\% by Prune.
+Thus, final path lengths provided by GB in both cases are equivalent, the major difference results in the computation time of GB.}}
+  \label{fig:pruningInfluence}
 \end{figure}
@@
 	}
-  \caption{Other trajectories comparisons of end-effectors or mobile bases (initial path in red, RS output in blue and GB output in green).}
+  \caption{Other trajectories comparisons of end-effectors or mobile bases (initial path in red, RS output in blue, PRS output in cyan and GB output in green).}
   \label{fig:trajectoriesbis}
@@
 	}
-  \caption{Convergence graphs of the three optimizers during the optimization process.
+  \caption{\textcolor{blue}{Convergence graphs of the three optimizers during the optimization process.
 For each benchmark, the considered intial paths correspond to the ones of Figures~\ref{fig:trajectories} and~\ref{fig:trajectoriesbis}.
@@
 The dashed blue line is the final result of GB.
-For RS and PRS, the averages and standard deviations (in grey) over 50 launches are plotted.}
+For RS and PRS, the averages and standard deviations (in grey) over 50 launches are plotted.}}
   \label{fig:remainingLength}

### 2016-01-27 paper.tex

@@
 The most common in motion planning are the path length, which penalize detours, the obstacle clearance for safety and the execution time, which is influenced by the number of via points, introducing velocity discontinuities.
-One can also mention the minimization of the acceleration or jerk peaks to implement safely the motion on a robot.}
+One can also mention the minimization of the acceleration or jerk peaks to safely play the motion on a robot.}
 
@@
 The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length expressed as a cost function and avoid collisions with linearized constraints.
-Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}
+Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the algorithm, presented in the last section.}
 
@@
 \end{equation}
-where \textcolor{blue}{$\p_i$ is computed by an unconstrained version of function \texttt{computeIterate}, and} $\alpha_i$ is a real valued parameter. Taking $\alpha_i=1$ yields the 
+where \textcolor{blue}{$\p_i$ is computed by an unconstrained version of function \texttt{computeIterate}, and} $\alpha_i$ is a real-valued parameter. Taking $\alpha_i=1$ yields the 
 unconstrained minimal cost path, i.e. all waypoints aligned on the straight line 
@@
 $$
-\textcolor{blue}{These linear constraints are built from the linearization of a differentiable function $f$, which will be detailed in the following section.}
+\textcolor{blue}{These linear constraints are built from the linearization of a collision-constraint function, which will be detailed in the following section.}
 
@@
 	\caption{Bodies representation in collision-configuration (left) and 
-	backtracked free-configuration (right). A contact point (i.e. any point in 
+	backtracked collision-free configuration (right). A contact point (i.e. any point in 
 	the intersection of the bodies) $\po_c$ can be returned by the Flexible 
-	Collision Library~\cite{fcl}, used to detect collisions. Constraint $f$ defined by~(\ref{eq:f}) aims at keeping $\po_2$ in plane $\Pi$ fixed to $\body_1$.
+	Collision Library~\cite{fcl}, used to detect collisions. 
+	\textcolor{blue}{Constraint defined by~\eqref{eq:new-constraint} aims at 
+	keeping $\po_2(\conf)$ in plane $\Pi$ fixed to $\body_1$.}
 	}
@@
 
-The principle of the method is to compute a \textcolor{blue}{linear} constraint on the initialized on the 
-free-configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
+The principle of the method is to compute a \textcolor{blue}{linear} constraint, 
+initialized on the 
+collision-free configuration $\xx_{i}(\tcolli)$ to avoid the encountered collision $\xx_{i+1}
 (\tcolli)$. To handle this, we introduce a one-dimensional constraint based on the 
@@
 \textcolor{blue}{
-  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
+  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2(\conf)}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
 \begin{equation}\label{eq:g}
@@
 \end {equation}
-aims at keeping point $\po_2$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
+aims at keeping point $\po_2(\conf)$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
 }
@@
 
-Finally, we refer to~\cite{nocedal2006numerical} for solving constrained quadratic 
-programs (\textcolor{blue}{LC}QP). \textcolor{blue}{The step computation corresponds to a constrained 
+Finally, we refer to~\cite{nocedal2006numerical} for solving \textcolor{blue}{LCQP. The step computation corresponds to a constrained 
 version of \texttt{computeIterate}}.
@@
 blocks using the Jacobian $\frac{\partial f}{\partial \conf}$ expressed in 
-each of the two waypoints and $\beta$. This stage \textcolor{blue}{is performed by \texttt{computeCollisionConstraint}}.
+each of the two waypoints and $\beta$. \textcolor{blue}{This stage is performed by \texttt{computeCollisionConstraint}}.
 
@@
 moving from 
-path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
+path $\xx_i$ along step $\p_i$ (see Figure~\ref{fig:convergence-diagram}). Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
 Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
@@
 \end{align}
+\begin{property}\label{prop:geometric-reasoning}
 From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $\po_2$ in reference frame $\body_1$ projected on vector $\U$.
+\end{property}
 }
@@
 	\centering
-	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
-	\caption{\textcolor{blue}{Considering a trajectory 
-	$\traj(t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
-	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
-	and $\kappa$ are respectively collision-free path, in-collision path, 
-	iteration step and collision-abcissa).
-	The trajectory is represented in the configuration space (left) and in the 
-	workspace with the colliding bodies (right).}
-	Note that a case with 
-	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, 
-	so $\p$ already complies 
-	with the constraint that should be added if no redundancy was 
-	tracked. Therefore, 
-	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
-	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
-	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
-	returning the new constraint.}
+	\includegraphics[width=8cm]{convergence-diagram2.pdf}
+	\caption{\textcolor{blue}{Representation in the robot configuration space of the trajectory $\traj$, defined in Equation~\eqref{eq:trajectory}.}}
 	\label{fig:convergence-diagram}
 \end{figure}
+
+\textcolor{blue}{
+Therefore the following expressions
+\begin{align*}
+\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
+\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
+\end{align*}
+correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$.
+Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0.
+According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$.
+Although very unlikely, this case might appear for some new constraint.
+In this case, inserting constraint $L_{j+1}$ is useless since
+$$
+\kernel \Jf_i = \kernel \Jf_{i+1}.
+$$
+}
+
+\subsubsection {Algorithm refinment}
+
+\textcolor{blue}{
+When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent.
+The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy.
+A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop.
+New iterations are chosen in the middle of this pair.
+\begin{itemize}
+\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
+\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
+\end{itemize}
+In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
+\begin{enumerate}
+\item at some point in the loop the new constraint is linearly independent from the previous constraints.
+The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
+\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
+\end{enumerate}
+In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$.
+$\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$.
+Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
+$\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision.
+Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
+}
+
+\textcolor{blue}{
+At each iteration, the new linear constraint
+$$
+L_{j+1} = \frac{\partial f}{\partial\xx} (\xx_{Free})
+$$
+is tested. As iterations $\xx_{Free}$ and $\xx_{Coll}$ tend toward $\bar{\xx}$, it can be established by a geometric reasoning analogous to Property~\ref{prop:geometric-reasoning}, that
+$L_{j+1} \p_i$ tends to the norm of the velocity of point $\bar{\po}$ belonging to $\body_2$ in the frame of $\body_1$. The following property summarizes the result of this section.
+}
+
+\textcolor{blue}{
+\begin{property}\label{prop:convergence}
+As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
+\end{property}
+The above property means that the gradient-based path length optimization algorithm described in this paper converges in all cases, except for ill-defined problems.
+}
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[b]
 \begin{algorithmic}%[1] % line numerotation
@@
 \textcolor{blue}{
-When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent.
-This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}.
+When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent.
 The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy.
@@
 }
-
-\textcolor{red}{Partie en cours de modif par Florent}
 

### 2016-01-28 paper.tex
Fix duplicated parts in 3.6 and typo

@@
 (Figure~\ref{contact-points} left).
-\textcolor{blue}{We denote by
+\textcolor{blue}{We denote by:
   \begin{itemize}
@@
 \end{align*}
-correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$.
-Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0.
-According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$.
-Although very unlikely, this case might appear for some new constraint.
-In this case, inserting constraint $L_{j+1}$ is useless since
-$$
-\kernel \Jf_i = \kernel \Jf_{i+1}.
-$$
-}
-
-\subsubsection {Algorithm refinment}
-
-\textcolor{blue}{
-When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop, performed by Algorithm~\ref{algo:solveRedundant}, in order to find a new constraint that is linearly independent.
-The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy.
-A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop.
-New iterations are chosen in the middle of this pair.
-\begin{itemize}
-\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
-\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
-\end{itemize}
-In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
-\begin{enumerate}
-\item at some point in the loop the new constraint is linearly independent from the previous constraints.
-The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
-\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
-\end{enumerate}
-In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$.
-$\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$.
-Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
-$\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision.
-Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
-}
-
-\textcolor{blue}{
-At each iteration, the new linear constraint
-$$
-L_{j+1} = \frac{\partial f}{\partial\xx} (\xx_{Free})
-$$
-is tested. As iterations $\xx_{Free}$ and $\xx_{Coll}$ tend toward $\bar{\xx}$, it can be established by a geometric reasoning analogous to Property~\ref{prop:geometric-reasoning}, that
-$L_{j+1} \p_i$ tends to the norm of the velocity of point $\bar{\po}$ belonging to $\body_2$ in the frame of $\body_1$. The following property summarizes the result of this section.
-}
-
-\textcolor{blue}{
-\begin{property}\label{prop:convergence}
-As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
-\end{property}
-The above property means that the gradient-based path length optimization algorithm described in this paper converges in all cases, except for ill-defined problems.
-}
-
-\textcolor{blue}{
-Therefore the following expressions
-\begin{align*}
-\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
-\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
-\end{align*}
 correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $\po_2$ in reference frame $\body_1$.
@@
 $$
-\kernel \Jf_i = \kernel \Jf_{i+1}.
-$$
-}
+\kernel \Jf_i = \kernel \Jf_{i+1}
+$$
+}
+
+\subsubsection {Algorithm refinment}
 
@@
 
-\subsubsection {Algorithm refinment}
-
 \textcolor{blue}{
@@
 \item at some point in the loop the new constraint is linearly independent from the previous constraints.
-The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
+The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$, and the loop is interrupted, or
+%the new constraint makes the dimension of the kernel decrease
 \item each new constraint is linearly dependent from the previous constraints and the loop never ends.
@@
 $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$.
-Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
-}
-
-Doing so, either the new constraint makes the dimension of the kernel decrease, or 
-makes the configuration in collision ($\xx_{i+1}$ or $\xx_{i+2}$ depending on the 
-case above) and the latest collision-free path ($\xx_{i}$ or $\xx_{i+2}$) become 
-closer and closer. And the continuity of $f$ implies that $\dot{\conf}(0)$ and 
-$\dot{\conf}(1)$
-converge toward the norm of the velocity of point $\po_2$ in the reference frame of 
-$\body_1$ along trajectory~(\ref{eq:trajectory}), when $\po_2$ and $\po_1$ come to 
-contact. We already handled this situation previously, admitting that the velocity norm 
-would not vanish at $t=1$. Therefore, assumption~(\ref{eq:assumption-1}) is verified.
-%$$
-%\frac{\partial f}{\partial \xx}(\xx_i)\p_i \ \ \ \mbox{and}\ \ \ \frac{\partial f}{\partial \xx}(\xx_{i+1})\p_i
-%$$
+$\bar{\xx}$ necessarily lies at the boundary between free paths and paths in collision.
+Let us denote by $\bar{\kappa}$ the abscissa along $\bar{\xx}$ where $\body_1$ and $\body_2$ come to contact, and let us denote by $\bar{\po}$ the contact point.
+}
+
+\textcolor{blue}{
+At each iteration, the new linear constraint
+$$
+L_{j+1} = \frac{\partial f}{\partial\xx} (\xx_{Free})
+$$
+is tested. As iterations $\xx_{Free}$ and $\xx_{Coll}$ tend toward $\bar{\xx}$, it can be established by a geometric reasoning analogous to Property~\ref{prop:geometric-reasoning}, that
+$L_{j+1} \p_i$ tends to the norm of the velocity of point $\bar{\po}$ belonging to $\body_2$ in the frame of $\body_1$. The following property summarizes the result of this section.
+}
+
+\textcolor{blue}{
+\begin{property}\label{prop:convergence}
+As long as along iteration $\p_i$, the trajectory $(\xx_i+t\alpha_i\p_i)(\bar{\kappa})$ does not enter in collision with contact point velocity equal to 0 in the frame of $\body_1$, Algorithm~\ref{algo:solveRedundant} converges in a finite number of steps.
+\end{property}
+The above property means that the gradient-based algorithm described in this paper converges in all cases, except for ill-defined problems.
+}
 
@@
 
-Figure~\ref{2D_long} shows the result of our optimizer on 2D cases. Contact points which have 
-led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
-Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
-the constraint must be created, since at some parts the robot is very close to obstacles.
-
-Besides, these examples give a better understanding of how the tuning of 
-$\alpha_{init}$ 
-has to balance lot of iterations and relevant collision-constraint addition. For 
-example, if we alter a lot the initial path with a large gradient step and 
-compute the corresponding collisions, constraints will be chosen on a 
-path closer-to-optimality but which is not relevant
-w.r.t. the obstacles we want to avoid.
-
-Figure~\ref{local_box_optim} illustrates a \textit{very long} path example which RS 
-or PRS will not manage to 
-optimize in an affordable time, because of probabilistically failing to sample 
-configurations in the box. The GB method succeeds to optimize the 
-path contained in the box, with the following cost coefficients:
-$$
-\lambda_{k-1} = \frac{1}{\sqrt{(\conf_{k,0}-\conf_{k-1,0})^T \weight^2 
-(\conf_{k,0}-\conf_{k-1,0})}}, \  k=1\cdots wp+1  
-$$
-aiming at keeping the same ratio between path segment lengths at 
-minimum as at 
-initial path, represented by the waypoints ($\conf_{i,0})_{k=1\cdots wp+1}$.
-Without these coefficients, the path that minimizes the cost corresponds to a 
-straight line with the waypoints equidistantly allocated. This is not relevant for 
-Figure~\ref{local_box_optim} type of problems where a local passage is very
-constrained by obstacles. Note that this cost is also working with all other 
-examples presented in this paper, and provides better quality results than the original cost.
-
-\begin{figure}
+\begin{figure}[b]
 	\centering
@@
 \end{figure}
+
+Figure~\ref{2D_long} shows the result of our optimizer on 2D cases. Contact points which have 
+led to constraints are represented. They permit to understand how the path is kept out of the obstacles. 
+Unfortunately, it is not possible in the GB method to choose at which distance of the obstacle 
+the constraint must be created, since at some parts the robot is very close to obstacles.
+
+Besides, these examples give a better understanding of how the tuning of 
+$\alpha_{init}$ 
+has to balance lot of iterations and relevant collision-constraint addition. For 
+example, if we alter a lot the initial path with a large gradient step and 
+compute the corresponding collisions, constraints will be chosen on a 
+path closer-to-optimality but which is not relevant
+w.r.t. the obstacles we want to avoid.
+
+Figure~\ref{local_box_optim} illustrates a \textit{very long} path example which RS 
+or PRS will not manage to 
+optimize in an affordable time, because of probabilistically failing to sample 
+configurations in the box. The GB method succeeds to optimize the 
+path contained in the box, with the following cost coefficients:
+$$
+\lambda_{k-1} = \frac{1}{\sqrt{(\conf_{k,0}-\conf_{k-1,0})^T \weight^2 
+(\conf_{k,0}-\conf_{k-1,0})}}, \  k=1\cdots wp+1  
+$$
+aiming at keeping the same ratio between path segment lengths at 
+minimum as at 
+initial path, represented by the waypoints ($\conf_{i,0})_{k=1\cdots wp+1}$.
+Without these coefficients, the path that minimizes the cost corresponds to a 
+straight line with the waypoints equidistantly allocated. This is not relevant for 
+Figure~\ref{local_box_optim} type of problems where a local passage is very
+constrained by obstacles. Note that this cost is also working with all other 
+examples presented in this paper, and provides better quality results than the original cost.
 

### 2016-01-28 answers-to-reviewers.tex
Update and finish answer letter with missing parts

@@
 \author{Myl\`{e}ne Campana, Florent Lamiraux and Jean-Paul Laumond}
-\date{Answer to reviewers}
+\date{Answer to editor and reviewers}
 \maketitle
@@
 
-\hrulefill\\
-
 \subsection{Comment}
-\hrulefill\\
 
@@
 
-\rule{\linewidth}{.1pt}
+\vspace{0.5cm}
+
 \textbf{Answer}:
@@
 
+\noindent
+\hrulefill\\
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Reviewer 1}
-\label{r2}
-\hrulefill\\
 
 \subsection{Comment 1}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 2}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 3}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 4}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 5}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 6}
-\hrulefill\\
 
@@
 We have rewritten the notations of this part to clarify it.
+We have used a subscript $loc$ to refer the local frames of the points, and a simplified matrix notation for the homogeneous matrix product.
+Thus the expressions of $\mathbf{u}$ and $g$ are clearer.
+
+\noindent
+\hrulefill\\
 
 \subsection{Comment 7}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 8}
-\hrulefill\\
 
@@
 
+\noindent
+\hrulefill\\
+
 \subsection{Comment 9}
-\hrulefill\\
 
@@
 optimization variables so the computation time reduces.
-However, less waypoints results in less possible actions to blend the path.
-In fact, we noted that we obtained shorter paths without the pruning stage.
+However, less waypoints results in less possible actions to blend the path and bypass obstacles.
+In fact, we have noted that we often obtain shorter paths without the pruning stage.
 
-\todo : voir si ajout d'une partie "comparaison avec et sans prune" ....
-globalement, prune ne change pas grand chose a la comparaison avec les autres optim, si ce n'est qu'on a en moyenne des chemins initiaux similaires (peu de waypoints)
+In the article, we have added a brief comparison of two convergence graphs, with and without a preliminary pruning.
+This comparison just relates the gain in terms of computation time.
+We prefer not extend this part, because the main idea of the paper is not to discuss on the relevance of the Prune use.
+
+\noindent
+\hrulefill\\
 
@@
 \subsection{Comment}
-\hrulefill\\
 
@@
 We thank the reviewer for the comment. We removed the reference from the paper.
+
+\noindent
+\hrulefill\\
 
@@
 \subsection{Comment 1}
-\hrulefill\\
 
@@
 
-\rule{\linewidth}{.1pt}
+\vspace{0.5cm}
+
 \textbf{Answer}:
 
-We thank the reviewer for the remark. We have improved the notations of the problem, avoiding to introduce unreferenced objects at the beginning.
+We thank the reviewer for the remark and have removed the unreferenced objects.
+We have improved the introduction of the problem, noting that the QP elements will be defined in the different parts of section 3.
+However, we prefer not give these definitions separately in a few rigorous subparts, as often encountered in the literature.
+We rather explain our motivations to insert constraints while encountering collisions, what type of constraint and how it allows to avoid the collision.
+Same argument is involved for the convergence analysis part.
 
-\todo : adapter / etoffer en fonction de ce que Florent a re-modifie
+\noindent
+\hrulefill\\
 
 \subsection{Comment 2}
-\hrulefill\\
 
@@
 
-\todo : verifier pertinence
+\noindent
+\hrulefill\\
 
 \subsection{Comment 3}
-\hrulefill\\
 
@@
 We want to highlight the fact that we optimize paths on the robot 
-configuration space in a proper mathematical way, specially regarding to the $SO3$ rotations. Thus we prefer to briefly define notations and operators.
+configuration space in a proper mathematical way, specially regarding to the $SO(3)$ rotations. Thus we prefer to briefly define notations and operators.
+
+\noindent
+\hrulefill\\
 
 \subsection{Comment 4}
-\hrulefill\\
 
@@
 
-\rule{\linewidth}{.1pt}
+\vspace{0.5cm}
+
 \textbf{Answer}:
 
-The authors agree with the reviewer that this part should be modified as it is unclear. We add the Figure 5 to help understanding the interest of the 3.6 part.
+The authors agree with the reviewer that this part should be modified as it is unclear.
+We add the Figure 5 to help understanding the interest of the convergence analysis of the 3.6 section.
+This latter section has been rewritten to clarify the link between the contact-point velocity and the constraint redundancy.
+Properties highlight the result of the analysis, that under some assumptions the additional loop in the global algorithm converges.
 
-\todo : adapter en fonction de ce que Florent a re-modifie
+\noindent
+\hrulefill\\
 
 \subsection{Comment 6}
-\hrulefill\\
 
@@
 
-The drawback of this method is that we ignore if the computation time of our method is too important.
-Thus, as suggested, we have represented the convergence curves for some benchmarks, to see if RS and PRS are not converging far rapidly than GB.
+The drawback of this method is that we ignore whether the convergence rate of our method is in the same order of magnitude than RS and PRS.
+Thus, as suggested, we have represented the convergence curves for some benchmarks.
+Averages curves for RS and PRS, rather than just one try as in the trajectories Figures, are provided.
+
+\noindent
+\hrulefill\\
 
 \subsection{Comment 7}
-\hrulefill\\
 
@@
 
-\rule{\linewidth}{.1pt}
+\vspace{0.5cm}
+
 \textbf{Answer}:
 
-As suggested, we have added at the end of the results some graphs for different values of $\alpha_{init}$.
+As suggested, we have added a section in the results parts gathering convergence graphs for different values of $\alpha_{init}$.
 
